Check fluoro data before uploading on EDI
3/28/23
Some annotations are general comments for Phoebe.

```{r}
# import packages
library(tidyverse)
library(deltamapr)
library(readxl)
library(sf)
library(sp)

# --Read in Data--
path <- 'C:/Users/pcarpent/Desktop/Fluroprobe_Data/2020' # manually set path bc Rmd is weird with dirs

df_fl <- list.files(path = 'C:/Users/pcarpent/Desktop/Fluoroprobe_Data/2020',
               pattern = '*.csv') %>% # this part pulls all files in that directory that end in ".csv"
  map_df(~read_csv(paste0(path,.), show_col_types = FALSE)) # this merges them all together into one dataframe

# don't worry about the parsing errors; it imports as expected
```



map_df(~read_csv(paste0(path,.), show_col_types = FALSE)) 

Check: lat/lons make sense (don't worry about this Phoebe, though we can go over it if you want)
```{r}
# import delta sf
sf_delta <- R_EDSM_Subregions_Mahardja

# convert wq to spdf
coords <- df_fl[,c('Latitude','Longitude')]
data   <- subset(df_fl, select = -c(Latitude, Longitude))
crs    <- CRS('+init=epsg:4326 +proj=longlat')
spdf_wq <- SpatialPointsDataFrame(coords = coords,
                               data = data, 
                               proj4string = crs)

# convert delta to spdf
spdf_delta <- as(sf_delta, 'Spatial')
spdf_delta <- spTransform(spdf_delta, CRS('+init=epsg:4326 +proj=longlat'))

# add subregion to df
col_sr <- sp::over(spdf_wq, spdf_delta[,'SubRegion'])
spdf_wq$SubRegion <- col_sr$SubRegion

# convert to shapefile
sf_wq <- st_as_sf(spdf_wq)
sf_wq <- st_transform(sf_wq, st_crs = sf_delta)
sf_wq <- sf_wq %>% filter(!is.na(SubRegion))

# check data
ggplot() +
  geom_sf(data = sf_delta) +
  geom_sf(data = sf_wq, aes(color = SubRegion))
```

Data checks, modified from Phoebe's code.

Primarily checking for extreme outliers or changes in the dataset that point to a QC issue. Best done with timeseries.

Since data has large gaps, it's best to check on a day-by-day basis. 

(Can do this in a for loop/with functions, but won't for learning purposes. Can teach if wanted though.)

*Confluence, Eastern Delta, Grizzly Bay, Sacramento River, San Pablo Bay, Suisun Bay, Western Delta*

*GreenAlgae, Bluegreen, Diatoms, Cryptophyta*
```{r}
# to be slightly fancy, let's set the variables that change up here. that way we don't have to change every instance of them later
region = 'Western Delta'
year = '2022'
month = 'December'

# add DateTime for graphing (eg. create x-axis)
df_fl$DateTime <- as.POSIXct(paste(df_fl$Date, df_fl$Time), format="%m/%d/%Y %H:%M:%S")

# factor month so it's in order
df_fl$Month <- factor(df_fl$Month, levels = month.name)

# best not to reuse names, so renamed these
Region_ug <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% # note that we set these to the objects we defined above 
  select(Month, DateTime, Date, GreenAlgae_2ug_L, Bluegreen_3ug_L, Diatoms_4ug_L, Cryptophyta_5ug_L) %>%
  pivot_longer(cols=c('GreenAlgae_2ug_L', 'Bluegreen_3ug_L', 'Diatoms_4ug_L', 'Cryptophyta_5ug_L'),
                                                 names_to='algae',
                                                 values_to='ug_L')

ggplot(Region_ug, aes(x=DateTime, y=ug_L)) + 
  geom_point(aes(color=algae)) +
  ggtitle(region) + # can use the object here too
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + # this will create a separate graph for each day with their own x-axes
  scale_color_manual(name='type', labels=c('green algae', 'bluegreen', 'diatoms', 'cryptophyta'),
                     values=c('green', 'steelblue', 'darkred', 'violet'))

Region_ml <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% 
  select(Month, DateTime, Date, GreenAlgae_14cells_mL, Bluegreen_15cells_mL, Diatoms_16cells_mL, Cryptophyta_17cells_mL) %>%
  pivot_longer(cols=c('GreenAlgae_14cells_mL', 'Bluegreen_15cells_mL', 'Diatoms_16cells_mL', 'Cryptophyta_17cells_mL'),
                                                 names_to='algae',
                                                 values_to='cells_mL')

ggplot(Region_ml, aes(x=DateTime, y=cells_mL)) + 
  geom_point(aes(color=algae)) +
  ggtitle(region) +
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + 
  scale_color_manual(name='type', labels=c('green algae', 'bluegreen', 'diatoms', 'cryptophyta'),
                     values=c('green', 'steelblue', 'darkred', 'violet'))
```

*Graphing bluegreen on its own*
```{r}
# to be slightly fancy, let's set the variables that change up here. that way we don't have to change every instance of them later
region = 'Western Delta'
year = '2022'
month = 'December'

# add DateTime for graphing (eg. create x-axis)
df_fl$DateTime <- as.POSIXct(paste(df_fl$Date, df_fl$Time), format="%m/%d/%Y %H:%M:%S")

# factor month so it's in order
df_fl$Month <- factor(df_fl$Month, levels = month.name)

# best not to reuse names, so renamed these
Region_ug <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% # note that we set these to the objects we defined above 
  select(Month, DateTime, Date, Bluegreen_3ug_L) %>%
  pivot_longer(cols=c('Bluegreen_3ug_L'),
                                                 names_to='algae',
                                                 values_to='ug_L')

ggplot(Region_ug, aes(x=DateTime, y=ug_L)) + 
  geom_point(aes(color=algae)) +
  ggtitle(region) + # can use the object here too
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + # this will create a separate graph for each day with their own x-axes
  scale_color_manual(name='type', labels=c('bluegreen'),
                     values=c('steelblue'))

Region_ml <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% 
  select(Month, DateTime, Date, Bluegreen_15cells_mL) %>%
  pivot_longer(cols=c('Bluegreen_15cells_mL'),
                                                 names_to='algae',
                                                 values_to='cells_mL')

ggplot(Region_ml, aes(x=DateTime, y=cells_mL)) + 
  geom_point(aes(color=algae)) +
  ggtitle(region) +
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + 
  scale_color_manual(name='type', labels=c('bluegreen'),
                     values=c('steelblue'))
```

*Graphing temperature*
```{r}
# to be slightly fancy, let's set the variables that change up here. that way we don't have to change every instance of them later
region = 'Western Delta'
year = '2022'
month = 'December'

# add DateTime for graphing (eg. create x-axis)
df_fl$DateTime <- as.POSIXct(paste(df_fl$Date, df_fl$Time), format="%m/%d/%Y %H:%M:%S")

# factor month so it's in order
df_fl$Month <- factor(df_fl$Month, levels = month.name)

# best not to reuse names, so renamed these
Region_ug <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% # note that we set these to the objects we defined above 
  select(Month, DateTime, Date, Temp_degC) %>%
  pivot_longer(cols=c('Temp_degC'),
                                                 names_to='temperature',
                                                 values_to='deg_C')

ggplot(Region_ug, aes(x=DateTime, y=deg_C)) + 
  geom_point(aes(color=temperature)) +
  ggtitle(region) + # can use the object here too
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + # this will create a separate graph for each day with their own x-axes
  scale_color_manual(name='type', labels=c('bluegreen'),
                     values=c('darkred'))

Region_ml <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% 
  select(Month, DateTime, Date, Temp_degC) %>%
  pivot_longer(cols=c('Temp_degC'),
                                                 names_to='temperature',
                                                 values_to='deg_C')

ggplot(Region_ml, aes(x=DateTime, y=deg_C)) + 
  geom_point(aes(color=temperature)) +
  ggtitle(region) +
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + 
  scale_color_manual(name='type', labels=c('temperature'),
                     values=c('darkred'))
```




*Graphing fluorescence*
```{r}
# to be slightly fancy, let's set the variables that change up here. that way we don't have to change every instance of them later
region = 'Western Delta'
year = '2022'
month = 'December'

# add DateTime for graphing (eg. create x-axis)
df_fl$DateTime <- as.POSIXct(paste(df_fl$Date, df_fl$Time), format="%m/%d/%Y %H:%M:%S")

# factor month so it's in order
df_fl$Month <- factor(df_fl$Month, levels = month.name)

# best not to reuse names, so renamed these
Region_ug <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% # note that we set these to the objects we defined above 
  select(Month, DateTime, Date, FLUORRFU_RFU) %>%
  pivot_longer(cols=c('FLUORRFU_RFU'),
                                                 names_to='relative_fluorescence',
                                                 values_to='RFU')

ggplot(Region_ug, aes(x=DateTime, y=RFU)) + 
  geom_point(aes(color=relative_fluorescence)) +
  ggtitle(region) + # can use the object here too
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + # this will create a separate graph for each day with their own x-axes
  scale_color_manual(name='type', labels=c('relative_fluorescence'),
                     values=c('green'))

Region_ml <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% 
  select(Month, DateTime, Date, FLUOR_ug_L) %>%
  pivot_longer(cols=c('FLUOR_ug_L'),
                                                 names_to='fluorescence',
                                                 values_to='ug_L')

ggplot(Region_ml, aes(x=DateTime, y=ug_L)) + 
  geom_point(aes(color=fluorescence)) +
  ggtitle(region) +
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + 
  scale_color_manual(name='type', labels=c('fluorescence'),
                     values=c('green'))


```





*Graphing turbitidy*

```{r}
# to be slightly fancy, let's set the variables that change up here. that way we don't have to change every instance of them later
region = 'Confluence'
year = '2021'
month = 'September'

# add DateTime for graphing (eg. create x-axis)
df_fl$DateTime <- as.POSIXct(paste(df_fl$Date, df_fl$Time), format="%m/%d/%Y %H:%M:%S")

# factor month so it's in order
df_fl$Month <- factor(df_fl$Month, levels = month.name)

# best not to reuse names, so renamed these
Region_ug <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% # note that we set these to the objects we defined above 
  select(Month, DateTime, Date, 'NTU_NTU') %>%
  pivot_longer(cols=c('NTU_NTU'),
                                                 names_to='turbidity',
                                                 values_to='NTU')

ggplot(Region_ug, aes(x=DateTime, y=NTU)) + 
  geom_point(aes(color=turbidity)) +
  ggtitle(region) + # can use the object here too
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + # this will create a separate graph for each day with their own x-axes
  scale_color_manual(name='type', labels=c('turbidity'),
                     values=c('darkorange'))

```

```{r}
# to be slightly fancy, let's set the variables that change up here. that way we don't have to change every instance of them later
region = 'Western Delta'
year = '2022'
month = 'December'

# add DateTime for graphing (eg. create x-axis)
df_fl$DateTime <- as.POSIXct(paste(df_fl$Date, df_fl$Time), format="%m/%d/%Y %H:%M:%S")

# factor month so it's in order
df_fl$Month <- factor(df_fl$Month, levels = month.name)

# best not to reuse names, so renamed these
Region_ug <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% # note that we set these to the objects we defined above 
  select(Month, DateTime, Date, 'FNU_FNU') %>%
  pivot_longer(cols=c('FNU_FNU'),
                                                 names_to='turbidity',
                                                 values_to='FNU')

ggplot(Region_ug, aes(x=DateTime, y=FNU)) + 
  geom_point(aes(color=turbidity)) +
  ggtitle(region) + # can use the object here too
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + # this will create a separate graph for each day with their own x-axes
  scale_color_manual(name='type', labels=c('turbidity'),
                     values=c('darkorange'))

```


*Graphing pH*

```{r}
# to be slightly fancy, let's set the variables that change up here. that way we don't have to change every instance of them later
region = 'Western Delta'
year = '2022'
month = 'December'

# add DateTime for graphing (eg. create x-axis)
df_fl$DateTime <- as.POSIXct(paste(df_fl$Date, df_fl$Time), format="%m/%d/%Y %H:%M:%S")

# factor month so it's in order
df_fl$Month <- factor(df_fl$Month, levels = month.name)

# best not to reuse names, so renamed these
Region_ug <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% # note that we set these to the objects we defined above 
  select(Month, DateTime, Date, pH) %>%
  pivot_longer(cols=c('pH'),
                                                 names_to='pH_value',
                                                 values_to='pH')

ggplot(Region_ug, aes(x=DateTime, y=pH)) + 
  geom_point(aes(color=pH_value)) +
  ggtitle(region) + # can use the object here too
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + # this will create a separate graph for each day with their own x-axes
  scale_color_manual(name='type', labels=c('pH'),
                     values=c('darkred'))

```





*Graphing specific conductance*

```{r}
# to be slightly fancy, let's set the variables that change up here. that way we don't have to change every instance of them later
region = 'Western Delta'
year = '2022'
month = 'December'

# add DateTime for graphing (eg. create x-axis)
df_fl$DateTime <- as.POSIXct(paste(df_fl$Date, df_fl$Time), format="%m/%d/%Y %H:%M:%S")

# factor month so it's in order
df_fl$Month <- factor(df_fl$Month, levels = month.name)

# best not to reuse names, so renamed these
Region_ug <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% # note that we set these to the objects we defined above 
  select(Month, DateTime, Date, SPC_uS_cm) %>%
  pivot_longer(cols=c('SPC_uS_cm'),
                                                 names_to='specific conductance',
                                                 values_to='uS_cm')

ggplot(Region_ug, aes(x=DateTime, y=uS_cm)) + 
  geom_point(aes(color='specific conductance')) +
  ggtitle(region) + # can use the object here too
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + # this will create a separate graph for each day with their own x-axes
  scale_color_manual(name='type', labels=c('specific conductance'),
                     values=c('darkred'))

```




*Graphing dissolved oxygen*
```{r}
# to be slightly fancy, let's set the variables that change up here. that way we don't have to change every instance of them later
region = 'Western Delta'
year = '2022'
month = 'December'

# add DateTime for graphing (eg. create x-axis)
df_fl$DateTime <- as.POSIXct(paste(df_fl$Date, df_fl$Time), format="%m/%d/%Y %H:%M:%S")

# factor month so it's in order
df_fl$Month <- factor(df_fl$Month, levels = month.name)

# best not to reuse names, so renamed these
Region_ug <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% # note that we set these to the objects we defined above 
  select(Month, DateTime, Date, 'DOSAT_%SAT') %>%
  pivot_longer(cols=c('DOSAT_%SAT'),
                                                 names_to='dissolved oxygen',
                                                 values_to='percent_saturation')

ggplot(Region_ug, aes(x=DateTime, y=percent_saturation)) + 
  geom_point(aes(color='dissolved oxygen')) +
  ggtitle(region) + # can use the object here too
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + # this will create a separate graph for each day with their own x-axes
  scale_color_manual(name='type', labels=c('dissolved oxygen'),
                     values=c('green'))

Region_ml <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% 
  select(Month, DateTime, Date, DO_mg_L) %>%
  pivot_longer(cols=c('DO_mg_L'),
                                                 names_to='dissolved oxygen',
                                                 values_to='mg_L')

ggplot(Region_ml, aes(x=DateTime, y=mg_L)) + 
  geom_point(aes(color='dissolved oxygen')) +
  ggtitle(region) +
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + 
  scale_color_manual(name='type', labels=c('dissolved oxygen'),
                     values=c('green'))


```




*Graphing water temperature*
```{r}
# to be slightly fancy, let's set the variables that change up here. that way we don't have to change every instance of them later
region = 'Western Delta'
year = '2022'
month = 'August'

# add DateTime for graphing (eg. create x-axis)
df_fl$DateTime <- as.POSIXct(paste(df_fl$Date, df_fl$Time), format="%m/%d/%Y %H:%M:%S")

# factor month so it's in order
df_fl$Month <- factor(df_fl$Month, levels = month.name)

# best not to reuse names, so renamed these
Region_ug <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% # note that we set these to the objects we defined above 
  select(Month, DateTime, Date, WT_C) %>%
  pivot_longer(cols=c('WT_C'),
                                                 names_to='water temperature',
                                                 values_to='degC')

ggplot(Region_ug, aes(x=DateTime, y=degC)) + 
  geom_point(aes(color='water temperature')) +
  ggtitle(region) + # can use the object here too
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + # this will create a separate graph for each day with their own x-axes
  scale_color_manual(name='type', labels=c('water temperature'),
                     values=c('aquamarine3'))

```





*Graphing PAR1 and PAR2*
```{r}
# to be slightly fancy, let's set the variables that change up here. that way we don't have to change every instance of them later
region = 'Confluence'
year = '2020'
month = 'December'

# add DateTime for graphing (eg. create x-axis)
df_fl$DateTime <- as.POSIXct(paste(df_fl$Date, df_fl$Time), format="%m/%d/%Y %H:%M:%S")

# factor month so it's in order
df_fl$Month <- factor(df_fl$Month, levels = month.name)

?drop_na

# best not to reuse names, so renamed these
Region_ug <- df_fl %>% 
  drop_na(`PAR1_umol_m-2_s-1`) %>%
  filter(Region == region, Year == year, Month == month) %>%
# note that we set these to the objects we defined above 
  select(Month, DateTime, Date, `PAR1_umol_m-2_s-1`) %>%
  pivot_longer(cols=c(`PAR1_umol_m-2_s-1`),
                                                 names_to='PAR1',
                                                 values_to='umol_smm')

ggplot(Region_ug, aes(x=DateTime, y=umol_smm)) + 
  geom_point(aes(color=PAR1)) +
  ggtitle(region) + # can use the object here too
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + # this will create a separate graph for each day with their own x-axes
  scale_color_manual(name='type', labels=c('PAR1'),
                     values=c('darkseagreen3'))

Region_ml <- df_fl %>% 
  filter(Region == region, Year == year, Month == month) %>% 
  select(Month, DateTime, Date, 'PAR2_umol_m-2_s-1') %>%
  pivot_longer(cols=c('PAR2_umol_m-2_s-1'),
                                                 names_to='PAR2',
                                                 values_to='umol_smm')

ggplot(Region_ml, aes(x=DateTime, y=umol_smm)) + 
  geom_point(aes(color='PAR2')) +
  ggtitle(region) +
  facet_wrap(. ~ Date, scales='free_x', ncol = 2) + 
  scale_color_manual(name='type', labels=c('PAR2'),
                     values=c('darkseagreen3'))


```







Create similar graphs for all other variables in dataset, including WQ (to check that they were integrated correctly)
```{r}

```

