# Goal:
Excel notebook with 4 sheets
- Lab
- Blanks
- Duplicates
- Field
# Part 1: Read in and Clean Data

# TODO: found out the old ... columns were real data, need to fix
```{r message = FALSE}
# import packages
library(tidyverse)
library(readxl)

# import data
df_raw <- read_excel('01_WaterQuality/wq-qaqc/raw-data/test-import.xlsx', skip = 1) # skip initial row; can do "1" because we know with 100% certainty it's always the first row

df_raw <- df_raw %>% select(-contains('...')) # remove columns containing '...'; know with 100% certainty that those columns are blanks

df_raw <- df_raw %>% 
  mutate(id = row_number())
```
```{r}
str(df_raw)
```

# Part 2: Split Field and Lab

```{r}
vari_id <- df_id$id[df_id$`Station Name` == '<<<Field Results>>>'][[1]]

df_field <- df_raw %>% subset(id > vari_id) # vari_id is id element for the row that starts the field data
df_lab <- df_raw %>% subset(id < vari_id)
```


Now that we have field and lab, let's look at them.

**For Lab**
- think about how to deal with lab dupes
    - do the RPD calcs in the same sheet? add them to their own sheet? do them in a separate column?
    
    - currently, add another row below the parent and the dupe and have it be the RPD
    
    
- fix up names (uploaded csvs)

**For Field**
- first row is actually column names
- rename columns

## Reassign column names
```{r}
# assign column names to the first row of the dataframe
colnames(df_field) <- df_field[1,]
```

## Rename column headers
```{r}
df_headers <- read_csv('01_WaterQuality/wq-qaqc/WDL_Column_Headers.csv')
df_headers <- df_headers %>% select(-contains('...')) 

# we know in this case that the order is correct, but need to double check
# we are NOT FUTURE PROOFING this code right now

# if you want to turn a tibble into a vector, use [[1]]
colnames(df_field) <- df_headers[,2][[1]]
```

```{r}

```


# Part 3: Fix up Field Data


# Part 4: Fix up Lab Data


# Part 5: Recombine Data


# Part 6: Split out Blanks


# Part 7: Split out Duplicates
