# Goal/General Plan:

Excel notebook with 4 sheets

Part 1: - Split Lab and Field

Part 2: - Fix up Lab and Field

Part 3: - Merge back together

Part 4: - Split blanks and duplicates

## Specifics

-   Fix up Lab

    -   delete columns with `…` in names

-   Fix up Field

    -   Splice based on `Station Name == '<< Field Results >>>'`

    -   Rename columns

    -   Reorder columns

-   *Blanks*

    -   *Splice based on `Sample Type`*

-   *Duplicates*

    -   *Splice based on `Sample Type`*
    -   *exists in both field and lab*
        -   *for lab, just chla volume*
            -   *add this column to main duplicate df*

## Part 1: Import Data

Let's start with our original code.

```{r message = FALSE}
# # import packages
# library(tidyverse)
# library(readxl)
# 
# # import data
# df_raw <- read_excel('01_WaterQuality/wq-qaqc/raw-data/test-import.xlsx', skip = 1) # skip initial row; can do "1" because we know with 100% certainty it's always the first row
# 
# df_raw <- df_raw %>% select(-contains('...')) # remove columns containing '...'; know with 100% certainty that those columns are blanks
# 
# df_test <- df_raw %>% subset(`Sample Type` == 'Blank; Equipment') # filter by Blanks
# 
# str(df_raw)
```

Oh no! We messed up! We actually need those "..." columns.

Let's not do that then, and just do a straight import. Then, we can move into splitting up field and lab.

```{r message = FALSE}
# import packages
library(tidyverse)
library(readxl)

# import data
df_raw <- read_excel('01_WaterQuality/wq-qaqc/raw-data/test-import.xlsx', skip = 1) # skip initial row; can do "1" because we know with 100% certainty it's always the first row
```

# Part 2: Split Field and Lab

```{r}
df_raw <- df_raw %>%
  mutate(id = row_number()) %>% # add index column
  relocate(id, .before = `Station Name`) # relocate to before Station Name, which we're 99% certain will be the first col

mid_idrow <- df_raw %>% subset(`Station Name` == '<<<Field Results>>>')

vari_id <- mid_idrow$id

df_field <- df_raw %>% subset(id > vari_id)
df_lab <- df_raw %>% subset(id < vari_id)
```

# Part 3: Fix up Lab Data

```{r}
# remove columns containing '...'; know with 100% certainty that those columns are blanks
# also remove id column
# (when removing more than one thing, they need to be put into a vector `c()`)
# to remove, use `-`
df_lab <- df_lab %>% select(-c(id,contains('...'))) 
```

# Part 4: Fix up Field Data

## Step 1: Remove extra rows of df_field
```{r}
# remove rows in df_field that we don't need (ie. the bottom 4 after `Station Name` == NA)
# use same process that we used to subset df_lab and df_field

field_idrow <- df_field %>% subset(is.na(df_field$`Station Name`))

vari_id <- field_idrow$id

df_field <- df_field %>% subset(id < vari_id)
```


## Step 2: Rename columns
```{r}
# We are going to replace ("map") the column headers of 'df_field' with the 'Field New' column of 'df_fieldnames'
# This assumes the columns in 'df_field' are ordered in same way as the rows for the 'Field New' column in 'df_fieldnames'.

# import field colnames csv
df_fieldnames <- read_csv('01_WaterQuality/wq-qaqc/WDL_Column_Headers.csv')

# get rid of ID column because 'df_fieldnames' doesn't have it
df_field <- df_field %>%
  select(-'id')

# "map" df_fieldnames$`Field New` onto the colnames for df_field
colnames(df_field) <- df_fieldnames$Field_New

# remove first row; always will be old column names
df_field <- df_field %>% slice(-1)
```

## Step 3: Reorder columns
```{r}
# import field order csv
df_fieldorder <- read_csv('01_WaterQuality/wq-qaqc/WDL_Column_Order.csv')

# select 'New Order' column from df_fieldorder
vec_fieldorder <- df_fieldorder$New_Order

# rebuild df_field with columns in the correct order
df_test <- df_field %>% select(vec_fieldorder)    

# check that the column names for df_field and df_test are the same 
colnames(df_field) %in% colnames(df_test)

# if true, rename df_test to df_field
df_field <- df_test  
```

# Part 4: Recombine Data

# Part 5: Split out Blanks

# Part 6: Split out Duplicates
