Title:
Author:
Description:
Date Modified:


```{r}
# import packages
library(tidyverse)
library(readxl)

# import data
df_raw <- read_excel('01_WaterQuality/wq-qaqc/raw-data/test-import.xlsx', skip = 1) # skip lets us skip the first row, which is just the lab data line
```
Oh no! It imported blank columns, which we don't want. So, we must **subset** the data.
Google has told us to use subset :)

```{r}
colnames(df_raw)
```
Any function used after the pipe will use the dataframe (or object) before the pipe.
```{r}
df_raw <- df_raw %>% select(-contains('...'))
```


# Subset blanks!
We see that blanks are indicated in the "Sample Type" column, so we can use that.
First, lets double check that there's no typos or anything else that could be "blank" in that column

```{r}
unique(df_raw$`Sample Type`)
```
Hmmm, that sample type is bad. Also the NAs?
We checked, we understand, it's the extra columns. We will ignore that for now.

```{r}
df_test <- df_raw %>% subset(`Sample Type` == 'Blank; Equipment')
```

It works, but the field data is really messing things up :(

```{r}
print('hi')

print('second hi')
```

