---
---
---

Format Field Data Sheets

INSTRUCTIONS:

-   run all code

-   edit values pre-defined sections

-   if checks are failed, fix in archived pdf and then run code again

# Set Up

## import packages

```{r}
# import packages
library(reticulate)
library(tidyverse)
library(lubridate)
library(janitor)
library(readxl)
source('00_GlobalFunctions/functions.R')
source('01_WaterQuality/database-migration/03_future-formatting/import-functions.R')
```

## EDIT: define paths

-   set `fp_rel` to the file path of the datasheets

    -   the "start" folder is the EMP SharePoint, so write out the path relative to that

-   set `fp_export` to the file path the csv should export to

    -   same as above, the "start" folder is the EMP SharePoint

```{r}
# path to main folder
fp_rel <- 'Water Quality/00 - Monthly Run Docs/Archived Field Data Sheets/2025/10 - October/'

# path to export folder
fp_export <- 'Water Quality/00 - Monthly Run Docs/Database Import Data/2025/'
```

## run file extraction code

```{python}
from os import environ, walk
from os.path import join, normpath
import fitz
import re
import pandas as pd
from datetime import datetime
from typing import List, Tuple, Dict, Optional
```

```{python}
exec(open('01_WaterQuality/database-migration/03_future-formatting/import_functions_py.py').read())

# --- get paths/files ---
fp_rel = r.fp_rel

fp_main = abs_path_emp(fp_rel)

# --- combine files ---
pdf_files = []
for root, dirs, files in os.walk(fp_main):
    for file in files:
        if (
            file.lower().endswith('.pdf') and
            'field data sheet' in file.lower()
        ):
            full_path = os.path.join(root, file)
            pdf_files.append(full_path)

# --- create df ---
df = structure_pdf_data(pdf_files[0])

df['value'] = df['value'].apply(to_scalar_string).astype('string')
```

```{r}
# write back to R variable
df <- py$df
```

# Format data

-   will give a warning if multiple entries are detected for duplicate/blank data in a run
-   will tell if any stations were excluded (eg. P8_DO)

```{r}
df_final <- format_FDS_pdf(df)
```

# Checks

## stations exist

check if stations are missing from pre-defined station list

-   can update file here

```{r}
# check stations
check_stations(df_final)
```

## extra data

check there's no extra data (ie. duplicates)

```{r}
# check extra data
check_extra_data(df_final)
```

## filter ID formatting

check filter IDs are all formatted correctly (one capital letter)

```{r}
# check filter container ID
check_filter_ID(df_final)
```

## analytes exist

check all expected analyte/station pairs exist

-   can update expected analytes here

special rules:

-   shore stations don't require Secchi Depth

-   equipment blank only has Churn Bucket and Filter Container ID

-   replicate only has Churn Bucket and Filter Container ID

-   LSZ stations have Latitude and Longitude

-   only D7, D16, and D19 have Dissolved Bromide

```{r}
# check analytes
check_analytes(df_final, type = 'field')
```

## analyte formatting

check analytes have correct format (eg. numeric vs text)

# Export

```{r}
# Export by month
fp_export <- abs_path_emp(fp_export)

df_list <- split(df_final,
                 list(format(as.POSIXct(df_final$`Observed DateTime`), '%Y-%m')),
                 drop = TRUE)

invisible(lapply(names(df_list), function(name) {
  name_parts <- unlist(strsplit(name, '\\.'))
  year_month <- name_parts[1]

  fp_exp <- paste0(fp_export, '/', year_month,'_FDS_EMP-DWQ.csv')
  write_csv(df_list[[name]], file = fp_exp, na = '')
}))
```
