# Format EMP phyto data for EDI publication
## Updated: 06/2023
## contact: Sarah Perry (sarah.perry@water.ca.gov)

=====
EDIT HERE
=====
```{r}
# filepath for most recent phyto data (must be linked to SharePoint)
cur_phyto <- '2022 Annual Report Data/2022_phyto_raw.csv'

# website path of most recent EDI publication
edi_phyto <- 'https://portal.edirepository.org/nis/dataviewer?packageid=edi.1320.4&entityid=1eee2c2a562a5b856398082c487dc1a7'
```


=====
CODE STARTS HERE
=====
```{r}
library(tidyverse)
source('03_Phyto/PESP/functions/phyto_funcs.R')
```

Read in Data
```{r message = FALSE}
# read in data
df_edi <- read_csv('03_Phyto/PESP/data/EMP/phytoedi_2022_wcoords.csv')
df_edi <- df_edi %>% relocate(c(WestLong, NorthLat), .after = StationCode)

df_cur <- read_csv(abs_path_ar(cur_phyto))

# synonym filepath
df_syn <- read_csv('03_Phyto/PESP/data/EMP/Phyto Classification.csv', show_col_types = FALSE) # synonym list for comparison
```

Basic Cleaning
```{r}
# remove unnecessary columns/rename
df_cur <- df_cur %>%
  select(-c(Taxonomist, `Full Code`, `BSA#`, BSA_TIN, `Percent of Sample Counted`)) %>%
  rename('Name' = 'Taxon',
         Colony_or_FilamentGroupCode = `Colony/Filament/Individual Group Code`,
         Total_Cells_Counted = `Total Number of Cells`,
         UnitAbundance = `Unit Abundance (# of Natural Units)`)

# fix dates/times, colony codes
df_cur$Colony_or_FilamentGroupCode[df_cur$Colony_or_FilamentGroupCode == 'n/p'] <- NA_character_
df_cur$SampleDate <- as.Date(df_cur$SampleDate, format = '%m/%d/%Y')
# df_cur<- df_cur %>%
#   mutate(
#     SampleTime = case_when(grepl(':', SampleTime) ~ SampleTime,
#                            is.na(SampleTime) ~ NA_character_,
#                            TRUE ~ paste0(substr(SampleTime, 1, nchar(SampleTime)-2),':',substr(SampleTime, nchar(SampleTime)-1, nchar(SampleTime)))
#                            )
#    )
```

Standardize dataset
  -Standardize station names
```{r}
df_cur <- df_cur %>%
  mutate(
    StationCode = case_when(StationCode == 'C3A-Hood' | StationCode == 'C3A - Hood' | StationCode == 'C3A Hood' ~ 'C3A',
                            StationCode == 'D16 - Twitchell' | StationCode == 'D16-Twitchell' |StationCode == 'D16 Twitchell' ~ 'D16',
                            StationCode == 'E22' ~ 'EZ2',
                            StationCode == 'E26' ~ 'EZ6',
                            StationCode == 'EZ2 SJR' | StationCode == 'EZ2 - SJR' ~ 'EZ2-SJR',
                            StationCode == 'EZ6 SJR' | StationCode == 'EZ6 - SJR' ~ 'EZ6-SJR',
                            TRUE ~ StationCode)
    )

unique(df_cur$StationCode)
```

standardize "unknown" identifiers
```{r}
unknown_syns <- 'unknown|unidentified|Unidentified|Undetermined|undetermined'

df_cur <- df_cur %>%
  # update Taxon column to standardize Unknown
  mutate(
    Name = case_when(grepl(unknown_syns,Name,ignore.case = TRUE) ~ str_replace_all(Name, unknown_syns, 'Unknown'),
                      TRUE ~ Name)
    ) %>%
  # Update Genus column if unknown Species
  mutate(
    Genus = case_when(grepl('Unknown', Name) ~ 'Unknown',
                      is.na(Genus) ~ 'Unknown',
                      Genus == 'Other' ~ 'Unknown',
                      Genus == 'genus' ~ 'Unknown',
                      TRUE ~ Genus)
    ) %>%
  # Update Species column in unknown
  mutate(
      Species = case_when(Genus == 'Unknown' ~ 'Unknown',
                          is.na(Species) ~ 'Unknown',
                          TRUE ~ Species)
  )
```

### Add in higher level taxa/current names
```{r}
# subset synonym/taxon columns
df_syn <- df_syn %>% select(c('Kingdom':'AlgalGroup','Name','CurrentName'))

# standardize spp. to sp.
df_cur$Name <- str_replace_all(df_cur$Name, 'spp.', 'sp.')
df_cur$Species <- str_replace_all(df_cur$Species, 'spp.', 'sp.')

# standardize unknown names before synonym join
df_cur <- df_cur %>%
  mutate(
    Name = case_when(Name == 'LGBs' ~ 'Little Green Balls',
                     Name == 'Unknown Banana Blue Green' | Name == 'Unknown Cyanobacteria sp.' ~ 'Unknown cyanobacterium',
                     Name == 'Unknown centrales sp.' | Name == 'Unknown centric sp.' | Name == 'Unknown Centric diatom'  ~ 'Unknown centric diatom',
                     Name == 'Unknown Chlorophyte alga sp.' | Name == 'Unknown Chlorophyte filament' ~ 'Unknown green alga',
                     Name == 'Unknown dinoflagellate sp.' | Name == 'Unknown Dinoflagellate sp.' | Name == 'small dinoflagellates'~ 'Unknown dinoflagellate',
                     Name == 'Unknown girdle sp.' | Name == 'Unknown pennales sp.' | Name == 'Unknown pennate girdle sp.' ~ 'Unknown pennate diatom',
                     Name == 'Small nano-flagellates' | Name == 'small nano-flagellates' ~ 'Unknown nanoflagellate',
                     Name == 'Unknown Euglenoids' ~ 'Unknown euglenoid',
                     Name == 'Unknown Genus' ~ 'Unknown genus',
                     TRUE ~ Name)
    )

# join with main dataset
df_cur_syn <- df_cur %>%
  left_join(df_syn, by = 'Name') %>%
  # mutate(Genus = Genus.y) %>%
  select(-c(ends_with('.y'), ends_with('.x'))) %>%
  relocate(c(Name, Kingdom:AlgalGroup), .after = StationCode) %>%
  relocate(c(Genus, Species), .after = AlgalGroup)
```

Update Names with Synonyms
```{r}
df_cur_syn <- df_cur_syn %>%
  mutate(
    Genus = case_when(Name == CurrentName | grepl('None|Unknown',CurrentName) | is.na(CurrentName) ~ Genus,
                      TRUE ~ str_remove(str_squish(str_remove(CurrentName, 'cf.')), ' .*')),
    Species = case_when(Name == CurrentName | grepl('None|Unknown',CurrentName) | is.na(CurrentName) ~ Species,
                        TRUE ~ str_remove(str_squish(str_remove(CurrentName, 'cf.')), '.*? ')),
    OrigName = case_when(Name == CurrentName | grepl('None|Unknown',CurrentName) | is.na(CurrentName) ~ NA_character_,
                         TRUE ~ Name),
    Name = case_when(grepl('None|Unknown',CurrentName) | is.na(CurrentName) ~ Name,
                      TRUE ~ CurrentName
                      )) %>%
  select(-CurrentName) %>%
  relocate(OrigName, .after = StationCode)

# standardize specific unknown name cases
df_cur_syn <- df_cur_syn %>%
  mutate(
    AlgalGroup = case_when(Name == 'Little Green Balls' ~ 'Unknown',
                           Name == 'Unknown cyanobacteria' ~ 'Cyanobacteria',
                           Name == 'Unknown centric diatom' ~ 'Centric Diatoms',
                           Name == 'Unknown green alga' ~ 'Green Algae',
                           Name == 'Unknown dinoflagellate' ~ 'Dinoflagellates',
                           Name == 'Unknown pennate diatom' ~ 'Pennate Diatoms',
                           Name == 'Unknown nanoflagellate' ~ 'Nanflagellates',
                           Name == 'Unknown euglenoid' ~ 'Euglenoids',
                           TRUE ~ AlgalGroup)
  )
```

Check taxa for errors
```{r}
df_error_check <- df_cur_syn %>% select(c('Name':'Species'))

check <- df_error_check %>% subset(is.na(AlgalGroup) | is.na(Class) | is.na(Phylum) | is.na(Kingdom) | is.na(Genus) | is.na(Species))
check <- check[!duplicated(check),]
check <- check %>% arrange(Name)
syn_check <- df_error_check %>% subset(is.na(Name) | Name == 'Unknown')
syn_check <- syn_check[!duplicated(syn_check),]
syn_check <- syn_check %>% arrange(Name)

print(check)
print(syn_check)
```

```{r}
df_cur_syn$Total_Cells_Counted <- as.numeric(df_cur_syn$Total_Cells_Counted)

df_cur_syn <- df_cur_syn %>%
  mutate(AverageBiovolume = round(rowMeans(select(., starts_with('Biovolume')), na.rm = TRUE),2),
         AverageBiovolume_per_mL = round(AverageBiovolume*Factor*Total_Cells_Counted,2)) %>% # Counted'
  select(-starts_with('Biovolume_')) %>%
  relocate(Comments, .after = Shape) #%>%
  # relocate(AverageBiovolume:AverageBiovolume_per_mL, .after = Organisms_per_mL)

df_cur_syn$AverageBiovolume[is.nan(df_cur_syn$AverageBiovolume)] <- NA
df_cur_syn$AverageBiovolume_per_mL[is.nan(df_cur_syn$AverageBiovolume_per_mL)] <- NA
```

Change Comments to QA
```{r}
df_cur_syn <- df_cur_syn %>%
  mutate(
    QC_1 = case_when(grepl('broken\\.|cannot meet|did not reach|delete', Comments, ignore.case = TRUE) ~ 'BadData'),
    QC_2 = case_when(grepl('degraded', Comments, ignore.case = TRUE) ~ 'Degraded'),
    QC_3 = case_when(grepl('poorly preserved', Comments, ignore.case = TRUE) ~ 'PoorlyPreserved'),
    QC_4 = case_when(grepl('fragment\\.', Comments, ignore.case = TRUE) ~ 'Fragmented')
  ) %>%
  unite(QualityCheck, starts_with('QC'), remove = TRUE, na.rm = TRUE, sep = ' ')

df_cur_syn$QualityCheck[df_cur_syn$QualityCheck == ''] <- 'Good'
```

Merge with Main Data
```{r}
df_edi$SampleDate <- as.Date(df_edi$SampleDate, '%m/%d/%Y')
df_joined <- bind_rows(df_edi, df_cur)

# check merge didn't add errors; if so, re-evaluate
rows <- nrow(df_edi)+nrow(df_cur) == nrow(df_joined)

sprintf('When joined, the number of rows is equal to the current dataframe + the EDI dataframe: %s', rows)
```
```{r}
year <- lubridate::year(Sys.Date())-1
write_csv(df_joined, glue::glue('03_Phyto/PESP/data/EMP/EMP_Phyto_Data_2008-{year}.csv'))
```

