```{r message = FALSE}
# read in packages
library(tidyverse)
library(lubridate)
library(stats)
library(plotly)
library(broom)
```

```{r}
# df_cwq <- read_csv('01_WaterQuality/00_archive/data/spc/MRZ_data_final.csv')
# 
# df_cwq <- df_cwq %>%
#   mutate(Month = factor(month(mdy(date))),
#          Year = factor(year(mdy(date)))) %>%
#   group_by(Month, Year) %>%
#   summarize(MRZ = round(mean(spc, na.rm = TRUE)), .groups = 'drop')
```


```{r message = FALSE}
# read in data
df_dwq <- read_csv('01_WaterQuality/00_archive/data/spc/dwq_20-23.csv')

# mutate df, drop NAs (< RL values)
df_dwq <- df_dwq %>%
  subset(select = c(`Short Station Name`, Analyte, Result, `Collection Date`)) %>%
  rename(Station = `Short Station Name`, Date = `Collection Date`) %>%
  drop_na() %>%
  group_by(Station,Date,Analyte) %>%
  summarize(Result = mean(Result, na.rm = TRUE))
```

```{r}
ggplot(df_dwq, aes(Analyte, Result)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_jitter(pch = 21, color = 'black', size = 2) +
  theme_bw()
```

```{r}
df_qq <- df_dwq %>%
  rowid_to_column() %>%
  pivot_wider(id_cols = c(Station, Date), names_from = Analyte, values_from = Result) %>%
  rename(SpC = `Specific Conductance`, FieldSpC = `Field Specific Conductance`)

df_qq <- df_qq %>%
  separate(col = Date, into = c('Date', 'Time'), sep = ' ') %>%
  mutate(Month = factor(month(mdy(Date))),
         Year = factor(year(mdy(Date))),
         RPD = (SpC-FieldSpC)/SpC,
         Dif = SpC-FieldSpC,
         Mean = rowMeans(cbind(SpC, FieldSpC), na.rm = TRUE))

df_qq <- df_qq[!is.na(df_qq$RPD),]
# df_qq <- df_qq %>% right_join(df_cwq, by = c('Year', 'Month')) %>%
#   pivot_longer(cols = c(SpC, FieldSpC, MRZ), names_to = 'SpC_Type', values_to = 'Value')

qqnorm(df_qq$RPD, pch = 1, frame = FALSE)
qqnorm(df_qq$Dif, pch = 1, frame = FALSE)

rpd_plt <- ggplot(df_qq, aes(Mean, RPD, label = Date)) +
  geom_point(pch=21, size = 3, color='black', alpha  = 0.5, fill = 'gray') +
  facet_wrap(~Year, scales = 'free_y') +
  geom_smooth(method = 'lm', color = 'blue') +
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'red', size=1) +
  theme_bw()


print(rpd_plt)

ggsave('01_WaterQuality/00_archive/data/spc/rpd.png', rpd_plt, height = 4, width = 7)
```

```{r}
# norm_plt <- ggplot(df_qq, aes(Mean, Dif, fill = Year, label = Date)) +
#   geom_point(pch=21, size = 3, color='black', alpha = 0.8) +
#   theme_bw() +
#   facet_wrap(~Year, scales = 'free_y') +
#   geom_smooth(method = 'lm') +
#   geom_hline(yintercept = 0, linetype = 'dashed', color = 'red', size=1)

```


```{r}
lm_overall <- lm(Dif ~ Mean, data = df_qq)
summary(lm_overall)

df_qq <- df_qq %>%
  mutate(Pred = lm_overall$coefficients[[1]]+Mean*lm_overall$coefficients[[2]],
         Norm = Pred - Dif)

lm_sim <- lm(abs_dif ~ samp_mean, data = df_sim)
summary(lm_sim)


df_sim <- df_sim %>%
  mutate(Pred = lm_sim$coefficients[[1]]+samp_mean*lm_sim$coefficients[[2]])

df_test <- df_qq %>% subset(Mean >= 4000 & Mean <= 6000)

length(df_test$Norm[df_test$Norm > 0])/length(df_test$Norm)

length(df_test$Dif[df_test$Dif < df_test$Pred])/length(df_test$Dif)

length(df_test$Dif[df_test$Dif > df_test$Pred])/length(df_test$Dif)

length(df_sim$abs_dif[df_sim$abs_dif < df_sim$Pred])/length(df_sim$abs_dif)

```

```{r}
norm_plt <- ggplot(df_test, aes(Mean, Dif, label = Date)) +
  geom_point(pch=21, size = 3, color='black', alpha = 0.8, aes(fill = Year)) +
  theme_bw() +
  # facet_wrap(~Year, scales = 'free_y') +
  # geom_smooth(method = 'lm', color = 'black') +
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'black', size=1)

norm_plt
# print(plt)
ggsave('01_WaterQuality/00_archive/data/spc/norm_all.png', norm_plt, height = 4, width = 7)

length(df_qq$Norm[df_qq$Norm > 0])
length(df_qq$Norm)

```


```{r}
ggplotly(plt, tooltip = 'Date')
```


    def __init__(self, value_independent_sd: float, value_dependent_sd: float):
        self.value_independent_sd: float = value_independent_sd
        self.value_dependent_sd: float = value_dependent_sd

    def sample(self, value: float) -> float:
        value_independent_error = random.normalvariate(0, self.value_independent_sd)
        value_dependent_error = random.normalvariate(0, value * self.value_dependent_sd)
        return value_independent_error + value_dependent_error
        
        
        value = true_value + self.combined_error_distribution.sample(true_value)


```{r}
set.seed(42)

sd = 5
df_sim <- data.frame(samp_one = as.integer(runif(n = 2000, min = 50, max = 45000))) %>%
  mutate(
    samp_two = as.integer((rnorm(n = 2000, mean = samp_one, sd = 6)+rnorm(n = 2000, mean = samp_one, sd = 0.001*samp_one))/2),
    samp_mean = as.integer((samp_one+samp_two)/2),
    abs_dif = samp_one - samp_two,
    rel_dif = (samp_one-samp_two)/(samp_mean))

plt_sim <- ggplot(df_sim, aes(samp_mean, abs_dif)) +
  geom_point(pch = 21, size = 3, fill = 'gray', alpha = 0.6) +
  theme_bw() +
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'red', size=1)

ggsave('01_WaterQuality/00_archive/data/spc/sim_abs.png', plt_sim, height = 4, width = 7)

```

```{r}
length(df_qq$RPD[df_qq$RPD < 0])/length(df_qq$RPD)

length(df_qq$Dif[df_qq$Dif < df_qq$Pred])/length(df_qq$Dif)
length(df_qq$Dif[df_qq$Dif < df_qq$Pred])/length(df_qq$Dif)
```

