# Author: Megan Pagliaro and Perry
# Title: Finish Cleaning Corrected Data (pulled from Aquarius)
# Date: 24 June 2026
#Data downloaded from Aquarius onto local computer

# Code Starts Here
```{r}
#run packages
library(tidyverse)
library(lubridate)
library(purrr)
library(dplyr)
library(stringr)
```

# Import data
##STOP## You will need to personalize code below

```{r}
#code to import

#first get your working directory
getwd()
#code to set working directory incase it is helpful:
#setwd()

#second, move the downloaded csv file from aquarius to the folder of your working directory

#third, import the data (replacing everything in brackets (and remove the brackets themselves) with your own working directory and the file name)
df_aqui <- read.csv("[YourWorkingDirectory]/[FileName.csv]")
#Example: df_aqui <- read.csv("c:/Users/mpagliar/OneDrive - California Department of Water Resources/Documents/Github/emp-projects/AQS-observations-2025.csv")

#Run if want to do a visual check
#View(df_aqui)
```


# Filter out un-needed data

```{r}
# filter out replicate/blank data
df_year <- df_aqui %>%
  filter(QC..Type != 'BLANK' & QC..Type != 'REPLICATE' )

# filter out Churn Bucket/Filter Container data
df_year <- df_year %>%
  filter(!Observed.Property.ID %in% c('Churn Bucket ID', 'Filter Container ID'))

#separate date and time column
df_year <- df_year %>%
  separate_wider_delim(cols = Observed.DateTime, delim = " ", names = c("ObservedDate", "ObservedTime"))

```

# Formating data for EDI

```{r}
#code which formats data for EDI
df_year <- df_year %>% 
  select(
    Location.ID,
    ObservedDate,
    ObservedTime,
    EA_Sampling.Depth,
    Observed.Property.ID,
    Result.Value,
    Result.Unit,
    Lab..Detection.Condition,
    Lab..MRL,
    Lab..Analysis.Method,
    Lab..Quality.Flag,
    Field..Comment
  )

# fix detection condition
# first changing text case of "NOT_DETECTED" to "Not detected"
df_year <- df_year %>% mutate(Lab..Detection.Condition = str_to_title(Lab..Detection.Condition))
df_year$Lab..Detection.Condition <- gsub("_", " ", df_year$Lab..Detection.Condition)
# Changing all empty cells to "Detected"
df_year <- df_year %>% 
  mutate(Lab..Detection.Condition = case_match(
    Lab..Detection.Condition, "" ~ "Detected", .default = Lab..Detection.Condition)
    )

# rename columns
df_year <- df_year %>%
  rename(
    Station = Location.ID,
    Date = ObservedDate,
    Time = ObservedTime,
    Sampling_Depth = EA_Sampling.Depth,
    Analyte = Observed.Property.ID,
    Result_Value = Result.Value,
    Result_Unit = Result.Unit,
    Detection_Status = Lab..Detection.Condition,
    MRL = Lab..MRL,
    Analysis_Method = Lab..Analysis.Method,
    QC_Flag = Lab..Quality.Flag,
    Field_Comment = Field..Comment
  )
  
# date to datetype
df_year <- df_year %>%
  mutate(
    Date = as.Date(Date, format = '%Y-%m-%d')
  )

#arrange the data so it is grouped by sample run
df_year <- df_year %>%
arrange(Date, Station, Analyte) 

#Run if want to do a visual check
#View(df_year)
```

# Checks

#Check Dates

```{r}
# check missing dates
df_year %>%
  filter(is.na(Date))
```

#Check Analytes

```{r}
#define source of csv's needed for checks
source(here::here('00_GlobalFunctions/functions.R'))

#Check analytes function creation 
check_analytes <- function(df, year = 'all', return_df = FALSE) {
  
# read analyte reference list
df_analytes <- read_csv(here::here('00_GlobalFunctions/dwq_analytes.csv'), show_col_types = FALSE)
  
# read active stations
df_stations <- read_csv(here::here('00_GlobalFunctions/station_names.csv'), show_col_types = FALSE) %>%
  filter(Status == 'active')
  
valid_analytes <- unique(df_analytes$Analyte)
valid_stations <- unique(df_stations$Station_new)
  
# stations exempt from full-month presence check
exempt_month_check <- c('LSZ2-SJR', 'LSZ6-SJR', 'NZ002', 'NZ004', 'NZ325')
  
# remove analytes that should never be checked
valid_analytes <- setdiff(valid_analytes, c('Filter Container ID', 'Churn Bucket ID'))
  
# create month field
df <- df %>%
  mutate(Month = month(Date))
  
# optionally filter to one year
  if (!identical(year, 'all')) {
    df <- df %>%
      filter(str_detect(Date, paste0('^', year)))
}
  
# keep only valid stations
df_valid <- df %>%
  filter(Station %in% valid_stations)
  
  # check for station-months with no data at all
all_months <- df_valid %>%
  distinct(Month) %>%
  arrange(Month)
  
missing_station_months <- expand_grid(
  Station = setdiff(valid_stations, exempt_month_check),
  all_months) %>%
  anti_join(df_valid %>%
    distinct(Station, Month), by = c('Station', 'Month')
  ) %>%
  arrange(Month, Station)
  
if (nrow(missing_station_months) > 0) {
  message(
    'Stations with no data for entire month(s):\n',
    missing_station_months %>%
      mutate(row_txt = paste0(Station, ' - ', month.abb[Month])) %>%
      pull(row_txt) %>%
      paste(collapse = '\n')
  )
} else {
  message('All active stations have at least some data in every month.')
  }
  
# join SamplingType
df_joined <- df_valid %>%
  left_join(
    df_stations %>% select(Station_new, SamplingType),
    by = c('Station' = 'Station_new')
  )
  
# gather observed stations/dates
station_dates <- df_joined %>%
  distinct(Station, Date)
  
# create expected combos (station x analyte x date)
expected <- expand_grid(
  station_dates,
  Analyte = valid_analytes
  )
  
# RULES
# 1. shore stations do not require Secchi
shore_stations <- df_stations %>%
  filter(SamplingType == 'shore') %>%
  pull(Station_new)
  
expected <- expected %>%
  filter(!(Station %in% shore_stations & Analyte == 'Secchi Depth'))
  
# 2. non-LSZ stations do not need Lat/Long
expected <- expected %>%
  filter(!(
    !grepl('LSZ', Station, ignore.case = TRUE) &
      Analyte %in% c('Latitude', 'Longitude')
  ))
  
# 3. only D7, D16, and D19 have Dissolved Bromide
expected <- expected %>%
  filter(!(Analyte == 'Dissolved Bromide' &
    !(Station %in% c('D7', 'D16', 'D19'))))
  
# 4. C10A does not require any analyte containing '(Bottom)'
expected <- expected %>%
  filter(!(Station == 'C10A' & str_detect(Analyte, fixed('(Bottom)'))))
  
# observed combos
observed <- df_joined %>%
    distinct(Station, Analyte, Date)
  
# find missing combinations
missing_combos <- anti_join(
  expected,
  observed,
  by = c('Station', 'Analyte', 'Date')
  ) %>%
  arrange(Date, Station, Analyte)
  
# 5. if detailed weather analytes are missing but Weather Observations exists,
# do not report Rain, Sky Conditions, or Wave Scale as missing
weather_observed <- observed %>%
  filter(Analyte == 'Weather Observations') %>%
  distinct(Station, Date) %>%
  mutate(has_weather_observations = TRUE)
  
missing_combos <- missing_combos %>%
  left_join(weather_observed, by = c('Station', 'Date')) %>%
  filter(!(
    Analyte %in% c('Rain', 'Sky Conditions', 'Wave Scale') &
      has_weather_observations %in% TRUE
  )) %>%
  select(-has_weather_observations) %>%
  arrange(Date, Station, Analyte)
  
# format message
fmt_row <- function(x) {
  parts <- x[!is.na(x) & x != '']
  paste(parts, collapse = ' - ')
}
  
if (nrow(missing_combos) > 0) {
  message(
    'Missing ', nrow(missing_combos),
    ' station/analyte combinations:\n',
    paste0(apply(missing_combos, 1, fmt_row), collapse = '\n')
  )
} else {
  message('All expected station/analyte combinations are represented.')
}
  
if (return_df) {
  return(missing_combos)
}}

#Check analytes on dataset itself by running function created above
check_analytes(df_year)

```

#Check Duplicates

```{r}
check_duplicates <- function(df, year = 'all', return_df = FALSE) {
  
  df_dups <- df %>%
    mutate(
      Year = year(Date),
      Month = month(Date)
    ) %>%
    {
      if (!identical(year, 'all')) {
        filter(., Year == as.character(year))
      } else {
        
      }
    } %>%
    group_by(Analyte, Station, Year, Month) %>%
    filter(n() > 1) %>%
    arrange(Analyte, Station, Year, Month, Date) %>%
    ungroup()
  
  dup_summary <- df_dups %>%
    distinct(Analyte, Station, Year, Month)
  
  if (nrow(df_dups) > 0) {
    message(
      'Found duplicate combinations: ',
      nrow(dup_summary), '\n',
      paste0(
        dup_summary %>%
          mutate(txt = paste(Analyte, Station, paste0(Year, '-', Month), sep = ' - ')) %>%
          pull(txt),
        collapse = '\n'
      )
    )
  } else {
    message('No duplicate combinations found.')
  }
  
  if (return_df) {
    return(df_dups)
  }
}




#Check if duplicates on data itself by running function created above
### NOTE: make sure the year is correct for the data you are formatting ###
check <- check_duplicates(df_year, year = '2025', return_df = TRUE) 

```


# Write file and convert to CSV
##STOP## You need to personalize code below

```{r}

#first convert all empty cells into NA
df_year[df_year == ""] <- NA

#Second, run the write.csv code replacing the brackets with your working directory and a new name for the file
write_csv(df_year, "[YourWorkingDirectory]/[NewNameForFile.csv")
#Example: write_csv(df_year,"c:/Users/mpagliar/OneDrive - California Department of Water Resources/Documents/Github/emp-projects/edi_2025Data_CreatedJul072426.csv")
```
