---
title: "Sonde Uptime for Table 5 Stations in D-1641"
author: "Ted Flynn"
date: "2023-06-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(lubridate)))
suppressWarnings(suppressMessages(library(janitor)))
suppressWarnings(suppressMessages(library(here)))
suppressWarnings(suppressMessages(library(cder)))

# Set output directory 
output <- here("01_WaterQuality","D1641-table5-sonde-uptime","output")

# Set visual theme in ggplot
theme_set(theme_bw())

# Suppress summarise info
options(tidyverse.quiet = TRUE)
options(dplyr.summarise.inform = FALSE)

# Create date variables --------------------------------------------------------
start <- as.Date("2020-01-01")
end <- as.Date("2024-05-08")

```

## Import sonde data from CDEC
Pull in data from six D-1641 continuous water quality stations that are maintained by USBR: Sacramento River @ Collinsville (CLL), San Joaquin River @ San Andreas Landing (SAL), Contra Costa Canal @ Pumping Plant #1 (CNT), Old River near Middle River (UNI), Sacramento River @ Port Chicago (PCT), Delta-Mendota Canal at Tracy Pumping Plant (DMC). 

```{r import}
# Pull USBR station data from CDEC
USBR_stations <- c("CLL","SAL","CNT","UNI","PCT","DMC","EMM")

df_WQ_all <- cdec_query(USBR_stations, sensors = c(100,25), 
                    start.date = start, 
                    end.date = end)

# Remove all data from years apart from 2023
df_WQ <- df_WQ_all %>% 
  mutate(Year = year(df_WQ_all$DateTime)) %>% 
  filter(Year == 2023)

# Add more detailed description of each station
df_WQ <- df_WQ %>% 
  mutate(Description = case_when(StationID == "CLL" ~ "C2 - Sac River @ Collinsville",
                                 StationID == "SAL" ~ "C4 - SJ River @ San Andreas Landing",
                                 StationID == "CNT" ~ "C5 - Contra Costa Canal @ Pumping Plant 1",
                                 StationID == "UNI" ~ "C8 - Old River near Middle River",
                                 StationID == "PCT" ~ "C14 - Sac River @ Port Chicago",
                                 StationID == "DMC" ~ "DMC1 - Delta-Mendota Canal @ Tracy Pumping Plant",
                                 StationID == "EMM" ~ "EMM - Sac River @ Emmaton"
                                 ))

```

## Clean Up Data
Create graphs that differentiate between different data types (Daily, Event, and Hourly) for SpC. 
```{r clean}
# Remove any duplicate rows
df_WQ <- df_WQ %>% distinct() # no rows removed
  
# Check CDEC column values
unique(df_WQ$DataFlag) # no values, not used for these stations

#quick check to see if there are the same number of samples per station per duration
table(df_WQ$SensorType, df_WQ$StationID, df_WQ$Duration)

# Filter out unneeded columns
df_WQ <- df_WQ %>% 
  select(!DataFlag) %>% # just a blank column
  select(!SensorNumber) %>% # CDEC codes for each analyte
  select(!SensorUnits) # we know what the units are

# Add additional date units for grouping and plotting
df_WQ <- df_WQ %>%
  mutate(Year = year(df_WQ$DateTime)) %>%
  mutate(Month = month(df_WQ$DateTime, label = TRUE)) %>%
  mutate(Date = date(df_WQ$DateTime))

# Rename output from CDEC for clarify
df_WQ <- df_WQ %>% 
  mutate(SensorType = str_replace(SensorType, "EL COND", "SpC")) %>% 
  mutate(SensorType = str_replace(SensorType, "TEMP W", "Water Temp"))

# Interpret content in the "Duration" column
unique(df_WQ$Duration) # H = hourly, E = event, D = daily

# Replace duration codes with more verbose descriptions
durations <- c("D" = "Daily","E" = "Event","H" = "Hourly")

df_WQ <- df_WQ %>% 
  mutate(Duration = str_replace_all(Duration, durations))

```

## Create a Blank Data Frame for the Data
```{r blank}
# Create a blank data frame with all combinations of stations and months for 2023
months <- seq(ymd("2024-01-01"), by = "month", length.out = 12)

df_intervals <- crossing(
  MonthNum = month(months, abbr = TRUE),
  StationID = USBR_stations,
  SensorType = c("SpC", "Water Temp")
  )

df_intervals <- df_intervals %>% 
  mutate(Daily = days_in_month(MonthNum)) %>% 
  mutate(Month = month(MonthNum, abbr = TRUE, label = TRUE)) %>% 
  mutate(Hourly = Daily * 24) %>% 
  mutate(Event = Daily * 96) %>% 
  select(-MonthNum) %>% 
  relocate(Month, .after = StationID)

df_intervals <- pivot_longer(df_intervals, 
                             cols = Daily:Event, 
                             values_to = "TotalIntervals",
                             names_to = "Duration")

# Remove "Daily" and "Event" categories for Water Temp data (only ever Hourly)
df_intervals <- df_intervals %>% 
  filter(!(SensorType == "Water Temp" & Duration %in% c("Daily","Event")))


```

## Summarize WQ Data by Month
```{r summarize data}

# Create summary table for how many individual measurements exist for each station
df_WQ_m <- df_WQ %>% 
  filter(!is.na(Value)) %>% # Remove entries with an NA for the measured parameter
  filter(Value > 0) %>% # Remove entries with a 0 or negative number for the measured parameter
  select(-Year) %>% 
  group_by(StationID,Description,Month,SensorType,Duration) %>% 
  summarize(Total_Measurements = sum(n())) %>% 
  ungroup()

# Merge with df with total intervals per month
df_WQ_sum <- left_join(df_intervals, df_WQ_m, join_by(Month,StationID,SensorType,Duration))

# Calculate Percent Data Recovery
df_WQ_sum <- df_WQ_sum %>% 
  mutate(Present = (Total_Measurements / TotalIntervals) * 100) %>% 
  mutate(Missing = 100 - Present) %>% 
  select(-c("TotalIntervals","Total_Measurements"))

# Pivot to long for plotting
df_WQ_sum <- pivot_longer(df_WQ_sum,
                          cols = Present:Missing,
                          values_to = "PercentData",
                          names_to = "PresenceAbsence")

  
```


## Plot Overall Status
```{r plotting}

# Use the black-and-white theme
theme_set(theme_bw())

# Set the theme
theme_update(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white", colour = NA),
    panel.background = element_rect(fill = "white", colour = NA),
    panel.border = element_rect(fill = NA, colour = "black"),
    strip.background = element_rect(fill = "gray", colour = "black"),
    legend.position = "bottom",
    legend.key = element_rect(fill = "white", colour = NA)
    )

# Plot sonde uptime for specific conductancy
p_SpC <- ggplot(df_WQ_sum, aes(x = Month,
                                  y = PercentData,
                                  fill = PresenceAbsence)) +
  geom_col(data = subset(df_WQ_sum, SensorType == "SpC" & Duration == "Event"),
           width = 0.7) +
  scale_y_continuous(breaks = seq(from = 0, to = 100, by = 50)) +
  labs(x = NULL,
       y = "Data Availability - SpC (% of Total)",
       fill = "Data Status on CDEC",
       title = "Data Recovery - SpC - 2023") +
  scale_fill_manual(values = c("red3","darkcyan"))

p_SpC +
    facet_wrap(Description ~ ., ncol = 1)

ggsave(path = output,
       filename = "USBR-sonde-status-2023-SpC.png",
       device = "png",
       scale=1.0,
       units="in",
       height=6,
       width=5,
       dpi="print")

# Plot sonde uptime for water temperature
p_Temp <- ggplot(df_WQ_sum, aes(x = Month,
                                  y = PercentData,
                                  fill = PresenceAbsence)) +
  geom_col(data = subset(df_WQ_sum, SensorType == "Water Temp" & Duration == "Hourly"),
           width = 0.7) +
  scale_y_continuous(breaks = seq(from = 0, to = 100, by = 50)) +
  labs(x = NULL,
       y = "Data Availability - Water T (% of Total)",
       fill = "Data Status on CDEC",
       title = "Data Recovery - Water Temp - 2023") +
  scale_fill_manual(values = c("red3","darkcyan")) # red-green is the worst for color blind people!

p_Temp +
    facet_wrap(Description ~ ., ncol = 1)

ggsave(path = output,
       filename = "USBR-sonde-status-2023-Temp.png",
       device = "png",
       scale=1.0,
       units="in",
       height=6,
       width=5,
       dpi="print")


```
```{r}
#quick plot of EC and temp over time
ggplot(filter(df_WQ, Duration == "Hourly", !is.na(Value), Value >0), aes(x = Date, y = Value, color = SensorType))+
  facet_wrap(~StationID, scales = "free_y")+ geom_point()

#what's goin gon with those super high values?
test = filter(df_WQ, Value >50000)
```
```{r}
#check the event data nad make sure no other weird shit happened.
ggplot(filter(df_WQ, Duration == "Event", !is.na(Value), Value >0), aes(x = Date, y = Value, color = SensorType))+
  facet_wrap(~StationID, scales = "free_y")+ geom_point()
#some of those points look odd
test2 = filter(df_WQ, Value <10 & SensorType=="SpC")
```


```{r}
#let's flag them
df_WQ = mutate(df_WQ, Flag = case_when(Value <10 & SensorType=="SpC" ~ "Suspicious",
                                       Value >100000 & SensorType=="SpC" ~ "Suspicious",
                                       Value <2 & SensorType=="Water Temp" ~ "Suspicious",
                                       TRUE ~ "Probably OK"))

ggplot(filter(df_WQ, Duration == "Event", !is.na(Value), Value >0), aes(x = Date, y = Value, color = Flag))+
  facet_wrap(~StationID, scales = "free_y")+ geom_point()


```

