---
title: "Sonde Uptime for Table 5 Stations in D-1641"
author: "Ted Flynn"
date: "2023-06-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(lubridate)))
suppressWarnings(suppressMessages(library(janitor)))
suppressWarnings(suppressMessages(library(here)))
suppressWarnings(suppressMessages(library(cder)))

# Set output directory 
plots <- here("01_WaterQuality","D1641-table5-sonde-uptime","output")

# Set visual theme in ggplot
theme_set(theme_bw())

# Suppress summarise info
options(tidyverse.quiet = TRUE)
options(dplyr.summarise.inform = FALSE)

# Create date variables --------------------------------------------------------
start <- as.Date("2020-01-01")
end <- as.Date("2024-12-31")

```

## Import sonde data from CDEC
Pull in data from seven D-1641 continuous water quality stations maintained by USBR: Sacramento River @ Collinsville (CLL), San Joaquin River @ San Andreas Landing (SAL), Contra Costa Canal @ Pumping Plant #1 (CNT), Old River near Middle River (UNI), Sacramento River @ Port Chicago (PCT), Delta-Mendota Canal at Tracy Pumping Plant (DMC), and Emmaton (EMM).

```{r import, echo=FALSE}
# Pull USBR station data from CDEC
stations <- c("CLL","SAL","CNT","UNI","PCT","DMC","EMM")
 
# df_WQ_all <- cdec_query(stations, sensors = c(100,25),
#                         start.date = start,
#                         end.date = end)
# 
# # Add more detailed description of each station
# df_WQ_all <- df_WQ_all %>% 
#    mutate(Description = case_when(StationID == "CLL" ~ "C2 - Sac River @ Collinsville",
#                                   StationID == "SAL" ~ "C4 - SJ River @ San Andreas Landing",
#                                   StationID == "CNT" ~ "C5 - Contra Costa Canal @ Pumping Plant 1",
#                                   StationID == "UNI" ~ "C8 - Old River near Middle River",
#                                   StationID == "PCT" ~ "C14 - Sac River @ Port Chicago",
#                                   StationID == "DMC" ~ "DMC1 - Delta-Mendota Canal @ Tracy Pumping Plant",
#                                   StationID == "EMM" ~ "EMM - Sac River @ Emmaton"
#                                   ))
# 
# # Takes a long time to download from CDEC so save as RData file
# save(df_WQ_all, file = here("01_WaterQuality","D1641-table5-sonde-uptime","df_WQ_all.RData"))

load(here("01_WaterQuality","D1641-table5-sonde-uptime","df_WQ_all.RData"))

```

## Clean Up Data
Create graphs that differentiate between different data types (Daily, Event, and Hourly) for SpC. 
```{r clean}
# Create new df for cleaning and remove duplicates
df_WQ <- df_WQ_all %>% distinct() # no rows removed

# Check CDEC column values
unique(df_WQ$DataFlag) # no values, not used for these stations

#quick check to see if there are the same number of samples per station per duration
table(df_WQ$SensorType, df_WQ$StationID, df_WQ$Duration)

# Filter out unneeded columns
df_WQ <- df_WQ %>% 
  select(!DataFlag) %>% # just a blank column
  select(!SensorNumber) %>% # CDEC codes for each analyte
  select(!SensorUnits) # we know what the units are

# Add additional date units for grouping and plotting
df_WQ <- df_WQ %>%
  mutate(Year = year(df_WQ$DateTime)) %>%
  mutate(Month = month(df_WQ$DateTime, label = TRUE)) %>%
  mutate(Date = date(df_WQ$DateTime))

# Rename output from CDEC for clarify
df_WQ <- df_WQ %>% 
  mutate(SensorType = str_replace(SensorType, "EL COND", "SpC")) %>% 
  mutate(SensorType = str_replace(SensorType, "TEMP W", "Water Temp"))

# Interpret content in the "Duration" column
unique(df_WQ$Duration) # H = hourly, E = event, D = daily

# Replace duration codes with more verbose descriptions
durations <- c("D" = "Daily","E" = "Event","H" = "Hourly")

df_WQ <- df_WQ %>% 
  mutate(Duration = str_replace_all(Duration, durations))

# Create summary table for how many individual measurements exist for each station
df_WQ_sum <- df_WQ %>% 
  filter(Duration == "Hourly") %>% 
  filter(!is.na(Value)) %>% # Remove entries with an NA for the measured parameter
  filter(Value > 0) %>% # Remove entries with a 0 or negative number for the measured parameter
  group_by(StationID,Description,Year,Month,SensorType,Duration) %>% 
  summarize(Total_Measurements = sum(n())) %>% 
  ungroup()

# Create data frame with all date time combinations from 1-1-2020 to 12-31-2024

hourly_date_times <- tibble(
  DateTime = seq.POSIXt(
    from = as.POSIXct("2020-01-01 00:00:00"),
    to = as.POSIXct("2024-12-31 23:00:00"),
    by = "hour"
  )
)

df_test <- crossing(StationID = stations, hourly_date_times)

# Filter CDEC data for hourly EC data only
df_WQ_EC <- df_WQ %>% 
  filter(SensorType == "SpC") %>% 
  filter(Duration == "Hourly") %>% 
  select(-Duration,-ObsDate)

df_test <- left_join(df_test,df_WQ_EC)

df_test <- df_test %>% 
  mutate(DataStatus = case_when(is.na(Value) ~ "Missing",
                                Value == 0 ~ "Zero",
                                TRUE ~ "Present"))

# Plot rating data faceted by probe
colors <- c("darkred","darkgreen","yellow3")

p1 <- ggplot(data = df_test, aes(x = DateTime,
                                 y = StationID,
                                 fill = DataStatus)) +
  geom_tile(height = 0.5) +
  scale_fill_manual(values = colors) 

p1 +
  labs(x = "Daily Rating of Data Collected",
       y = "Station ID",
       fill = "Sonde Rating",
       title = "EXO2 Probe Ratings - 2023")

ggsave(path = output,
       filename = "CEMP_sonde_scores_test.png", 
       device = "png",
       scale=1.0, 
       units="in",
       height=6.5,
       width=9, 
       dpi="print")
```

```

## Basic Plots for Data Review

```{r basic plots}

# Create history plots for SpC at each stations
stations <- unique(df_WQ$StationID) 
stations <- sort(stations, decreasing = F, na.last = T)

for (x in stations) {
  df_temp <- df_WQ %>%
    filter(StationID == x) %>% 
    filter(SensorType == "SpC")
  
  p <- ggplot(data = df_temp, aes(x = DateTime,
                                  y = Value)) +
    geom_point(size = 1) + 
    scale_color_brewer(palette = "Set2") + 
    facet_wrap(Duration ~ ., ncol = 1)
  
  print(p)

  ggsave(path = plots,
         filename = paste0(x,"-metadata-timeline-CDEC.png"),
         device = "png",
         scale=1.0,
         units="in",
         height=7,
         width=6.5,
         dpi="print")
  
}

```





## Create a Blank Data Frame for the 2023 Data
```{r blank}
# Create a blank data frame with all combinations of stations and months for 2023
months <- seq(ymd("2024-01-01"), by = "month", length.out = 12)

days <- as_tibble(seq(ymd("2024-01-01"), by = "day", length.out = 366))

df_intervals <- crossing(
  Month = month(months, abbr = TRUE, label = TRUE),
  Year = c("2020","2021","2022","2023","2024"),
  StationID = stations,
  SensorType = c("SpC", "Water Temp")
  )

df_intervals <- df_intervals %>% 
  mutate(Daily = days_in_month(MonthNum)) %>% 
  mutate(Month = month(MonthNum, abbr = TRUE, label = TRUE)) %>% 
  mutate(Hourly = Daily * 24) %>% 
  mutate(Event = Daily * 96) %>% 
  select(-MonthNum) %>% 
  relocate(Month, .after = StationID)

df_intervals <- pivot_longer(df_intervals, 
                             cols = Daily:Event, 
                             values_to = "TotalIntervals",
                             names_to = "Duration")

# Remove "Daily" and "Event" categories for Water Temp data (only ever Hourly)
df_intervals <- df_intervals %>% 
  filter(!(SensorType == "Water Temp" & Duration %in% c("Daily","Event")))


```

## Summarize WQ Data by Month
```{r summarize data}
# 


# Create summary table for how many individual measurements exist for each station
df_WQ_m <- df_WQ %>% 
  filter(!is.na(Value)) %>% # Remove entries with an NA for the measured parameter
  filter(Value > 0) %>% # Remove entries with a 0 or negative number for the measured parameter
  select(-Year) %>% 
  group_by(StationID,Description,Month,SensorType,Duration) %>% 
  summarize(Total_Measurements = sum(n())) %>% 
  ungroup()

# Merge with df with total intervals per month
df_WQ_sum <- left_join(df_intervals, df_WQ_m, join_by(Month,StationID,SensorType,Duration))

# Calculate Percent Data Recovery
df_WQ_sum <- df_WQ_sum %>% 
  mutate(Present = (Total_Measurements / TotalIntervals) * 100) %>% 
  mutate(Missing = 100 - Present) %>% 
  select(-c("TotalIntervals","Total_Measurements"))

# Pivot to long for plotting
df_WQ_sum <- pivot_longer(df_WQ_sum,
                          cols = Present:Missing,
                          values_to = "PercentData",
                          names_to = "PresenceAbsence")

  
```


## Plot Overall Status
```{r plotting}

# Use the black-and-white theme
theme_set(theme_bw())

# Set the theme
theme_update(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white", colour = NA),
    panel.background = element_rect(fill = "white", colour = NA),
    panel.border = element_rect(fill = NA, colour = "black"),
    strip.background = element_rect(fill = "gray", colour = "black"),
    legend.position = "bottom",
    legend.key = element_rect(fill = "white", colour = NA)
    )

# Plot sonde uptime for specific conductancy
p_SpC <- ggplot(df_WQ_sum, aes(x = Month,
                                  y = PercentData,
                                  fill = PresenceAbsence)) +
  geom_col(data = subset(df_WQ_sum, SensorType == "SpC" & Duration == "Event"),
           width = 0.7) +
  scale_y_continuous(breaks = seq(from = 0, to = 100, by = 50)) +
  labs(x = NULL,
       y = "Data Availability - SpC (% of Total)",
       fill = "Data Status on CDEC",
       title = "Data Recovery - SpC - 2023") +
  scale_fill_manual(values = c("red3","darkcyan"))

p_SpC +
    facet_wrap(Description ~ ., ncol = 1)

ggsave(path = output,
       filename = "USBR-sonde-status-2023-SpC.png",
       device = "png",
       scale=1.0,
       units="in",
       height=6,
       width=5,
       dpi="print")

# Plot sonde uptime for water temperature
p_Temp <- ggplot(df_WQ_sum, aes(x = Month,
                                  y = PercentData,
                                  fill = PresenceAbsence)) +
  geom_col(data = subset(df_WQ_sum, SensorType == "Water Temp" & Duration == "Hourly"),
           width = 0.7) +
  scale_y_continuous(breaks = seq(from = 0, to = 100, by = 50)) +
  labs(x = NULL,
       y = "Data Availability - Water T (% of Total)",
       fill = "Data Status on CDEC",
       title = "Data Recovery - Water Temp - 2023") +
  scale_fill_manual(values = c("red3","darkcyan")) # red-green is the worst for color blind people!

p_Temp +
    facet_wrap(Description ~ ., ncol = 1)

ggsave(path = output,
       filename = "USBR-sonde-status-2023-Temp.png",
       device = "png",
       scale=1.0,
       units="in",
       height=6,
       width=5,
       dpi="print")


```
```{r}
#quick plot of EC and temp over time
ggplot(filter(df_WQ, Duration == "Hourly", !is.na(Value), Value >0), aes(x = Date, y = Value, color = SensorType))+
  facet_wrap(~StationID, scales = "free_y")+ geom_point()

#what's goin gon with those super high values?
test = filter(df_WQ, Value >50000)
```
```{r}
#check the event data nad make sure no other weird shit happened.
ggplot(filter(df_WQ, Duration == "Event", !is.na(Value), Value >0), aes(x = Date, y = Value, color = SensorType))+
  facet_wrap(~StationID, scales = "free_y")+ geom_point()
#some of those points look odd
test2 = filter(df_WQ, Value <10 & SensorType=="SpC")
```


```{r}
#let's flag them
df_WQ = mutate(df_WQ, Flag = case_when(Value <10 & SensorType=="SpC" ~ "Suspicious",
                                       Value >100000 & SensorType=="SpC" ~ "Suspicious",
                                       Value <2 & SensorType=="Water Temp" ~ "Suspicious",
                                       TRUE ~ "Probably OK"))

ggplot(filter(df_WQ, Duration == "Event", !is.na(Value), Value >0), aes(x = Date, y = Value, color = Flag))+
  facet_wrap(~StationID, scales = "free_y")+ geom_point()


```

