# Variables and Types

R is completely object oriented, and not "statically typed". You do not need to declare variables before using them, or declare their type. Every variable in R is an object.

This tutorial will go over a few basic types of variables.


## Numbers

R supports two types of numbers - integers and doubles (floating point numbers). (It also supports complex numbers, which will not be explained in this tutorial).

To define an integer, use the following syntax:


```{r}
myint <- 7
print(myint)

Hello <- 8
print(Hello)

floating_point <- 3.14159
print(floating_point)
```

In addition, we have characters! Can use single or double quotes.

```{r}
a_character <- 'a character'
print(a_character)

z <- "also a character"
print(z)

# numbers can also be characters
num_char <- '5'
num <- 5

typeof(num_char)
typeof(num)

num_char + 5 # if you run into this error, check if everything is actually a number
```
Other important types:

```{r}
# vectors!
a_vector <- c(5,3,54,6)
not_a_vector <- c(5,'abc',3,'eafea') # ie. a list

# vectors apply mathmatical functions element-wise by default
multi_vector <- a_vector * 2
print(multi_vector)

# boolean!
typeof(1) == typeof('a character') # false, one is 'double' while the other is 'character'
```

What is a dataframe?

Dataframes are composed of vectors.
```{r}
# these individual vectors will make up a dataframe
employee <- c('John Doe','Peter Gynn','Jolie Hope')
salary <- c(21000, 23400, 26800)
startdate <- as.Date(c('2010-11-1','2008-3-25','2007-3-14'))

# here is our dataframe!
employ_data <- data.frame(employee, salary, startdate) #

print(employ_data)

# to look at an invidual column (vector) in a dataframe, use the $ sign
print(employ_data$salary)
```

Can manipulate elements in a vector.

```{r}
print(salary)

# index a vector using []; what's in the [] is called the condition
salary[3]

# index multiple elements in a row (a 'slice') using [:]
salary[2:3]

# can subset by true/false conditions
salary[salary < 22000]
```

Can reassign elements in a vector (or column)
```{r}
salary[2] <- 1000
salary

# redefine salary v quick
salary <- c(21000, 23400, 26800)

# reassign multiple things
salary[2:3] <- 'reassigning :D'

salary
```


