Title: Nutrient Preservation Comparison Study
Contact: sarah.perry@water.ca.gov
Date: 01/2023

Summary: 2-way paired t-test to determine if there's a significant difference between acidified and frozen samples from Bryte

# Read in data
```{r message = FALSE}
# read in packages
library(tidyverse)
library(stats)
```

```{r message = FALSE}
# read in data
df_wq <- read_csv('01_WaterQuality/nutrient-comparison-study/data/nutrient_data.csv')

vec_analytes <- unique(df_wq$Analyte)

# mutate df, drop NAs (< RL values)
df_wq_pivot <- df_wq %>%
   mutate(Difference = Frozen-Acidified,
          StationType = case_when(
             StationName %in% c('D22','P8','D19','D16 - Twitchell') ~ 'freshwater',
             StationName == 'D7' ~ 'brackish',
             StationName == 'D41' ~ 'saline'
          )) %>%
   drop_na()

df_wq_pivot <- rowid_to_column(df_wq_pivot, 'sample')

df_wq <- df_wq_pivot %>%
   pivot_longer(cols = c(Frozen, Acidified),
               names_to = 'PresType',
               values_to = 'Result')
```

# Descriptive analysis

Summary by preservation type
```{r}
by(data = df_wq, 
   INDICES = c(df_wq[,'Analyte'],df_wq[,'PresType']), 
   FUN = summary)
```

# Check boxplots (individual and difference)

First, we look at the individual preservation types to see if there are any outliers
```{r}
ggplot(df_wq, aes(PresType, Result)) +
  geom_boxplot()+
   facet_wrap(~Analyte)
```

As we don't see any, we can move on.

The boxplot of the differences gives us a qualitative take on the data.
```{r}
ggplot(df_wq, aes(,Difference)) +
  geom_boxplot()+
   facet_wrap(~Analyte)
```
TKN seems to have a higher variance, but a mean close to 0 still. All the rest look okay.

# Check normality

Residuals aren't normal; however, we are assuming we have enough data for the CLT to be valid, and therefore are unconcerned. To be safe, we will also run the model using log-transformed values (which more closely approximate normality):

# linear regression
```{r}
for (analyte in vec_analytes){
   df_test <- df_wq %>% subset(Analyte == analyte)
   
   df_test_pivot <- df_test %>%
      pivot_wider(names_from = 'PresType',
                  values_from = 'Result')
   
   plt <- ggplot(df_test_pivot, aes(Acidified, Frozen, color = StationType)) +
      geom_point(size=2) +
      geom_smooth(method='lm') +
      theme_bw() +
      facet_wrap(~StationName, scales = 'free') +
      ggtitle(analyte)
   
   print(plt)

   ggsave(paste0('01_WaterQuality/nutrient-comparison-study/plots/lm_',analyte,'.jpg'),
          plot = plt,
          height = 3,
          width = 4.5)
   }
```
```{r}
for (analyte in vec_analytes){
   df_test <- df_wq %>% subset(Analyte == analyte)
   test <- bartlett.test(df_test$Result ~ df_test$PresType)
   print(test)
}
```

# f-test
```{r}
var.test(df_wq_pivot$Frozen, df_wq_pivot$Acidified, alternative = 'two.sided')
```

# t-test

First, we run the model with no transformations.
```{r}
for (analyte in vec_analytes){
   df_test <- df_wq %>% subset(Analyte == analyte) 
   
   ttest <- t.test(formula = df_test$Result ~ df_test$PresType,
                   alternative = 'two.sided',
                   mu = 0, 
                   paired = TRUE,   
                   var.equal = FALSE,
                   conf.level = 0.95)
   
   print(paste0('~~~',analyte,'~~~'))
   print(ttest)
   print('~~~~~~~~')

   # (for fun, also do with linear mixed effect model; *assumes equal variance*, which we won't, so results don't matter)
   df_test <- df_wq %>% subset(df_wq$Analyte == analyte)
   model <- summary(lmerTest::lmer(Result ~ PresType + (1|sample), df_test), ddf = 'Kenward-Roger')
   print(model)
}
```
   }
```
p > 0.05 for all analytes.

Accept null hypothesis of no difference -- the two are the same -- for all analytes.


To check that our violation of the non-normality assumption does not affect the 