```{r results='asis', message=FALSE, echo=FALSE}
library(kableExtra)
library(tibble)

df <- tribble(
  ~Design, ~Sample_Size, ~Recommended_Test, ~R_Functions, ~Why,

  # LM
  "Balanced", "Any", "F-test (classical ANOVA)", "anova(lm(...))",
  "Sums of squares partition cleanly due to orthogonality",

  "Unbalanced", "Any", "LRT (nested models)", "anova(lm1, lm2)",
  "More robust than contrast-dependent F-tests; handles imbalance",

  # LMM
  "Balanced", "Large", "LRT (ML fits) or Kenward–Roger", "anova(...), KRmodcomp()",
  "Both valid under large n; KR adjusts df, LRT compares nested models",

  "Balanced", "Small", "Kenward–Roger F-test", "lmerTest::anova(), KRmodcomp()",
  "Adjusts for small-sample bias in denominator df",

  "Unbalanced", "Large", "LRT (ML fits)", "anova(..., REML = FALSE)",
  "Robust if model is large and well-estimated",

  "Unbalanced", "Small", "Kenward–Roger or Bootstrap LRT", "KRmodcomp(), PBmodcomp()",
  "LRT may be anti-conservative; KR/bootstrap fix small-sample bias",

  # GLM
  "Balanced", "Large", "Wald or LRT", "anova(glm(...), test = \"Chisq\")",
  "Wald test matches LRT under regularity; both valid",

  "Unbalanced", "Large", "LRT (nested models)", "anova(glm1, glm2, test = \"Chisq\")",
  "More reliable than Wald under imbalance",

  "Any", "Small", "Avoid Wald; cautious LRT", "anova(glm1, glm2)",
  "Wald test has inflated Type I error in small samples",

  # GLMM
  "Balanced", "Large", "LRT (ML fits)", "anova(glmm1, glmm2)",
  "Valid for nested models; REML not allowed for fixed effect testing",

  "Balanced", "Small", "LRT (caution)", "anova(...)",
  "Works better than Wald, but not perfect at small n",

  "Unbalanced", "Large", "LRT or AIC", "anova(...), AIC()",
  "Handles complex structures; bootstrapping safer",

  "Unbalanced", "Small", "Parametric bootstrap LRT", "PBmodcomp(), afex::mixed()",
  "Wald and LRT unreliable near boundary; bootstrap gives correct null distribution"
)

kbl(df, format = "html", escape = FALSE,
    col.names = c("Design", "Sample Size", "Recommended Test", "R Function", "Why"),
    align = "l") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  row_spec(0, bold = TRUE) %>%
  pack_rows("LM", 1, 2, background = "lightgray") %>%
  pack_rows("LMM", 3, 6, background = "white") %>%
  pack_rows("GLM", 7, 9, background = "lightgray") %>%
  pack_rows("GLMM", 10, 13, background = "white")
```

# Additional Notes

## Type III ANOVA vs. LRT

Type III ANOVA
Type III ANOVA tests the marginal effect of each term after adjusting for all other terms in the model, including interactions.

It is commonly used in software like SAS and in car::Anova(..., type = 3) in R.

It is sometimes recommended for unbalanced designs because it ensures each term is tested in the full model context, regardless of order.

However:

Results depend critically on contrast coding:

For Type III ANOVA to work sensibly in R, you must set sum-to-zero contrasts, e.g.:

r
Copy
Edit
options(contrasts = c("contr.sum", "contr.poly"))
If you use treatment (default) contrasts, Type III tests lose interpretability, especially for interactions.

Even with proper contrasts, Type III tests can produce F-statistics that don't correspond to clear model comparisons, especially when design is unbalanced and effects are not orthogonal.

In mixed models, Type III ANOVA assumes denominator degrees of freedom in complex, often heuristic ways — e.g., via Satterthwaite or Kenward–Roger approximations.

Likelihood Ratio Tests (LRTs)
LRTs compare nested models: one full model, and one reduced model without the effect of interest.

They ask: "Does removing this term significantly worsen model fit?"

In R, this is done via:

Advantages:

Contrast coding is irrelevant — the test compares overall model fit.

More robust to imbalance, especially when models are nested.

Interpretable: it answers the exact question of whether the effect improves the model.

Preferred for testing fixed effects in unbalanced or complex models, or when you want reproducibility across software environments.

Limitations:

Must use ML (REML = FALSE) — REML is only valid when fixed effects are identical.

Only applicable for nested model comparisons.

The LRT is asymptotic — small samples can lead to anti-conservative p-values (especially for random effects or boundary parameters).

So Which to Prefer?
Scenario	Preferred Test	Why
Balanced design	Type III ANOVA (with correct contrasts) or LRT	Either is fine; orthogonality makes results align
Unbalanced design	LRT	Type III depends on contrasts; LRT compares real models
Mixed models (LMM/GLMM)	LRT or Kenward–Roger, not Type III	Type III less reliable for mixed models, LRT is clearer
Complex interactions	LRT	Avoid Type III’s ambiguous marginal tests

Final Recommendation
If you’ve set contrasts correctly (e.g., contr.sum) and the design is not too messy, Type III ANOVA can be used cautiously.
But for general-purpose, model-based inference, especially with unbalanced designs or mixed models, Likelihood Ratio Tests are more robust, easier to interpret, and not contrast-dependent.

Note
If you want to use Type III ANOVA, make sure to set cont.sum correctly.

## REML

Use REML = FALSE (i.e., maximum likelihood) when comparing nested models with different fixed effects, such as in likelihood ratio tests (LRTs).

Use REML = TRUE (i.e., restricted maximum likelihood) when testing fixed effects within a single model, especially when using Kenward–Roger adjusted F-tests.

REML is only valid for comparing models with the same fixed-effects structure — otherwise, the likelihoods are not comparable.

**Summary: REML = TRUE is only for Kenward-Roger adjusted F-tests**

## Kenward-Roger vs Satterthwaite

Both Kenward–Roger and Satterthwaite approximations are used to estimate denominator degrees of freedom for F-tests or t-tests in linear mixed models, particularly when sample sizes are small or unbalanced.

Satterthwaite is:

Simpler and faster

Used in lmerTest::anova() by default for t-tests

Based on approximating the distribution of a ratio of variances

Often sufficient for random-intercept models or balanced data

Kenward–Roger is:

More computationally intensive, but more accurate

Adjusts both the degrees of freedom and the estimated covariance matrix

Preferred in small samples, complex random effects, or when testing F-statistics

Used via pbkrtest::KRmodcomp() or lmerTest::anova(type = 3)

✅ Use Kenward–Roger when precision matters (especially for F-tests);
✅ Use Satterthwaite for quicker, approximate inference (especially for t-tests).


## Stepwise Selection

📌 Note on Stepwise Selection, dredge(), and Variable Screening
Stepwise Selection
Stepwise model selection procedures (e.g., forward, backward, or both) automatically add or remove predictors based on criteria like AIC, BIC, or p-values.

Though widely used, stepwise selection has serious limitations:

Inflated Type I error rates

Biased parameter estimates

Overfitting and unstable model structure

Results highly dependent on the data at hand and on variable entry order

It’s often data-driven rather than theory-driven, which undermines reproducibility and interpretability.

Alternative: MuMIn::dredge()
The dredge() function (from the MuMIn package) fits all possible subsets of a global model and ranks them by information criteria (e.g., AICc).

Advantages:

Provides a global view of model fit across variable combinations

Allows model averaging for uncertainty in variable selection

Caveats:

Can be computationally expensive

Not immune to the overfitting risks of stepwise procedures

Should be used with care and a limited, theory-justified set of predictors

Preliminary Variable Screening
In large datasets or models with many potential predictors, it’s common to:

Perform univariate tests or visual exploration to filter irrelevant variables

Use domain knowledge to pre-specify likely important variables

Assess multicollinearity (e.g., via VIF) before modeling

✅ Preliminary screening is acceptable if transparently reported and justified — it helps reduce noise and computational burden, especially before fitting complex models (e.g., GLMMs).

✅ Best Practices
Prefer theory-driven model building over automated selection.

If variable selection is necessary:

Use information criteria (AIC/AICc) over p-values

Report the full set of candidate models and selection process

Consider model averaging to account for model uncertainty

Always assess the stability and interpretability of the final model.


```{r}
# Load required package
library(car)
```


```{r}
# Set contrasts to contr.sum to allow Type III SS
options(contrasts = c("contr.sum", "contr.poly"))

# Fit a linear model
model <- lm(Sepal.Length ~ Species + Petal.Width + Species:Petal.Width, data = iris)

# Perform Type III ANOVA
Anova(model, type = 3, test = 'F')
```

```{r}
summary(model)
```
```{r}
m1 <- glm(Sepal.Length ~ Petal.Width + Species, data = iris, family = gaussian())
m2 <- glm(Sepal.Length ~ Species + Petal.Width + Species:Petal.Width, data = iris, family = gaussian())

anova(m1, m2, test = "LRT")
```

```{r}
drop1(m1)
```


```{r}
summary(m2)
```

```{r}
# Set contrasts to contr.sum to allow Type III SS
options(contrasts = c("contr.sum", "contr.poly"))

# Fit a linear model
model <- lm(Sepal.Length ~ Species + Petal.Width, data = iris)

# Perform Type III ANOVA
Anova(model, type = 3, test = 'F')
```

```{r}
# Set contrasts to contr.sum to allow Type III SS
options(contrasts = c("contr.sum", "contr.poly"))

# Fit a linear model
model <- lm(Sepal.Length ~ Species + Petal.Width, data = iris)

# Perform Type III ANOVA
Anova(model, type = 3, test = 'Chisq')
```

```{r}
m1 <- glm(Sepal.Length ~ Petal.Width, data = iris, family = gaussian())
m2 <- glm(Sepal.Length ~ Species + Petal.Width, data = iris, family = gaussian())

library(lme4)
anova(m1, m2)
```

```{r}
library(MuMIn)
m2 <- glm(Sepal.Length ~ Species + Petal.Width + Species:Petal.Width, data = iris, family = gaussian(), na.action = 'na.fail')

dredge(m2, extra = alist(AIC, BIC, ICOMP, Cp))
```


```{r}
anova(m2)
```

```{r}
anova(model)
```


```{r}
summary(model)
```

```{r}
summary(m2)
```

```{r}
mtcars$cyl <- factor(mtcars$cyl)
mtcars$am  <- factor(mtcars$am, labels = c("auto", "manual"))

m <- glm(mpg ~ cyl * wt + hp + qsec + am, data = mtcars,
         family = gaussian(), na.action = na.fail)

summary(m)
```
```{r}
m <- lm(mpg ~ cyl * wt + hp + qsec + am, data = mtcars, na.action = na.fail)

anova(m)
```
```{r}
dredge(m, extra = c(BIC, Cp))
```

```{r}
me <- lm(mpg ~ wt + qsec + am, data = mtcars, na.action = na.fail)

summary(me)
```

```{r}
Anova(m2, type = 3)
Anova(me, type = 3)
```

```{r}
Anova(m, type = 3)
```
```{r}
Anova(m2, type = 3)
```
```{r}
dredge(m2)
```

```{r}
mlme <- lmer(mpg ~ wt + qsec + am + (1|1), data = mtcars, na.action = na.fail)

lmerTest::anova(m2)
```
```{r}
library(pbkrtest)
KRmodcomp(m2)

m2
```




```{r}
dredge(m2)
```

```{r}
m2 <- glm(mpg ~ wt + qsec + am, data = mtcars, na.action = na.fail)
m3 <- glm(mpg ~ wt + qsec, data = mtcars, na.action = na.fail)
m4 <- glm(mpg ~ wt + am, data = mtcars, na.action = na.fail)
m5 <- glm(mpg ~ qsec + am, data = mtcars, na.action = na.fail)
```

```{r}
# ensure correct contrasts
options(contrasts = c("contr.sum", "contr.poly"))

# fit model
m2 <- lm(mpg ~ wt + qsec + am, data = mtcars)

# get Type III ANOVA
a3 <- Anova(m2, type = "III")

# extract sums of squares
ss_term <- a3["am", "Sum Sq"]
ss_res  <- a3["Residuals", "Sum Sq"]

# compute f^2
f2_am <- ss_term / ss_res

# degrees of freedom
df_term <- a3["am", "Df"]
df_res  <- a3["Residuals", "Df"]

# compute power
pwr.f2.test(u = df_term, v = df_res, f2 = f2_am, sig.level = 0.042)
```


```{r}
drop1(m2, )
```

```{r}
library(effectsize)
eta_squared(m2, partial = TRUE, type = 3)
```
```{r}
true_beta_am <- coef(m2)['am']

library(MASS)

simulate_lrt_power <- function(n = nrow(mtcars), B = 1000) {
  design <- model.matrix(~ wt + qsec + am, data = mtcars)
  beta <- coef(m2)
  sigma <- sigma(m2)
  pvals <- numeric(B)

  for (b in 1:B) {
    y_sim <- rnorm(n, mean = design %*% beta, sd = sigma)
    full_sim <- lm(y_sim ~ wt + qsec + am, data = mtcars)
    red_sim  <- lm(y_sim ~ wt + qsec, data = mtcars)

    lrt <- -2 * (logLik(red_sim)[1] - logLik(full_sim)[1])
    pvals[b] <- pchisq(lrt, df = 1, lower.tail = FALSE)
  }

  power_estimate <- mean(pvals < 0.05)
  return(power_estimate)
}

simulate_type3_power <- function(model_full, term = "am", B = 1000, alpha_nominal = 0.05) {
  X <- model.matrix(model_full)
  beta <- coef(model_full)
  sigma <- sigma(model_full)
  n <- nrow(X)

  pvals <- numeric(B)

  for (b in 1:B) {
    y_sim <- rnorm(n, mean = X %*% beta, sd = sigma)
    data_sim <- model.frame(model_full)
    data_sim[[as.character(formula(model_full)[[2]])]] <- y_sim

    model_b <- update(model_full, data = data_sim)
    a3 <- car::Anova(model_b, type = "III")
    pvals[b] <- a3[term, "Pr(>F)"]
  }

  mean(pvals < alpha_nominal)  # empirical power
}

simulate_type3_power(m2)
simulate_lrt_power()
```

```{r}
m2 <- lm(mpg ~ wt + qsec + am, data = mtcars)
m3 <- lm(mpg ~ wt + qsec, data = mtcars)  # null model (no am)

simulate_lrt_alpha <- function(full_model, reduced_model, B = 1000, alpha_nominal = 0.05) {
  # design matrix and parameters from reduced model (null)
  X <- model.matrix(reduced_model)
  beta <- coef(reduced_model)
  sigma <- sigma(reduced_model)
  n <- nrow(X)

  pvals <- numeric(B)

  for (b in 1:B) {
    # simulate new response under null
    y_sim <- rnorm(n, mean = X %*% beta, sd = sigma)

    # refit models with simulated data
    full_sim <- update(full_model, data = transform(mtcars, mpg = y_sim))
    red_sim  <- update(reduced_model, data = transform(mtcars, mpg = y_sim))

    # compute LRT
    lrt <- -2 * (logLik(red_sim)[1] - logLik(full_sim)[1])
    pvals[b] <- pchisq(lrt, df = 1, lower.tail = FALSE)
  }

  # return estimated alpha (false positive rate)
  list(
    estimated_alpha = mean(pvals < alpha_nominal),
    nominal_alpha = alpha_nominal
  )
}

simulate_type3_alpha(m2, B = 1000)
simulate_lrt_alpha(m2,m3, B = 1000)
```
```{r}
Anova(m2, type = 3)
```
```{r}
library(pwr)
pwr.f2.test(u = 1, v = 28, f2 = 0.1547, sig.level = 0.05)
```

```{r}
simulate_type3_alpha <- function(model_full, term = "am", B = 1000, alpha_nominal = 0.05) {
  # 1. Fit null model without the target term
  terms_all <- attr(terms(model_full), "term.labels")
  terms_null <- setdiff(terms_all, term)
  formula_null <- reformulate(terms_null, response = all.vars(formula(model_full))[1])
  model_null <- update(model_full, formula_null)

  # 2. Extract design matrix, coefficients, and sigma from null model
  X <- model.matrix(model_null)
  beta <- coef(model_null)
  sigma <- sigma(model_null)
  n <- nrow(X)

  # 3. Run simulation
  pvals <- numeric(B)

  for (b in 1:B) {
    y_sim <- rnorm(n, mean = X %*% beta, sd = sigma)
    data_sim <- model.frame(model_full)
    data_sim[[as.character(formula(model_full)[[2]])]] <- y_sim

    model_b <- update(model_full, data = data_sim)
    a3 <- Anova(model_b, type = "III")
    pvals[b] <- a3[term, "Pr(>F)"]
  }

  # Return estimated alpha
  list(
    estimated_alpha = mean(pvals < alpha_nominal),
    nominal_alpha = alpha_nominal
  )
}
```


```{r}
anova(m2, m3)
anova(m2, m4)
anova(m2, m5)
```


```{r}
Anova(m2, type = 3)
```
```{r}
library(glmnet)

# Prepare the design matrix (X) and response vector (y)
X <- model.matrix(mpg ~ wt + qsec + am, data = mtcars)[, -1]  # remove intercept column
y <- mtcars$mpg

```
```{r}
ridge_fit <- glmnet(X, y, alpha = 0)
```

```{r}
cv_ridge <- cv.glmnet(X, y, alpha = 0)
best_lambda <- cv_ridge$lambda.min

# Coefficients at optimal lambda
coef(cv_ridge, s = 'lambda.min')
```

