---
title: 'Turbidity Post-Cal Exploration'
author: 'Perry'
date: '`r Sys.Date()`'
format:
    html:
      toc: true
      highlight-style: github
      embed-resources: true
      self-contained-math: true
---

# Read in Data

```{r}
#| message: FALSE
#| warning: FALSE

library(tidyverse)
library(lme4)
library(lmerTest)
library(lattice)
library(emmeans)
```

```{r}
#| message: FALSE
#| warning: FALSE

df_turb <- read_csv(here::here('05_SpecialStudies/QA_TurbiditySensors/moreturbiditydata.csv'))

df_post <- df_turb %>% filter(Type == 'postcal')
df_pre <- df_turb %>% filter(Type == 'precal')

# all post-cal is new standard (24E) and all pre-cal is old standard (23F, except for 24B)
```

```{r}
df_turb <- df_post %>%
  mutate(PostCalCentered = Turbidity_FNU - 124, # center the data at 0
         PostCalDate = as.Date(Date, format = '%m/%d/%Y'))
```

# Assumptions

- sources of error accumulated between calibration and post-calibration are the same for all data points
  - ie. no significant differences in drift between sondes or sites

- sources of error accumulated between calibration and post-calibration are negligible
  - ie. the expected value at post-cal remains 124 (the standard)
  
- the person doing the calibrations did not significantly affect the values

- all sondes were calibrated in the old standard and post-calibrated in the new standard
  - **checking this one**

- the post-cal standard was 124 for all sondes

# Plots

## Original

```{r}
ggplot(df_turb, aes(x = PostCalCentered)) +
  geom_histogram(bins = 10, fill = 'steelblue', color = 'white') +
  theme_bw()
```

```{r}
ggplot(df_turb, aes(x = '', y = PostCalCentered)) +
  geom_boxplot(alpha = 0.8) +
  geom_hline(yintercept = 0, color = 'red', linetype = 'dashed', linewidth = 1) +
  geom_jitter(width = 0.2, shape = 21, size = 2, fill = 'gray') +
  theme_bw()
```

**Conclusions:**
Data is lower than expected, looks bimodal again.

## Population Split

```{r}
df_split <- df_turb %>%
  mutate(
    ProgramSplit = case_when(
      PostCalCentered < -5 ~ 'CEMP_low',
      PostCalCentered >= -5 ~ 'CEMP_high',
    )
  ) %>%
  mutate(ProgramSplit = factor(ProgramSplit, levels = c('CEMP_high','CEMP_low')))
```

```{r}
ggplot(df_split, aes(x = ProgramSplit, y = PostCalCentered)) +
  geom_boxplot(alpha = 0.8, outliers = FALSE) +
  geom_hline(yintercept = 0, color = 'red', linetype = 'dashed', linewidth = 1) +
  geom_jitter(width = 0.2, shape = 21, size = 2, fill = 'gray') +
  theme_bw()
```

# Models

## Original (Mod1)

```{r}
mod1 <- lm(PostCalCentered ~ 1, data = df_turb)

summary(mod1)
```

```{r}
#| layout-ncol: 2

plot(mod1)
```

**Conclusions:**

Not a great fit, look into split model.

## Split (Mod3)

```{r}
mod2r <- lmer(PostCalCentered ~ (1|ProgramSplit), data = df_split, REML = FALSE)

summary(mod2r)
```

```{r}
ranef(mod2r)
dotplot(ranef(mod2r, condVar = TRUE))
```

```{r}
mod2f <- lm(PostCalCentered ~ ProgramSplit, data = df_split)

summary(mod2f)
```

```{r}
#| layout-ncol: 2

plot(mod2f)
```

```{r}
emm3f <- emmeans(mod2f, ~ ProgramSplit)
summary(emm3f, infer = TRUE)
```

**Conclusions:**

The high-valued data is not statistically different from 124, but the low-valued data is.

## Compare models

```{r}
AIC(mod1, mod2f)
BIC(mod1, mod2f)
```

**Conclusions:**

We prefer the insight from our models that account for the bimodal nature of the program data.

# More Plots

```{r}
ggplot(df_split, aes(x = PostCalDate, y = PostCalCentered+124, fill = ProgramSplit)) +
  geom_hline(yintercept = 124, color = 'black', linetype = 'dashed', linewidth = 0.8) +
  geom_point(size = 2.5, shape = 21, color = 'black') +
  theme_bw()
```

# Conclusions

Something appears to have happened after July.

```{r}
# write_csv(df_split, '05_SpecialStudies/QA_TurbiditySensors/split_data.csv')
```
