## Title: EMLassemblyline template for EDI
## Purpose: a template to create relevant metadata for EDI datasets

Instructions are not exhaustive, use in conjunction with "DIY Instructions EMLassemblyLine in R.docx"

```{r}
# load packages
library(EMLassemblyline)
```

## Variables
Set the variables that will be referenced throughout the notebook.

```{r Variables}
datPath <- 'C:/R/FishRegressions/EDI/Data'

# path to empty folder to populate metadata files in
metaDatPath <- 'C:/R/FishRegressions/EDI/Metadata'

# path to empty folder to populate EML files
emlPath <- 'C:/R/FishRegressions/EDI/EML'

# name(s) of dataset files
datTables <- c('BiomassRegression_Compiled.csv',
               'BiomassRegression_Raw.csv') 

# define which license to use
ediLicense = 'CC0'

# define which keywordThesasurs are used
thesauruses = c('lter')
```

## Metadata
Create the metadata templates.

```{r Create Metadata Templates}
# create core metadata templates
template_core_metadata(
  metaDatPath,
  license = ediLicense,
  file.type = '.txt',
  write.file = TRUE
  )

# create attribute tables for data tables; will be placed with other metadata
template_table_attributes(
  path = metaDatPath,
  data.path = datPath,
  data.table = datTables,
  write.file = TRUE
  )

# create categorical codes for categorical variables
template_categorical_variables(
    path = metaDatPath,
    data.path = datPath,
    write.file = TRUE
    )
```

## Edit Metadata Files
Note: for all metadata .txt files, do not use special characters.

1) Open "abstract.txt" and type out abstract.

2) Open "methods.txt" and add in methods.

3) Open "additional_info.txt" and add any additional info not covered in the abstract/methods.

4) Open "personnel.txt" and enter relevant information.

5) Open "keywords.txt" and add in keywords (can be multiple words, hit enter after each entry). Can use keywords from a controlled vocabulary, such as LTER (https://vocab.lternet.edu/vocab/vocab/index.php), or write your own. Leave "keywordThesaurus" blank. Once finished, run the code below to autopopulate "keywordThesaurus" for all keywords that are from LTER.

```{r KeywordThesarus}
# validate thesarus keywords
for (thesaurus in thesauruses) {
  EDIutils::validate_keywords(
    path = metaDatPath,
    cv = thesaurus
    )
}
```

6a) Add units to template attribute tables. These files are stored in the the metadata path and start with "attributes_". Replace !insert units here! with correct units; if there are none, use "dimensionless". View the dictionary below for list of standard units.

```{r Unit Dictionary}
view_unit_dictionary()
```

If the unit is not in this directory:
6b) Create the unit under "custom_units.txt".

7) Add geographic coverage. If data contains lat/lon values, run the code below. If not:
- create a file in the metadata folder called "geographic_coverage.txt". Treat it as a tab-delimited file.
- add column headers: geographicDescription, nwestBoundingCoordinate, eastBoundingCoordinate, northBoundingCoordinate, southBoundingCoordinate

```{r}
extract_geocoverage(
  path = metaDatPath,
  data.path = datPath,
  data.file = 'name',
  lat.col = "lake_latitude",
  lon.col = "lake_longitude",
  site.col = "lake_name"
  )
```


## Make EML
Edit the parameters in this code as needed.

```{r}
# Run this function
make_eml(path = metaDatPath, # path to metadata
         data.path = datPath, # path to data
         eml.path = emlPath,
         dataset.title = 'Synthesized Dataset of Length-Weight Regression Coefficients for Delta Fish',
         data.table = datTables,
         data.table.name = c('Biomass Regressions - Compiled Data','Biomass Regressions - Raw Data'),
         data.table.description = c('Average values for the length-weight regression parameters, separated by length type.', 
                                    'Raw data used to calculate the average length-weight regression parameters.'),
         data.table.quote.character = c('\"', '\"'),
         other.entity = c('RegressionCalc_Data.csv','RegressionCalc_Code.R'),
         other.entity.name = c('Regression Calculations - Data','Regression Calculations - Code'),
         other.entity.description = c('The raw data used to calculate the parameters for fish species not found in the database or literature (the "calc" source in the compiled dataset).', 'The code used to calculate the parameters for fish species not found in the database or literature (the "calc" source in the compiled dataset).'),
         temporal.coverage = c('1950-01-01', '2018-12-31'),
         geographic.description = 'San Francisco Bay-Delta',
         geographic.coordinates = c('38.369', '-121.262', '37.678', '-122.503'), # north east south west
         maintenance.description = 'completed',
         write.file = TRUE,
         user.id = 'sperry',
         user.domain = 'EDI',
         package.id = 'edi.508.1'
         )
```

