---
title: "EMP Continuous WQ Metrics - 2023"
author: "Ted Flynn"
date: "2024-02-16"
output: html_document
---
Process and plot Wagner rating data from C-EMP sondes.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
                      
suppressWarnings(suppressMessages(library(lubridate)))
suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(RColorBrewer)))
suppressWarnings(suppressMessages(library(janitor)))
suppressWarnings(suppressMessages(library(here)))
suppressWarnings(suppressMessages(library(scales)))

# Set output directory 
output <- here("04_Other","performance-metrics","output")

# Set visual theme in ggplot
theme_set(theme_bw())

# Suppress summarise info
options(tidyverse.quiet = TRUE)
options(dplyr.summarise.inform = FALSE)

# Define negative %in% operator
`%nin%` = Negate(`%in%`)
                      
```

## Read in Raw Data
Import the raw data from C-EMP and perform basic QA-QC. Check station 
completeness and screen for duplicates.

```{r read input, echo = TRUE, results = FALSE}
# Import C-EMP data files 
df_CEMP <- read_csv(file = here("04_Other",
                                "performance-metrics",
                                "data",
                                "2023_wagner.csv"),
                    show_col_types = FALSE)

## Remove date time column (will only consider rating by date)
df_CEMP <- df_CEMP %>% 
  select(site:visit_date,lname:sonde_verif) 

# Check for missing dates
df_CEMP %>% filter(is.na(visit_date)) # No missing dates

# Select only needed columns
df_CEMP <- df_CEMP %>% 
  select(site:visit_date,analyte_name,rating)

# Check for station completeness
unique(df_CEMP$site)
summary(unique(df_CEMP$site)) # Lists all 15 sites

# Check for duplicated lines with 
df_CEMP %>% janitor::get_dupes(site:rating) # 1 duplicated row

# Retain only distinct rows (remove duplicates)
df_CEMP <- distinct(df_CEMP)


```

## Process Sonde Data
Calculate sonde rating intervals and summarize.

```{r process CEMP data, echo = TRUE}
# Get ratings to sort in order good -> bad
rating.order <- c("Excellent","Good","Fair","Poor","MAL")

df_CEMP$rating <- factor(as.character(df_CEMP$rating), 
                             levels = rating.order)

df_CEMP_w <- pivot_wider(df_CEMP, names_from = analyte_name, values_from = rating)

# Sort samples by date (early to late)
# Add previous exchange data as end point for range of rating data

df_CEMP_w <- df_CEMP_w %>% 
  group_by(site) %>% 
  arrange(visit_date) %>% 
  mutate(start_date = lag(visit_date)) %>%
  ungroup()

df_CEMP_w <- df_CEMP_w %>% 
  mutate(start_date = case_when(is.na(start_date) ~ ymd("2023-01-01"),
                                TRUE ~ ymd(start_date)))

# Relocate Start Date column
df_CEMP_w <- df_CEMP_w %>% 
  relocate(start_date, .before = visit_date)

# Calculate number of days between sonde exchanges
df_CEMP_w <- df_CEMP_w %>%
  mutate(days_rated = as.integer(visit_date - start_date))


```

## Expand rating data from intervals to daily
Only data rating intervals are given but we need a daily value to properly
graph the results. Also remove turbidity data because it is not useful
to apply Wagner method to this data.
```{r expand data, echo = TRUE}
# Create tibble with exchange intervals for each station
df_CEMP_l <- pivot_longer(df_CEMP_w, 
                          names_to = "Parameter", 
                          values_to = "Rating",
                          cols = `Water Temperature`:Turbidity)

df_CEMP_l <- df_CEMP_l %>% 
  select(site:visit_date,Parameter:Rating)

# Fill in ratings data for each individual day
df_CEMP_c <- df_CEMP_l %>% 
  mutate(across(ends_with("_date"), ymd)) %>% 
  mutate(Date = map2(start_date, 
                     visit_date, ~ seq.Date(.x, .y, by = "1 day")), 
         .before = start_date) %>% 
  unnest(Date) %>% 
  select(!ends_with("_date"))

# Convert date formats and add Julian Date
df_CEMP_c <- df_CEMP_c %>% 
  mutate(Julian = yday(Date)) %>% 
  ungroup()

# Remove turbidity data and data after 1/1/2024
df_CEMP_c <- df_CEMP_c %>% 
  filter(Parameter != "Turbidity") %>% 
  filter(Date <= "2024-01-01")

```

## Results
Plot sonde rating for all stations by probe.

```{r plot CEMP, echo = FALSE, fig.cap = "C-EMP Sonde Ratings.", fig.align="center"}
# Plot rating data faceted by probe
colors = c("darkgreen","darkblue","yellow3","darkred","purple")

p1 <- ggplot(data = df_CEMP_c, aes(x = Date,
                                   y = site,
                                   color = Rating)) +
  geom_line(linewidth = 5) +
  scale_color_manual(values = colors) + 
  scale_y_discrete(limits=rev) +
  scale_x_date(breaks = ymd(c("2023-01-01","2023-05-01","2023-09-01","2024-01-01")),
               labels = date_format("%b"))

p1 +
  labs(x = "Daily Rating of Data Collected",
       y = "Station ID",
       fill = "Sonde Rating",
       title = "EXO2 Probe Ratings - 2023") +
  facet_wrap(Parameter ~ ., ncol = 4)

ggsave(path = output,
       filename = "CEMP_sonde_scores_test.png", 
       device = "png",
       scale=1.0, 
       units="in",
       height=6.5,
       width=9, 
       dpi="print")
```

Plot sonde rating for all probes at each station.

```{r, echo = FALSE, fig.cap = "C-EMP Sonde Ratings for each station.", fig.align="center"}
# Plot rating data faceted by station
p2 <- ggplot(data = df_CEMP_c, aes(x = Date,
                                   y = Parameter,
                                   color = Rating)) +
  geom_line(linewidth = 5) +
  scale_color_manual(values = colors) + 
  scale_y_discrete(limits=rev) +
  scale_x_date(breaks = ymd(c("2023-01-01","2023-05-01","2023-09-01","2024-01-01")),
               labels = date_format("%b"))

p2 +
  labs(x = "Daily Rating of Data Collected)",
       y = "Sensor",
       fill = "Sonde Rating",
       title = "EXO2 Probe Ratings - 2023 ") +
  facet_wrap(site ~ ., ncol = 5)

ggsave(path = output,
       filename = "CEMP_sonde_scores_2023.png", 
       device = "png",
       scale=1.0, 
       units="in",
       height=6.5,
       width=9, 
       dpi="print")
```

