# Clean and Format 1975-2007 Phyto Data

# Initial setup

```{r}
library(pespr)
library(lubridate)
library(tidyverse)
source(here::here('00_GlobalFunctions/functions.R'))
```


## Read in data

```{r}
# phyto data
fp_raw <- abs_path_emp('Phytoplankton/AQUARIUS Samples Database/Database Migration/Raw Phyto Data/EMP_Phyto Data 75-07.csv')

df_phyto <- read_csv(fp_raw, col_types = cols(.default = col_character())) %>%
  rename(Date = SampleDate,
         Time = SampleTime,
         Station = StationCode)


# tin data
load(here::here('03_Phyto/aquarius/historical-imports/data/tins.rda'))

# dwq data
load(here::here('03_Phyto/aquarius/historical-imports/data/dwq_data.rda'))

# phyto taxa data
df_taxon <- pespr::phyto_taxonomy
```

# Dates

Convert Dates
```{r}
df_phyto <- df_phyto %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y'))

df_phyto <- df_phyto %>%
  mutate(Time = str_replace(Time, "^0(\\d:)", "\\1"))

df_wq <- df_wq %>%
  mutate(Date = as.Date(Date, format = '%Y-%m-%d'))
```

```{r}
unique(df_phyto$Time)
```
# Times

Fix non-military

```{r}
df_phyto <- df_phyto %>%
  mutate(Time = ifelse(
    !is.na(Time) & as.numeric(sub(':.*', '', Time)) < 5,
    sprintf('%d:%s',
            as.numeric(sub(':.*', '', Time)) + 12,
            sub('^[^:]*:', '', Time)),
    Time
  ))
```

Round Times

```{r}
round_time <- function(x) {
  out <- rep(NA_character_, length(x))
  ok <- !is.na(x)
  out[ok] <- format(round_date(as_datetime(paste0('2000-01-01 ', x[ok], ':00')),
                               '5 minutes'), '%H:%M')
  out
}

df_phyto <- df_phyto %>% mutate(Time = round_time(Time))

unique(df_phyto$Time)
```
# Stations

Rename EZs
```{r}
df_phyto <- df_phyto %>%
  mutate(Station = recode(Station, 'EZ2' = 'LSZ2', 'EZ6' = 'LSZ6'))
```


Determine stations that don't match with DWQ

TODO: get metadata from Tiffany on these

```{r}
setdiff(df_phyto$Station, df_wq$Station)
```

# Data Checks

## Numeric Metadata

```{r}
unique(df_phyto$CellArea)
```

```{r}
unique(df_phyto$GridArea)
```

```{r}
unique(df_phyto$NumberOfFields)
```
```{r}
unique(df_phyto$Magnification)
```

```{r}
unique(df_phyto$SettlingVolume)
```

Looks good, convert to numeric.

```{r}
df_phyto <- df_phyto %>%
  mutate(
    CellArea = as.numeric(CellArea),
    GridArea = as.numeric(GridArea),
    NumberOfFields = as.numeric(NumberOfFields),
    Magnification = as.numeric(Magnification),
    SettlingVolume = as.numeric(SettlingVolume)
  )
```

```{r}
colnames(df_phyto)
```

## String Metadata

For group, convert n/p to NA

```{r}
df_phyto <- df_phyto %>%
  mutate(Group = na_if(Group, 'n/p'))

unique(df_phyto$Group)
```

```{r}
unique(df_phyto$Remark)
```
## Check Data

```{r}
unique(df_phyto$Count)
```

```{r}
unique(df_phyto$`Organisms per mL`)
```

Looks good, convert.

```{r}
df_phyto <- df_phyto %>%
  mutate(
    Count = as.numeric(Count),
    `Organisms per mL` = as.numeric(`Organisms per mL`)
  )
```

Double check numbers.

```{r}
ggplot(df_phyto, aes(x = "", y = log(Count))) +
  geom_boxplot(outlier.shape = NA, fill = "lightblue", alpha = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.5, shape = 21) +
  theme_minimal()
```

```{r}
ggplot(df_phyto, aes(x = "", y = log(`Organisms per mL`))) +
  geom_boxplot(outlier.shape = NA, fill = "lightblue", alpha = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.5, shape = 21) +
  theme_minimal()
```

## Check Taxon

Correct spellings

```{r}
taxon_fixes <- tribble(
  ~Taxon,                        ~new_Taxon,                               ~new_Species,      ~new_Genus,
  'Diatoma hiemale',             'Diatoma hyemalis',                       'hyemalis',        NA,
  'Diatoma elongatum',           'Diatoma elongata',                       'elongata',        NA,
  'Chlorosphaeralean sp.',       'Chlorosphaera sp.',                      NA,                'Chlorosphaera',
  'Treubaria setigerum',         'Treubaria setigera',                     'setigera',        NA,
  'Attheya zachariasi',          'Attheya zachariasii',                    'zachariasii',     NA,
  'Microcystis flosaquae',       'Microcystis flos-aquae',                 'flos-aquae',      'Microcystis',
  'Bacillaria paxillifer',       'Bacillaria paxillifera',                 'paxillifera',     NA,
  'Diploneis didyma',            'Diploneis didymus',                      'didymus',         NA,
  'Chlorallantus sp.',           'Chlorallanthus sp.',                     NA,                'Chlorallanthus',
  'Anabaena flosaquae',          'Anabaena flos-aquae',                    'flos-aquae',      'Anabaena',
  'Dolichospermum flosaquae',    'Dolichospermum flos-aquae',              'flos-aquae',      'Dolichospermum',
  'Aphanizomenon flosaquae',     'Aphanizomenon flos-aquae',               'flos-aquae',      'Aphanizomenon',
  'Centric Diatom sp.',          'Unidentified Centric Diatom sp. 1',      NA,                NA,
  'Pennate Diatom sp.',          'Unidentified Pennate Diatom sp. 1',      NA,                NA,
  'Unidentified flagellate',     'Unidentified Flagellate sp. 1',          NA,                NA,
  'Nanoflagellates',             'Unidentified Nanoflagellate sp. 1',      NA,                NA,
  'Chodatella wratislaviensis',  'Chodatella wratislawiensis',             'wratislawiensis', NA,
  'Eucampia zoodiacus',          'Eucampia zodiacus',                      'zodiacus',        NA,
  'NONE NONE',                   'No organisms found',                     NA,                NA,
  'Unidentified sp.',            'Unidentified sp. 1',                     NA,                NA,
  'Unknown unknown',             'Unidentified sp. 1',                     NA,                NA,
)

df_phyto <- df_phyto %>%
  left_join(taxon_fixes, by = 'Taxon') %>%
  mutate(
    Species = coalesce(new_Species, Species),
    Genus   = coalesce(new_Genus, Genus),
    Taxon   = coalesce(new_Taxon, Taxon)
  ) %>%
  select(-new_Taxon, -new_Species, -new_Genus)
```

Compare to TIN code list

```{r}
df_eataxa <- anti_join(df_phyto, df_codes, by = 'Taxon') %>% distinct(Taxon)

df_eataxa
```

Compare ones not in TIN codes to taxonomy list

```{r}
df_newinfo <- inner_join(df_taxon, df_eataxa, by = 'Taxon') %>%
  select(Taxon, CurrentTaxon)
```

See what's left

```{r}
df_missing <- anti_join(df_eataxa, df_taxon, by = 'Taxon')

df_missing
```

# Special Study Data

```{r}
df_special <- df_phyto %>%
  filter(MethodCode == 'SPECIAL')
```

Check unique dates and stations
```{r}
df_special %>%
  distinct(Date, Station) %>%
  group_by(Date) %>%
  summarise(Station = paste(sort(Station), collapse = ", "), .groups = "drop")
```

Check dates that match with routine data
```{r}
special_wq_matches <- df_special %>%
  distinct(Station, Date) %>%
  inner_join(df_wq %>% distinct(Station, Date), by = c("Station", "Date")) %>%
  arrange(Station, Date)

special_wq_routine_check <- special_wq_matches %>%
  left_join(
    df_routine %>% distinct(Station, Date) %>% mutate(has_routine = TRUE),
    by = c("Station", "Date")
  ) %>%
  mutate(has_routine = if_else(is.na(has_routine), FALSE, has_routine))

special_wq_routine_check
```

```{r}
special_wq_routine_check_month <- df_special %>%
  mutate(Year = year(Date), Month = month(Date)) %>%
  group_by(Station, Year, Month) %>%
  summarise(special_dates = paste(sort(unique(Date)), collapse = ", "), .groups = "drop") %>%
  inner_join(
    df_wq %>%
      mutate(Year = year(Date), Month = month(Date)) %>%
      group_by(Station, Year, Month) %>%
      summarise(wq_dates = paste(sort(unique(Date)), collapse = ", "), .groups = "drop"),
    by = c("Station", "Year", "Month")
  ) %>%
  left_join(
    df_routine %>%
      mutate(Year = year(Date), Month = month(Date)) %>%
      distinct(Station, Year, Month) %>%
      mutate(has_routine_month = TRUE),
    by = c("Station", "Year", "Month")
  ) %>%
  mutate(has_routine_month = if_else(is.na(has_routine_month), FALSE, has_routine_month)) %>%
  arrange(Station, Year, Month)

special_wq_routine_check_month
```

# Routine Data

```{r}
df_routine <- df_phyto %>%
  filter(MethodCode == 'PHYTO') %>%
  mutate(Time = paste0(Time, ":00"))


df_routine <- df_routine %>%
  mutate(Time = str_replace(Time, "^0(\\d:)", "\\1"))
```

## Check stations

Stations that have phyto data but no corresponding DWQ data

TODO: get some sort of info on these

```{r}
setdiff(df_routine$Station, df_wq$Station)
```

```{r}
missing_stations <- setdiff(df_routine$Station, df_wq$Station)

df_routine %>%
  filter(Station %in% missing_stations) %>%
  mutate(Year = format(Date, "%Y"), Month = format(Date, "%m")) %>%
  distinct(Station, Year, Month) %>%
  arrange(Station, Year, Month)
```

## Check Dates

Compare against DWQ

```{r}
missing_stations <- setdiff(df_routine$Station, df_wq$Station)

wq_events <- df_wq %>%
  mutate(Year = year(Date), Month = month(Date)) %>%
  distinct(Year, Month, Station, Date) %>%
  group_by(Year, Month, Station) %>%
  summarise(
    n_events_wq = n(),
    wq_dates = paste(sort(Date), collapse = ", "),
    .groups = "drop"
  )

phyto_events <- df_routine %>%
  filter(!Station %in% missing_stations) %>%      # exclude stations that were never in df_wq to begin with
  mutate(Year = year(Date), Month = month(Date)) %>%
  distinct(Year, Month, Station, Date) %>%
  group_by(Year, Month, Station) %>%
  summarise(
    n_events_phyto = n(),
    phyto_dates = paste(sort(Date), collapse = ", "),
    .groups = "drop"
  )

comparison <- wq_events %>%
  full_join(phyto_events, by = c("Year", "Month", "Station")) %>%
  mutate(
    status = case_when(
      is.na(n_events_wq)             ~ "phyto only",
      is.na(n_events_phyto)          ~ "wq only (missing in phyto)",
      n_events_wq != n_events_phyto  ~ "mismatch in # events",
      TRUE                           ~ "match"
    )
  ) %>%
  arrange(status, Year, Month, Station)

problems <- comparison %>% filter(status != "match")
problems %>%
  arrange(Year, Month, Station) %>%
  filter(status != 'wq only (missing in phyto)')
```

Apply fixes
```{r}
df_routine <- df_routine %>%

  mutate(Time = if_else(Date == as.Date('2004-07-30') & Station == 'C3',
                         '16:55', Time)) %>%

  filter(!(Date == as.Date('1987-08-19') & Station == 'D7' & Taxon == 'NONE')) %>%
  
  mutate(Date = if_else(Date == as.Date('1994-11-29') & Station == 'D7',
                         as.Date('1994-11-16'), Date)) %>%
  
  mutate(Date = if_else(Date == as.Date('1998-07-15') & 
                           Station %in% c('D26', 'MD10A', 'P8', 'D28A'),
                         as.Date('1998-07-16'), Date))
```

Check and update mismatched dates (one event per Month/Year/Station in each)
- update to DWQ dates

```{r}
wq_single <- df_wq %>%
  mutate(Year = year(Date), Month = month(Date)) %>%
  distinct(Year, Month, Station, Date) %>%
  group_by(Year, Month, Station) %>%
  summarise(
    n_wq = n(),
    date_wq = paste(sort(Date), collapse = ", "),
    .groups = "drop"
  ) %>%
  filter(n_wq == 1)

routine_single <- df_routine %>%
  mutate(Year = year(Date), Month = month(Date)) %>%
  distinct(Year, Month, Station, Date) %>%
  group_by(Year, Month, Station) %>%
  summarise(
    n_routine = n(),
    date_routine = paste(sort(Date), collapse = ", "),
    .groups = "drop"
  ) %>%
  filter(n_routine == 1)

single_event_mismatches <- wq_single %>%
  inner_join(routine_single, by = c("Year", "Month", "Station")) %>%
  filter(date_wq != date_routine) %>%
  arrange(Year, Month, Station)

single_event_mismatches
```

```{r}
df_routine <- df_routine %>%
  mutate(Year = year(Date), Month = month(Date)) %>%
  left_join(
    single_event_mismatches %>%
      transmute(Year, Month, Station, new_date = as.Date(date_wq)),
    by = c("Year", "Month", "Station")
  ) %>%
  mutate(Date = if_else(!is.na(new_date), new_date, Date)) %>%
  select(-Year, -Month, -new_date)
```

Check Times

```{r}
wq_times <- df_wq %>%
  distinct(Station, Date, Time) %>%
  rename(Time_wq = Time)

routine_times <- df_routine %>%
  distinct(Station, Date, Time) %>%
  rename(Time_routine = Time)

time_mismatches <- wq_times %>%
  inner_join(routine_times, by = c("Station", "Date")) %>%
  filter(Time_wq != Time_routine | (is.na(Time_wq) != is.na(Time_routine))) %>%
  arrange(Station, Date)
```

Check which Dates have > Time
```{r}
routine_times %>% count(Station, Date) %>% filter(n > 1)
```

Fix those
```{r}
df_routine <- df_routine %>%
  filter(!(Date == '1992-03-26' & Station == 'D7' & Time == '9:15:00')) %>%
  filter(!(Date == '1994-10-26' & Station == 'D7' & Time == '8:35:00')) %>%
  mutate(
    Taxon = if_else(
      Date == '1994-11-23' & Station == 'C3' & Time == '13:30:00' & Taxon == 'Unidentified Flagellate sp. 1',
      'Unidentified Flagellate sp. 2',
      Taxon
    )
  )
```

Update Times
```{r}
df_routine <- df_routine %>%
  left_join(
    time_mismatches %>%
      select(Station, Date, Time_routine, Time_wq),
    by = c("Station", "Date", "Time" = "Time_routine")
  ) %>%
  mutate(Time = if_else(!is.na(Time_wq), Time_wq, Time)) %>%
  select(-Time_wq)
```

## Check missing data

```{r}
# build the full Month x Year grid, then flag presence/absence
routine_coverage <- df_routine %>%
  mutate(Year = year(Date), Month = month(Date)) %>%
  distinct(Year, Month) %>%
  mutate(has_data = TRUE) %>%
  right_join(
    expand_grid(Year = 1975:2007, Month = 1:12),
    by = c("Year", "Month")
  ) %>%
  mutate(has_data = if_else(is.na(has_data), FALSE, has_data))

ggplot(routine_coverage, aes(x = factor(Month), y = factor(Year), fill = has_data)) +
  geom_tile(color = "white", linewidth = 0.2) +
  scale_fill_manual(values = c("TRUE" = "steelblue", "FALSE" = "grey90"),
                     labels = c("TRUE" = "Data", "FALSE" = "No data"),
                     name = NULL) +
  scale_x_discrete(labels = month.abb) +
  scale_y_discrete(limits = rev(as.character(1975:2007))) +
  labs(x = "Month", y = "Year", title = "df_routine data coverage, 1975–2007") +
  theme_minimal() +
  theme(panel.grid = element_blank())
```
# Format Data
TODO: add lab column, format for DB

# Save data
Combine routine and special
```{r}
df_newphyto <- bind_rows(df_routine, df_special)
```

Remove trailing zeros in Time
```{r}
df_newphyto <- df_newphyto %>%
  mutate(Time = str_replace(Time, ":00$", ""))
```

Write data
```{r}
save(df_newphyto, file = here::here('03_Phyto/aquarius/historical-imports/data/1975-2007.rda'))
```

