# Clean and Format 2008-2013 Phyto Data

Special Studies aren't already marked.

# Initial setup

```{r}
library(pespr)
library(lubridate)
library(tidyverse)
source(here::here('00_GlobalFunctions/functions.R'))
```


## Read in data

```{r}
# phyto data
fp_raw <- abs_path_emp('Phytoplankton/AQUARIUS Samples Database/Database Migration/Raw Phyto Data/EMP_Phyto Data 08-13.csv')

df_phyto <- read_csv(fp_raw, col_types = cols(.default = col_character())) %>%
  rename(Date = SampleDate,
         Time = SampleTime,
         Station = StationCode,
         Taxon = TAXON)


# tin data
load(here::here('03_Phyto/aquarius/historical-imports/data/tins.rda'))

# dwq data
load(here::here('03_Phyto/aquarius/historical-imports/data/dwq_data.rda'))

# phyto taxa data
df_taxon <- pespr::phyto_taxonomy
```

# Dates

Convert Dates
```{r}
df_phyto <- df_phyto %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y'))

df_phyto <- df_phyto %>%
  mutate(Time = str_replace(Time, '^0(\\d:)', '\\1'))

df_wq <- df_wq %>%
  mutate(Date = as.Date(Date, format = '%Y-%m-%d'))

start_date <- as.Date('2008-01-01')
end_date   <- as.Date('2013-12-30')

df_wq <- df_wq %>%
  filter(between(Date, start_date, end_date))
```

```{r}
unique(df_phyto$Time)
```
# Times

Add in colons

```{r}
fix_time <- function(x) {
  case_when(
    is.na(x)                      ~ NA_character_,
    str_detect(x, ':')            ~ x, 
    nchar(x) == 3                  ~ str_replace(x, '^(\\d)(\\d{2})$', '\\1:\\2'),
    nchar(x) == 4                  ~ str_replace(x, '^(\\d{2})(\\d{2})$', '\\1:\\2'),
    TRUE                            ~ x
    )
}

df_phyto <- df_phyto %>%
  mutate(Time = fix_time(Time))

unique(df_phyto$Time)
```


Fix non-military

```{r}
df_phyto <- df_phyto %>%
  mutate(Time = ifelse(
    !is.na(Time) & as.numeric(sub(':.*', '', Time)) < 5,
    sprintf('%d:%s',
            as.numeric(sub(':.*', '', Time)) + 12,
            sub('^[^:]*:', '', Time)),
    Time
  ))
```

Round Times

```{r}
round_time <- function(x) {
  out <- rep(NA_character_, length(x))
  ok <- !is.na(x)

  rounded <- round_date(
    as_datetime(paste0('2000-01-01 ', x[ok], ':00')),
    '5 minutes'
  )

  out[ok] <- sub('^0', '', format(rounded, '%H:%M'))
  out
}

df_phyto <- df_phyto %>% mutate(Time = round_time(Time))

unique(df_phyto$Time)
```

# Stations

```{r}
unique(df_phyto$Station)
```


Rename EZs
```{r}
df_phyto <- df_phyto %>%
  mutate(Station = recode(Station, 'EZ2' = 'LSZ2', 'EZ6' = 'LSZ6'))
```


Determine stations that don't match with DWQ

```{r}
setdiff(df_phyto$Station, df_wq$Station)
```

# Data Checks

## Numeric Metadata

```{r}
unique(df_phyto$`Volume Received (mL)`)
```

```{r}
unique(df_phyto$`Volume Analyzed (mL)`)

df_phyto <- df_phyto %>%
  mutate(`Volume Analyzed (mL)` = round(as.numeric(`Volume Analyzed (mL)`), 2))
```

```{r}
unique(df_phyto$`Percent of Sample Counted`)

# remove percent sign
df_phyto <- df_phyto %>%
  mutate(`Percent of Sample Counted` = str_remove(`Percent of Sample Counted`, '%'))
```

```{r}
unique(df_phyto$`Field-of-view (mm2)`)

df_phyto <- df_phyto %>%
  mutate(`Field-of-view (mm2)` = round(as.numeric(`Field-of-view (mm2)`), 3))
```

```{r}
unique(df_phyto$`Slide Chamber Area (mm2)`)
```
```{r}
unique(df_phyto$`Number of Transects Counted`)

df_phyto <- df_phyto %>%
  mutate(`Number of Transects Counted` = round(as.numeric(`Number of Transects Counted`), 2))
```
```{r}
unique(df_phyto$`Number of Fields Counted`)

df_phyto <- df_phyto %>%
  mutate(`Number of Fields Counted` = if_else(
    !is.na(`Number of Fields Counted`) & str_detect(`Number of Fields Counted`, '\\.\\d{3,}'),
    as.character(round(as.numeric(`Number of Fields Counted`), 2)),
    `Number of Fields Counted`
  ))
```
```{r}
unique(df_phyto$`Number of Fields Counted`)

df_phyto <- df_phyto %>%
  mutate(`Number of Fields Counted` = if_else(
    !is.na(`Number of Fields Counted`) & str_detect(`Number of Fields Counted`, '\\.\\d{3,}'),
    as.character(round(as.numeric(`Number of Fields Counted`), 2)),
    `Number of Fields Counted`
  ))
```

```{r}
unique(df_phyto$`Total Number of Fields`)
```

```{r}
df_phyto %>%
  filter(`Total Number of Fields` == '6050')
```



Looks good, convert to numeric.

```{r}
colnames(df_phyto)
```

## String Metadata

For group, convert n/p to NA

```{r}
unique(df_phyto$`Diatom or SoftBody`)
```
Fix Filament Group Codes

```{r}
df_phyto <- df_phyto %>%
  mutate(`Colony or Filament Group Code` = str_to_lower(`Colony or Filament Group Code`)) %>%
  mutate(`Colony or Filament Group Code` = case_when(
    `Colony or Filament Group Code` %in% c('0', 'n/p', '#n/a') ~ NA_character_,
    `Colony or Filament Group Code` == '1' ~ 'i',
    TRUE ~ `Colony or Filament Group Code`
  ))

unique(df_phyto$`Colony or Filament Group Code`)
```
Shape

```{r}
unique(df_phyto$Shape)
```

```{r}
shape_lookup <- c(
  'cuboid'                                = 'cuboid',
  'ellipsoid'                             = 'ellipsoid',
  'other: achnanthes type'                = 'other: achnanthes type',
  'other: achnanthes'                     = 'other: achnanthes type',
  'cylinder'                               = 'cylinder',
  'sphere'                                 = 'sphere',
  'ellipsoid + cone'                       = 'ellipsoid + cone',
  'ellipsoid+cone'                         = 'ellipsoid + cone',
  'ellipsoid with cone'                    = 'ellipsoid + cone',
  'ellipsoid + 2 cones + cylinder'         = 'ellipsoid + two cones + cylinder',
  'ellipsoid + two cones + cylinder'       = 'ellipsoid + two cones + cylinder',
  'peridinium'                             = 'other: peridinium type',
  'other: peridinium type'                 = 'other: peridinium type',
  'two cones'                              = 'two cones',
  'other: crucigenia tetrapedia type'      = 'other: crucigenia tetrapedia type',
  'box'                                    = 'box',
  'ovoid'                                  = 'ovoid',
  'two wedges'                             = 'two wedges',
  'prism on elliptic base'                 = 'prism on elliptical base',
  'prism on elliptical base'               = 'prism on elliptical base',
  'prism on ellptical base'                = 'prism on elliptical base',
  'prism on eliiptical base'               = 'prism on elliptical base',
  'ellipsoid/prolate spheroid'             = 'ellipsoid/prolate spheroid',
  'cylinder + two cones'                   = 'cylinder + two cones',
  'cylinder with two cones'                = 'cylinder + two cones',
  '2 half ellipsoids'                      = 'two half ellipsoids',
  'two half ellipsoids'                    = 'two half ellipsoids',
  'wedge'                                  = 'wedge',
  'rectangular pyramid'                    = 'rectangular pyramid',
  'prolate spheriod'                       = 'prolate spheroid',
  'prolate spheroid'                       = 'prolate spheroid',
  'cone + half sphere'                     = 'cone + half sphere',
  'cone+ half sphere'                      = 'cone + half sphere',
  'cone with half sphere'                  = 'cone + half sphere',
  'cone with 1/2 sphere'                   = 'cone + half sphere',
  'cube'                                   = 'cube',
  'prism on triangle base'                 = 'prism on triangular base',
  'three cones'                            = 'three cones',
  'triangular pyramid'                     = 'triangular pyramid',
  'monoraphidioid'                         = 'other: monoraphidium type',
  'cone'                                   = 'cone'
)

df_phyto <- df_phyto %>%
  mutate(
    Shape = str_to_lower(str_trim(Shape)),
    Shape = recode(Shape, !!!shape_lookup, .default = Shape)
  )

unique(df_phyto$Shape)
```

Comments

TODO: determine what to do with 'delete' ones

```{r}
unique(df_phyto$Comments)
```

## Check Data

Check non-numeric values in numeric columns

```{r}
nonnum_vals <- df_phyto %>%
  select(`Organisms per mL`, `Unit Abundance`, `Cells per Unit`) %>%
  pivot_longer(
    everything(),
    names_to = 'column',
    values_to = 'raw_value'
  ) %>%
  mutate(
    raw_value = as.character(raw_value),
    cleaned_value = str_trim(raw_value),
    valid_numeric = str_detect(
      cleaned_value,
      '^[+-]?(\\d+\\.?\\d*|\\.\\d+)([eE][+-]?\\d+)?$'
    )
  ) %>%
  filter(
    !is.na(cleaned_value),
    cleaned_value != '',
    !valid_numeric
  ) %>%
  count(column, raw_value, sort = TRUE)

nonnum_vals
```

Fix the Cells per Unit col

```{r}
df_phyto <- df_phyto %>%
  mutate(
    standardized_note = case_when(
      `Cells per Unit` %in% c(
        '10 microns/unit',
        '10microns/unit',
        '10 micron unit'
      ) ~ '10 microns per unit',

      `Cells per Unit` == '.315 - 10 micron sections' ~
        '.315-10 microns per unit',

      `Cells per Unit` == '6 - 10micron units' ~
        '6-10 microns per unit',

      `Cells per Unit` == '9 -10 micron units' ~
        '9-10 microns per unit',

      TRUE ~ NA_character_
    ),

    Comments = case_when(
      is.na(standardized_note) ~ Comments,
      is.na(Comments) | trimws(Comments) == '' ~ standardized_note,
      TRUE ~ paste(Comments, standardized_note)
    ),

    `Cells per Unit` = if_else(
      !is.na(standardized_note),
      NA_character_,
      `Cells per Unit`
    )
  ) %>%
  select(-standardized_note)
```

```{r}
df_phyto %>%
  filter(is.na(`Cells per Unit`))
```


Looks good, convert.

```{r}
df_phyto <- df_phyto %>%
  mutate(`Organisms per mL` = as.numeric(`Organisms per mL`),
         `Unit Abundance` = as.numeric(`Unit Abundance`),
         `Cells per Unit` = as.numeric(`Cells per Unit`))
```

## Check Taxon

Correct spellings

```{r}
taxon_fixes <- tribble(
  ~Taxon,                        ~new_Taxon,                               ~new_Species,      ~new_Genus,
  'Diatoma hiemale',             'Diatoma hyemalis',                       'hyemalis',        NA,
  'Diatoma elongatum',           'Diatoma elongata',                       'elongata',        NA,
  'Chlorosphaeralean sp.',       'Chlorosphaera sp.',                      'sp.',             'Chlorosphaera',
  'Treubaria setigerum',         'Treubaria setigera',                     'setigera',        NA,
  'Attheya zachariasi',          'Attheya zachariasii',                    'zachariasii',     NA,
  'Bacillaria paxillifer',       'Bacillaria paxillifera',                 'paxillifera',     NA,
  'Microcystis flosaquae',       'Microcystis flos-aquae',                 'flos-aquae',      'Microcystis',
  'Diploneis didyma',            'Diploneis didymus',                      'didymus',         NA,
  'Chlorallantus sp.',           'Chlorallanthus sp.',                     'sp.',             'Chlorallanthus',
  'Centric Diatom sp.',          'Unidentified Centric Diatom sp. 1',      NA,                NA,
  'unknown centric sp.',         'Unidentified Centric Diatom sp. 1',      NA,                NA, 
  'unknown Centric diatom',      'Unidentified Centric Diatom sp. 1',      NA,                NA,
  'Pennate Diatom sp.',          'Unidentified Pennate Diatom sp. 1',      NA,                NA,
  'Unidentified flagellate',     'Unidentified Flagellate sp. 1',          NA,                NA,
  'chrysophyte',                 'Unidentified Chrysophyte sp. 1',         NA,                NA,
  'Gomphonema constricta',       'Gomphonema constrictum',                 'constrictum',     'Gomphonema',
  'Merismopedia warmigiana',     'Merismopedia warmingiana',               'warmingiana',     'Merismopedia',
  'Gonyaulax digitale',          'Gonyaulax digitalis',                    'digitalis',       'Gonyaulax',
  'Leptolyngbya foveolara',      'Leptolyngbya foveolarum',                'foveolarum',      'Leptolyngbya',
  'Nanoflagellates',             'Unidentified Nanoflagellate sp. 1',      NA,                NA,
  'crysophyte',                  'Unidentified Chrysophyte sp. 1',         NA,                NA,
  'unknown crysophyte',          'Unidentified Chrysophyte sp. 1',         NA,                NA,
  'unknown chrysophyte',         'Unidentified Chrysophyte sp. 1',         NA,                NA,
  'small nano-flagellates',      'Unidentified Nanoflagellate sp. 1',      NA,                NA,
  'small dinoflagellates',       'Unidentified Dinoflagellate sp. 1',      NA,                NA,
  'unknown girdle sp.',          'Unidentified sp. (girdle) 1',            NA,                NA,
  'unknown pennate girdle sp.',  'Unidentified Pennate Diatom sp. (girdle) 1',    NA,         NA,
  'Unknown Chlorophyte alga sp.','Unidentified Green Alga sp. 1',          NA,                NA,
  'unknown Chlorophyte filament','Unidentified Green Alga sp. 1',          NA,                NA,
  'Undetermined Euglenoids',     'Unidentified Euglenoid sp. 1',          NA,                NA,
  'Chodatella wratislaviensis',  'Chodatella wratislawiensis',             'wratislawiensis', 'Chodatella',
  'Eucampia zoodiacus',          'Eucampia zodiacus',                      'zodiacus',        'Eucampia',
  'Schroederia',                 'Schroederia sp.',                        'sp.',             'Schroederia',
  'NONE NONE',                   'No organisms found',                     NA,                NA,
  'Unidentified sp.',            'Unidentified sp. 1',                     NA,                NA,
  'Unknown unknown',             'Unidentified sp. 1',                     NA,                NA,
  'Pseudoanabaena sp.',          'Pseudanabaena sp.',                      'sp.',             'Pseudanabaena',
  'Tetraedron',                  'Tetraedron sp.',                         'sp.',             'Tetraedron',
  'Didymosphenia sp',            'Didymosphenia sp.',                      'sp.',             'Didymosphenia',
  'Chrysochromulina parva cf.',  'Chrysochromulina cf. parva',             'parva',           'Chrysochromulina',
  'Pseudonitzschia sp.',         'Pseudo-nitzschia sp.',                   'sp.',             'Pseudo-nitzschia',
  'Alexandrium sp. 2',           'Alexandrium sp.',                        'sp.',             'Alexandrium',
  'LGBs',                        'Little Green Balls',                     NA,                NA,
  'Tetraedron trigonum var. gracile', 'Tetraedron trigonum f. gracile', 'trigonum',         'Tetraedron',
  'Hemidiscus cuneiforme var. ventricosa', 'Hemidiscus cuneiformis var. ventricosus', 'cuneiformis var. ventricosus', 'Hemidiscus',
  'Scenedesmus quadricauda var. longispina fo. Granulatus', 'Scenedesmus quadricauda var. longispina f. granulatus', 'quadricauda var. longispina', 'Scenedesmus'
  
)

df_phyto <- df_phyto %>%
  left_join(taxon_fixes, by = 'Taxon') %>%
  mutate(
    Species = coalesce(new_Species, Species),
    Genus   = coalesce(new_Genus, Genus),
    Taxon   = coalesce(new_Taxon, Taxon)
  ) %>%
  select(-new_Taxon, -new_Species, -new_Genus)
```

Compare to TIN code list

```{r}
df_eataxa <- anti_join(df_phyto, df_codes, by = 'Taxon') %>% distinct(Taxon)

df_eataxa
```

Compare ones not in TIN codes to taxonomy list

```{r}
df_newinfo <- inner_join(df_taxon, df_eataxa, by = 'Taxon') %>%
  select(Taxon, CurrentTaxon)
```

See what's left

```{r}
df_missing <- anti_join(df_eataxa, df_taxon, by = 'Taxon')

df_missing
```

## Fixes

Filter out missing data (broken vial)

```{r}
df_phyto <- df_phyto %>%
  filter(!is.na(Taxon))
```

Combine the Melosira sp. 2

```{r}
df_phyto %>%
  filter(
    Date == as.Date('2009-04-16'),
    Station == 'P8',
    str_detect(Taxon, regex('Melosira', ignore_case = TRUE))
  )
```

```{r}
rows_to_combine <- df_phyto %>%
  mutate(.row_id = row_number()) %>%
  filter(
    Date == as.Date('2009-04-16'),
    Station == 'P8',
    Taxon %in% c('Melosira sp.', 'Melosira sp. 2')
  )

# Use the Melosira sp. row as the template, including its EcoA_TIN.
combined_row <- rows_to_combine %>%
  filter(Taxon == 'Melosira sp.') %>%
  slice(1) %>%
  mutate(
    `Organisms per mL` = sum(
      rows_to_combine$`Organisms per mL`,
      na.rm = TRUE
    ),
    `Unit Abundance` = sum(
      rows_to_combine$`Unit Abundance`,
      na.rm = TRUE
    )
  ) %>%
  select(-.row_id)

# Remove the two original rows and insert the combined row.
df_phyto <- df_phyto %>%
  mutate(.row_id = row_number()) %>%
  filter(!.row_id %in% rows_to_combine$.row_id) %>%
  select(-.row_id) %>%
  bind_rows(combined_row)
```


## Check stations

Stations that have phyto data but no corresponding DWQ data

TODO: get some sort of info on these

```{r}
setdiff(df_phyto$Station, df_wq$Station)
```

```{r}
missing_stations <- setdiff(df_phyto$Station, df_wq$Station)

df_phyto %>%
  filter(Station %in% missing_stations) %>%
  mutate(Year = format(Date, '%Y'), Month = format(Date, '%m'), Day = format(Date, '%d')) %>%
  distinct(Station, Year, Month, Day) %>%
  arrange(Station, Year, Month, Day)
```

assign the TB data to SPECIAL MethodCode

TODO: is this right?

```{r}
df_phyto <- df_phyto %>%
  mutate(
    MethodCode = if_else(
      Station == 'TB',
      'SPECIAL',
      'PHYTO'
    )
  )
```


## Check Dates

Compare against DWQ

```{r}
missing_stations <- setdiff(
  df_phyto$Station,
  df_wq$Station
)

wq_events <- df_wq %>%
  mutate(
    Year = year(Date),
    Month = month(Date)
  ) %>%
  distinct(Year, Month, Station, Date) %>%
  group_by(Year, Month, Station) %>%
  summarise(
    n_events_wq = n(),
    wq_dates = paste(sort(Date), collapse = ', '),
    .groups = 'drop'
  )

phyto_events <- df_phyto %>%
  filter(!Station %in% missing_stations) %>%
  mutate(
    Year = year(Date),
    Month = month(Date)
  ) %>%
  distinct(Year, Month, Station, Date) %>%
  group_by(Year, Month, Station) %>%
  summarise(
    n_events_phyto = n(),
    phyto_dates = paste(sort(Date), collapse = ', '),
    .groups = 'drop'
  )

comparison <- wq_events %>%
  full_join(
    phyto_events,
    by = c('Year', 'Month', 'Station')
  ) %>%
  mutate(
    status = case_when(
      is.na(n_events_wq)            ~ 'phyto only',
      is.na(n_events_phyto)         ~ 'wq only (missing in phyto)',
      n_events_wq != n_events_phyto ~ 'mismatch in # events',
      TRUE                          ~ 'match'
    )
  ) %>%
  arrange(status, Year, Month, Station)

problems <- comparison %>%
  filter(status != 'match') %>%
  arrange(Year, Month, Station)

problems
```

```{r}
problem_station_summary <- problems %>%
  count(Station, name = 'n_times', sort = TRUE)

problem_station_summary
```

Check for multiple events per month/year

```{r}
multiple_phyto_events <- phyto_events %>%
  filter(n_events_phyto > 1) %>%
  arrange(Station, Year, Month)

multiple_phyto_events
```


Check and update mismatched dates (one event per Month/Year/Station in each)
- update to DWQ dates

```{r}
wq_single <- df_wq %>%
  mutate(Year = year(Date), Month = month(Date)) %>%
  distinct(Year, Month, Station, Date) %>%
  group_by(Year, Month, Station) %>%
  summarise(
    n_wq = n(),
    date_wq = paste(sort(Date), collapse = ', '),
    .groups = 'drop'
  ) %>%
  filter(n_wq == 1)

routine_single <- df_phyto %>%
  mutate(Year = year(Date), Month = month(Date)) %>%
  distinct(Year, Month, Station, Date) %>%
  group_by(Year, Month, Station) %>%
  summarise(
    n_routine = n(),
    date_routine = paste(sort(Date), collapse = ', '),
    .groups = 'drop'
  ) %>%
  filter(n_routine == 1)

single_event_mismatches <- wq_single %>%
  inner_join(routine_single, by = c('Year', 'Month', 'Station')) %>%
  filter(date_wq != date_routine) %>%
  arrange(Year, Month, Station)

single_event_mismatches
```


```{r}
df_phyto <- df_phyto %>%
  mutate(Year = year(Date), Month = month(Date)) %>%
  left_join(
    single_event_mismatches %>%
      transmute(Year, Month, Station, new_date = as.Date(date_wq)),
    by = c('Year', 'Month', 'Station')
  ) %>%
  mutate(Date = if_else(!is.na(new_date), new_date, Date)) %>%
  select(-Year, -Month, -new_date)
```

## Check Times

```{r}
wq_times <- df_wq %>%
  distinct(Station, Date, Time) %>%
  mutate(Time = sub(':00$', '', Time)) %>%
  rename(Time_wq = Time)

routine_times <- df_phyto %>%
  distinct(Station, Date, Time) %>%
  rename(Time_routine = Time)

time_mismatches <- wq_times %>%
  inner_join(routine_times, by = c('Station', 'Date')) %>%
  filter(Time_wq != Time_routine | (is.na(Time_wq) != is.na(Time_routine))) %>%
  arrange(Station, Date)

time_mismatches
```

Check which Dates have > Time
```{r}
routine_times %>% count(Station, Date) %>% filter(n > 1)
```

Update Times
```{r}
df_phyto <- df_phyto %>%
  left_join(
    time_mismatches %>%
      select(Station, Date, Time_routine, Time_wq),
    by = c('Station', 'Date', 'Time' = 'Time_routine')
  ) %>%
  mutate(Time = if_else(!is.na(Time_wq), Time_wq, Time)) %>%
  select(-Time_wq)
```

## Check missing data

```{r}
# build the full Month x Year grid, then flag presence/absence
routine_coverage <- df_phyto %>%
  mutate(Year = year(Date), Month = month(Date)) %>%
  distinct(Year, Month) %>%
  mutate(has_data = TRUE) %>%
  right_join(
    expand_grid(Year = 2008:2013, Month = 1:12),
    by = c('Year', 'Month')
  ) %>%
  mutate(has_data = if_else(is.na(has_data), FALSE, has_data))

ggplot(routine_coverage, aes(x = factor(Month), y = factor(Year), fill = has_data)) +
  geom_tile(color = 'white', linewidth = 0.2) +
  scale_fill_manual(values = c('TRUE' = 'steelblue', 'FALSE' = 'grey90'),
                     labels = c('TRUE' = 'Data', 'FALSE' = 'No data'),
                     name = NULL) +
  scale_x_discrete(labels = month.abb) +
  scale_y_discrete(limits = rev(as.character(2008:2013))) +
  labs(x = 'Month', y = 'Year', title = 'df_phyto data coverage, 2008–2013') +
  theme_minimal() +
  theme(panel.grid = element_blank())
```

# Format Data
TODO: add lab column, format for DB (here, Ecoanalysts)

# Save data

Write data

```{r}
save(df_phyto, file = here::here('03_Phyto/aquarius/historical-imports/data/2008-2013.rda'))
```



