# Clean and Format 2013-2021 Phyto Data

Special Studies aren't already marked. Sept/Nov 2013 data wasn't in previous batch.

# Initial setup

```{r}
library(pespr)
library(lubridate)
library(tidyverse)
source(here::here('00_GlobalFunctions/functions.R'))
```


## Read in data

```{r}
# phyto data
fp_raw <- abs_path_emp('Phytoplankton/AQUARIUS Samples Database/Database Migration/Raw Phyto Data/EMP_Phyto Data 13-21.csv')

df_phyto <- read_csv(fp_raw, col_types = cols(.default = col_character())) %>%
  rename(Date = SampleDate,
         Time = SampleTime,
         Station = StationCode,
         `Field-of-view (mm2)` = `Field-of-view (mm²)`,
         `Slide Chamber Area (mm2)` = `Slide/ Chamber Area (mm²)`,
          `Diatom or SoftBody` = `Diatom/SoftBody`)


# tin data
load(here::here('03_Phyto/aquarius/historical-imports/data/tins.rda'))

# dwq data
load(here::here('03_Phyto/aquarius/historical-imports/data/dwq_data.rda'))

# phyto taxa data
df_taxon <- pespr::phyto_taxonomy

unique(df_phyto$Factor)
```

# Dates

Convert Dates
```{r}
df_phyto <- df_phyto %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y'))

df_phyto <- df_phyto %>%
  mutate(Time = str_replace(Time, '^0(\\d:)', '\\1'))

df_wq <- df_wq %>%
  mutate(Date = as.Date(Date, format = '%Y-%m-%d'))

start_date <- as.Date('2013-09-01')
end_date   <- as.Date('2021-12-30')

df_wq <- df_wq %>%
  filter(between(Date, start_date, end_date))
```

```{r}
unique(df_phyto$Time)
```
# Times

Fix non-military

```{r}
df_phyto <- df_phyto %>%
  mutate(Time = ifelse(
    !is.na(Time) & as.numeric(sub(':.*', '', Time)) < 4,
    sprintf('%d:%s',
            as.numeric(sub(':.*', '', Time)) + 12,
            sub('^[^:]*:', '', Time)),
    Time
  ))
```

Round Times

```{r}
round_time <- function(x) {
  out <- rep(NA_character_, length(x))
  ok <- !is.na(x)

  rounded <- round_date(
    as_datetime(paste0('2000-01-01 ', x[ok], ':00')),
    '5 minutes'
  )

  out[ok] <- sub('^0', '', format(rounded, '%H:%M'))
  out
}

df_phyto <- df_phyto %>% mutate(Time = round_time(Time))

unique(df_phyto$Time)
```

# Stations

```{r}
unique(df_phyto$Station)
```


Rename EZs
```{r}
df_phyto <- df_phyto %>%
  mutate(
    Station = recode(
      Station,
      "EZ2"           = "LSZ2",
      "E22"           = "LSZ2",
      "E26"           = "LSZ2",
      "EZ6"           = "LSZ6",
      "D16-Twitchell" = "D16",
      "D16 Twitchell" = "D16",
      "C3A-Hood"      = "C3A",
      "EZ2 SJR"       = "LSZ2-SJR",
      "EZ2-SJR"       = "LSZ2-SJR",
      "EZ6-SJR"       = "LSZ6-SJR",
      "E22"           = "D22",
      "NZ328"         = "NZ325"
    )
  )
```


Determine stations that don't match with DWQ

```{r}
setdiff(df_phyto$Station, df_wq$Station)
```


# Data Checks

## Numeric Metadata

```{r}
unique(df_phyto$`Volume Received (mL)`)
```

```{r}
unique(df_phyto$`Volume Analyzed (mL)`)
```

```{r}
# remove percent sign
df_phyto <- df_phyto %>%
  mutate(`Percent of Sample Counted` = str_remove(`Percent of Sample Counted`, '%'))

unique(df_phyto$`Percent of Sample Counted`)
```

```{r}
unique(df_phyto$`Field-of-view (mm2)`)
```

```{r}
unique(df_phyto$`Slide Chamber Area (mm2)`)
```

```{r}
df_phyto <- df_phyto %>%
  mutate(
    `Number of Fields Counted` = na_if(
      as.character(`Number of Fields Counted`),
      "N/A"
    )
  )

unique(df_phyto$`Number of Fields Counted`)
```
```{r}
unique(df_phyto$Factor)
```
```{r}
unique(df_phyto$GALD)
```

Looks good, convert to numeric.

```{r}
df_phyto <- df_phyto %>%
  mutate(
    across(
      all_of(c(
        "Volume Received (mL)",
        "Volume Analyzed (mL)",
        "Percent of Sample Counted",
        "Field-of-view (mm2)",
        "Slide Chamber Area (mm2)",
        "Area Counted",
        "Number of Fields Counted",
        "Total Cells Counted",
        "Organisms per mL",
        "GALD",
        "Factor",
        paste0("Biovolume ", 1:10)
      )),
      as.numeric
    )
  )
```

## Check Calcs

Area Counted (looks good, rounding errors)

```{r}
df_test <- df_phyto %>%
  mutate(Area2 = (`Field-of-view (mm2)`*`Number of Fields Counted`))

df_test <- df_test %>%
  mutate(
    AreaCounted = round(`Area Counted`,2),
    Area_same = round(AreaCounted,2) == round(Area2, 2)
  )

df_test %>%
  summarise(
    all_same = all(Area_same, na.rm = TRUE),
    n_different = sum(!Area_same, na.rm = TRUE)
  )

df_test %>%
  filter(!Area_same) %>%
  select(AreaCounted, Area2)
```

Factor
```{r}
df_test <- df_phyto %>%
  mutate(Factor2 = (`Slide Chamber Area (mm2)`)/(`Area Counted`*`Volume Analyzed (mL)`))

df_test <- df_test %>%
  mutate(
    Factor = as.numeric(Factor),
    Factor_same = Factor == round(Factor2, 2)
  )

df_test %>%
  summarise(
    all_same = all(Factor_same, na.rm = TRUE),
    n_different = sum(!Factor_same, na.rm = TRUE)
  )
```

Organisms per mL (looks good, differences are minor)

```{r}
df_test <- df_test %>%
  mutate(UnitpermL = (Factor)*(`Unit Abundance`))

df_test <- df_test %>%
  mutate(
    Org_same = `Organisms per mL` == round(UnitpermL, 2)
  )

df_test %>%
  summarise(
    all_same = all(Org_same, na.rm = TRUE),
    n_different = sum(!Org_same, na.rm = TRUE)
  )

df_test %>%
  filter(!Org_same) %>%
  select(`Organisms per mL`, UnitpermL)
```

## Create Cells/Biovol cols

Create Cells per mL
```{r}
df_phyto <- df_phyto %>%
  mutate(`Cells per mL` = (Factor)*(`Total Cells Counted`))
```

## Check Cells

Cells per mL < Units per mL check
Very minor differences are due to rounding. When this occurs, set Cells per mL to Units per mL (within 1)
```{r}
df_phyto %>%
  filter(
    !near(
      `Cells per mL`,
      `Organisms per mL`
    ) &
    `Cells per mL` < `Organisms per mL`
  ) %>%
  select(`Cells per mL`, `Organisms per mL`)
```

```{r}
df_phyto <- df_phyto %>%
  mutate(
    `Cells per mL` = if_else(
      abs(`Cells per mL` - `Organisms per mL`) <= 1,
      as.numeric(`Organisms per mL`),
      `Cells per mL`
    )
  )

df_phyto %>%
  filter(
    !near(
      as.numeric(`Cells per mL`),
      as.numeric(`Organisms per mL`)
    ) &
    as.numeric(`Cells per mL`) < as.numeric(`Organisms per mL`)
  ) %>%
  select(`Cells per mL`, `Organisms per mL`)
```

## Create Biovol per mL

```{r}
df_phyto <- df_phyto %>%
  mutate(
    `Average Biovolume per mL` = round(rowMeans(pick(all_of(paste0("Biovolume ", 1:10))), na.rm = TRUE) * `Cells per mL`, 2),
    `Average Biovolume per mL` = replace(`Average Biovolume per mL`, is.nan(`Average Biovolume per mL`), NA)
  )
```


## String Metadata

Check Diatom/Softbody

```{r}
unique(df_phyto$`Diatom or SoftBody`)
```

Filament Group Codes

```{r}
unique(df_phyto$`Colony or Filament Group Code`)
```

Shape

```{r}
unique(df_phyto$Shape)
```

```{r}
shape_lookup <- c(
  'cuboid'                                 = 'cuboid',
  'ellipsoid'                              = 'ellipsoid',
  'other: achnanthes type'                 = 'other: achnanthes type',
  'other: achnanthes'                      = 'other: achnanthes type',
  'cylinder'                               = 'cylinder',
  'sphere'                                 = 'sphere',
  'ellipsoid + cone'                       = 'ellipsoid + cone',
  'ellipsoid+cone'                         = 'ellipsoid + cone',
  'ellipsoid with cone'                    = 'ellipsoid + cone',
  'ellipsoid + 2 cones + cylinder'         = 'ellipsoid + two cones + cylinder',
  'ellipsoid + two cones + cylinder'       = 'ellipsoid + two cones + cylinder',
  'peridinium'                             = 'other: peridinium type',
  'other: peridinium type'                 = 'other: peridinium type',
  'two cones'                              = 'two cones',
  'other: crucigenia tetrapedia type'      = 'other: crucigenia tetrapedia type',
  'box'                                    = 'box',
  'ovoid'                                  = 'ovoid',
  'two wedges'                             = 'two wedges',
  'half-elliptic prism'                    = 'half prism on elliptical base',
  'prism on elliptic base'                 = 'prism on elliptical base',
  'prism on elliptical base'               = 'prism on elliptical base',
  'prism on ellptical base'                = 'prism on elliptical base',
  'prism on eliiptical base'               = 'prism on elliptical base',
  'elliptic prism'                         = 'prism on elliptical base',
  'elliptic prism + cylinder'              = 'prism on elliptical base + cylinder',
  'prism on parallelogram'                 = 'prism on parallelogram base',
  'ellipsoid/prolate spheroid'             = 'ellipsoid/prolate spheroid',
  'cylinder + two cones'                   = 'cylinder + two cones',
  'cylinder with two cones'                = 'cylinder + two cones',
  'cylinder + 2 cones'                     = 'cylinder + two cones',
  'cylinder + cone'                        = 'cylinder + cone',
  'cylinder + 2 half spheres'              = 'cylinder + two half spheres',
  'box + 2 cylinders'                      = 'box + two cylinders',
  'prolate spheroid + 2 cylinders'         = 'prolate spheroid + two cylinders',
  '2 half ellipsoids'                      = 'two half ellipsoids',
  'two half ellipsoids'                    = 'two half ellipsoids',
  'half ellipsoid + cone on an elliptic base' = 'half ellipsoid + cone on an elliptical base',
  'sickle-shaped prism (elliptic prism)'   = 'sickle-shaped prism (prism on elliptical base)',
  'wedge'                                  = 'wedge',
  'rectangular pyramid'                    = 'rectangular pyramid',
  'prolate spheriod'                       = 'prolate spheroid',
  'prolate spheroid'                       = 'prolate spheroid',
  'cone + half sphere'                     = 'cone + half sphere',
  'cone+ half sphere'                      = 'cone + half sphere',
  'cone with half sphere'                  = 'cone + half sphere',
  'cone with 1/2 sphere'                   = 'cone + half sphere',
  'cube'                                   = 'cube',
  'prism on triangle base'                 = 'prism on triangular base',
  'prism on triangle'                      = 'prism on triangular base',
  'three cones'                            = 'three cones',
  'triangular pyramid'                     = 'triangular pyramid',
  'monoraphidioid'                         = 'other: monoraphidium type',
  'gomphonemoid'                           = 'other: gomphonema type',
  'cone'                                   = 'cone',
  '2 cones'                                = 'two cones',
  '3 cones'                                = 'three cones'
)

df_phyto <- df_phyto %>%
  mutate(
    Shape = str_to_lower(str_trim(Shape)),
    Shape = recode(Shape, !!!shape_lookup, .default = Shape)
  )

unique(df_phyto$Shape)
```

Comments

Fix obvious typos

```{r}
df_phyto$Comments <- stringr::str_replace_all(
  df_phyto$Comments,
  c(
    "detitus" = "detritus",
    "detrtitus" = "detritus",
    "detrtius" = "detritus",
    "sedimend" = "sediment",
    "sedimnet" = "sediment",
    "disintigrated" = "disintegrated",
    "mucilagenous" = "mucilaginous",
    "Ditylum brightwellei" = "Ditylum brightwellii",
    "Degraded\\.Moderate" = "Degraded. Moderate",
    "detritus\\.Low" = "detritus. Low",
    "detritus\\.Cannot" = "detritus. Cannot",
    "1mL" = "1 mL"
  )
)
```


## Check Taxon

Correct spellings

```{r}
taxon_fixes <- tribble(
  ~Taxon,                        ~new_Taxon,                               ~new_Species,      ~new_Genus,
  'Diatoma hiemale',             'Diatoma hyemalis',                       'hyemalis',        NA,
  'Diatoma elongatum',           'Diatoma elongata',                       'elongata',        NA,
  'Chlorosphaeralean sp.',       'Chlorosphaera sp.',                      'sp.',             'Chlorosphaera',
  'Treubaria setigerum',         'Treubaria setigera',                     'setigera',        NA,
  'Attheya zachariasi',          'Attheya zachariasii',                    'zachariasii',     NA,
  'Bacillaria paxillifer',       'Bacillaria paxillifera',                 'paxillifera',     NA,
  'Microcystis flosaquae',       'Microcystis flos-aquae',                 'flos-aquae',      'Microcystis',
  'Diploneis didyma',            'Diploneis didymus',                      'didymus',         NA,
  'Chlorallantus sp.',           'Chlorallanthus sp.',                     'sp.',             'Chlorallanthus',
  'Centric Diatom sp.',          'Unidentified Centric Diatom sp. 1',      NA,                NA,
  'unknown centric sp.',         'Unidentified Centric Diatom sp. 1',      NA,                NA, 
  'unknown Centric diatom',      'Unidentified Centric Diatom sp. 1',      NA,                NA,
  'Pennate Diatom sp.',          'Unidentified Pennate Diatom sp. 1',      NA,                NA,
  'Unidentified flagellate',     'Unidentified Flagellate sp. 1',          NA,                NA,
  'chrysophyte',                 'Unidentified Chrysophyte sp. 1',         NA,                NA,
  'Gomphonema constricta',       'Gomphonema constrictum',                 'constrictum',     'Gomphonema',
  'Merismopedia warmigiana',     'Merismopedia warmingiana',               'warmingiana',     'Merismopedia',
  'Gonyaulax digitale',          'Gonyaulax digitalis',                    'digitalis',       'Gonyaulax',
  'Leptolyngbya foveolara',      'Leptolyngbya foveolarum',                'foveolarum',      'Leptolyngbya',
  'Nanoflagellates',             'Unidentified Nanoflagellate sp. 1',      NA,                NA,
  'crysophyte',                  'Unidentified Chrysophyte sp. 1',         NA,                NA,
  'unknown crysophyte',          'Unidentified Chrysophyte sp. 1',         NA,                NA,
  'unknown chrysophyte',         'Unidentified Chrysophyte sp. 1',         NA,                NA,
  'small nano-flagellates',      'Unidentified Nanoflagellate sp. 1',      NA,                NA,
  'small dinoflagellates',       'Unidentified Dinoflagellate sp. 1',      NA,                NA,
  'Unidentified Dinoflagellate sp.','Unidentified Dinoflagellate sp. 1',   NA,                NA,
  'Unidentified dinoflagellate sp.','Unidentified Dinoflagellate sp. 1',   NA,                NA,
  'unknown girdle sp.',          'Unidentified sp. (girdle) 1',            NA,                NA,
  'Unidentified Cyanobacteria sp.','Unidentified Cyanobacterium sp. 1',    NA,                NA,
  'unknown pennate girdle sp.',  'Unidentified Pennate Diatom sp. (girdle) 1',    NA,         NA,
  'Unknown Chlorophyte alga sp.','Unidentified Green Alga sp. 1',          NA,                NA,
  'unknown Chlorophyte filament','Unidentified Green Alga sp. 1',          NA,                NA,
  'Undetermined Euglenoids',     'Unidentified Euglenoid sp. 1',          NA,                NA,
  'Chodatella wratislaviensis',  'Chodatella wratislawiensis',             'wratislawiensis', 'Chodatella',
  'Eucampia zoodiacus',          'Eucampia zodiacus',                      'zodiacus',        'Eucampia',
  'Schroederia',                 'Schroederia sp.',                        'sp.',             'Schroederia',
  'NONE NONE',                   'No organisms found',                     NA,                NA,
  'Unidentified sp.',            'Unidentified sp. 1',                     NA,                NA,
  'Unknown unknown',             'Unidentified sp. 1',                     NA,                NA,
  'Pseudoanabaena sp.',          'Pseudanabaena sp.',                      'sp.',             'Pseudanabaena',
  'Tetraedron',                  'Tetraedron sp.',                         'sp.',             'Tetraedron',
  'Didymosphenia sp',            'Didymosphenia sp.',                      'sp.',             'Didymosphenia',
  'Chrysochromulina parva cf.',  'Chrysochromulina cf. parva',             'parva',           'Chrysochromulina',
  'Pseudonitzschia sp.',         'Pseudo-nitzschia sp.',                   'sp.',             'Pseudo-nitzschia',
  'Alexandrium sp. 2',           'Alexandrium sp.',                        'sp.',             'Alexandrium',
  'LGBs',                        'Little Green Balls',                     NA,                NA,
  'Tetraedron trigonum var. gracile', 'Tetraedron trigonum f. gracile', 'trigonum',         'Tetraedron',
  'Hemidiscus cuneiforme var. ventricosa', 'Hemidiscus cuneiformis var. ventricosus', 'cuneiformis var. ventricosus', 'Hemidiscus',
  'Scenedesmus quadricauda var. longispina fo. Granulatus', 'Scenedesmus quadricauda var. longispina f. granulatus', 'quadricauda var. longispina', 'Scenedesmus'
  
)

df_phyto <- df_phyto %>%
  left_join(taxon_fixes, by = 'Taxon') %>%
  mutate(
    Species = coalesce(new_Species, Species),
    Genus   = coalesce(new_Genus, Genus),
    Taxon   = coalesce(new_Taxon, Taxon)
  ) %>%
  select(-new_Taxon, -new_Species, -new_Genus)
```

Compare to TIN code list (cf normalized)

```{r}
taxon_fixes <- tribble(
  ~Taxon,                        ~new_Taxon,                               ~new_Species,      ~new_Genus,
  'Bacillaria paxillifer',       'Bacillaria paxillifera',                 'paxillifera',     NA,
  'cf. Bacillaria paxillifer',   'cf. Bacillaria paxillifera',             'paxillifera',     NA,
  'Bacillaria cf. paxillifer',   'Bacillaria cf. paxillifera',             'paxillifera',     NA,
  'Gonyaulax cf. digitale',      'Gonyaulax cf. digitalis',                'digitalis',       NA,
  'Aulacoseira cf. granulata fragment',   'Aulacoseira cf. granulata',     'granulata',       'Aulacoseira',
  'cf. Gymnodinium',             'cf. Gymnodinium sp.',                    'sp.',             'Gymnodinium'
  
)

df_phyto <- df_phyto %>%
  left_join(taxon_fixes, by = 'Taxon') %>%
  mutate(
    Species = coalesce(new_Species, Species),
    Genus   = coalesce(new_Genus, Genus),
    Taxon   = coalesce(new_Taxon, Taxon)
  ) %>%
  select(-new_Taxon, -new_Species, -new_Genus)
```


```{r}
df_eataxa <- df_phyto %>%
  mutate(
    Taxon_match = Taxon %>%
      str_remove_all(regex("\\bcf\\.", ignore_case = TRUE)) %>%
      str_squish()
  ) %>%
  anti_join(
    df_codes %>%
      mutate(
        Taxon_match = Taxon %>%
          str_remove_all(regex("\\bcf\\.", ignore_case = TRUE)) %>%
          str_squish()
      ) %>%
      distinct(Taxon_match),
    by = "Taxon_match"
  ) %>%
  distinct(Taxon, BSA_TIN) %>%
  arrange(Taxon)

df_eataxa
```

Compare ones not in TIN codes to taxonomy list

```{r}
df_newinfo <- inner_join(df_taxon, df_eataxa, by = 'Taxon') %>%
  select(Taxon, CurrentTaxon)
```

See what's left

```{r}
df_missing <- anti_join(df_eataxa, df_taxon, by = 'Taxon')

df_missing
```

## Fixes

Fix missing Taxon

```{r}
df_phyto <- df_phyto %>%
  mutate(
    Taxon = if_else(
      is.na(Taxon),
      "Entomoneis sp.",
      Taxon
    )
  )
```

Remove fragment

```{r}
df_phyto <- df_phyto %>%
  mutate(
    Taxon = str_replace(
      Taxon,
      "^Aulacoseira cf\\. granulata fragment$",
      "Aulacoseira cf. granulata"
    )
  )
```


## Check stations

Stations that have phyto data but no corresponding DWQ data

TODO: get some sort of info on these

```{r}
setdiff(df_phyto$Station, df_wq$Station)
```

```{r}
missing_stations <- setdiff(df_phyto$Station, df_wq$Station)

df_phyto %>%
  filter(Station %in% missing_stations) %>%
  mutate(Year = format(Date, '%Y'), Month = format(Date, '%m'), Day = format(Date, '%d')) %>%
  distinct(Station, Year, Month, Day) %>%
  arrange(Station, Year, Month, Day)
```

assign the TB data to SPECIAL MethodCode

TODO: is this right?

```{r}
df_phyto <- df_phyto %>%
  mutate(
    MethodCode = if_else(
      Station == 'TB',
      'SPECIAL',
      'PHYTO'
    )
  )
```


## Check Dates

Compare against DWQ

```{r}
missing_stations <- setdiff(
  df_phyto$Station,
  df_wq$Station
)

wq_events <- df_wq %>%
  mutate(
    Year = year(Date),
    Month = month(Date)
  ) %>%
  distinct(Year, Month, Station, Date) %>%
  group_by(Year, Month, Station) %>%
  summarise(
    n_events_wq = n(),
    wq_dates = paste(sort(Date), collapse = ', '),
    .groups = 'drop'
  )

phyto_events <- df_phyto %>%
  filter(!Station %in% missing_stations) %>%
  mutate(
    Year = year(Date),
    Month = month(Date)
  ) %>%
  distinct(Year, Month, Station, Date) %>%
  group_by(Year, Month, Station) %>%
  summarise(
    n_events_phyto = n(),
    phyto_dates = paste(sort(Date), collapse = ', '),
    .groups = 'drop'
  )

comparison <- wq_events %>%
  full_join(
    phyto_events,
    by = c('Year', 'Month', 'Station')
  ) %>%
  mutate(
    status = case_when(
      is.na(n_events_wq)            ~ 'phyto only',
      is.na(n_events_phyto)         ~ 'wq only (missing in phyto)',
      n_events_wq != n_events_phyto ~ 'mismatch in # events',
      TRUE                          ~ 'match'
    )
  ) %>%
  arrange(status, Year, Month, Station)

problems <- comparison %>%
  filter(!(status %in% c('match', 'wq only (missing in phyto)'))) %>%
  arrange(Year, Month, Station)

problems
```

Check for multiple events per month/year

```{r}
multiple_phyto_events <- phyto_events %>%
  filter(n_events_phyto > 1) %>%
  arrange(Station, Year, Month)

multiple_phyto_events
```

Fixes

```{r}
df_phyto <- df_phyto %>%
  mutate(
    Date = if_else(
      Station == 'D28A' & Date == as.Date('2016-05-10'),
      as.Date('2016-05-11'),
      Date
    ),
    Time = if_else(
      Station == 'D28A' & Date == as.Date('2016-05-11'),
      '9:25',
      Time
    )
  )


df_phyto <- df_phyto %>%
  mutate(
    Date = if_else(
      Station == 'D26' & Date == as.Date('2020-09-08'),
      as.Date('2020-09-09'),
      Date
    )
  )

df_phyto <- df_phyto %>%
  mutate(
    Date = if_else(
      Station == 'D19' & Date == as.Date('2021-11-19'),
      as.Date('2021-11-09'),
      Date
    )
  )
```

Check and update mismatched dates (one event per Month/Year/Station in each)
- update to DWQ dates

```{r}
wq_single <- df_wq %>%
  mutate(Year = year(Date), Month = month(Date)) %>%
  distinct(Year, Month, Station, Date) %>%
  group_by(Year, Month, Station) %>%
  summarise(
    n_wq = n(),
    date_wq = paste(sort(Date), collapse = ', '),
    .groups = 'drop'
  ) %>%
  filter(n_wq == 1)

routine_single <- df_phyto %>%
  mutate(Year = year(Date), Month = month(Date)) %>%
  distinct(Year, Month, Station, Date) %>%
  group_by(Year, Month, Station) %>%
  summarise(
    n_routine = n(),
    date_routine = paste(sort(Date), collapse = ', '),
    .groups = 'drop'
  ) %>%
  filter(n_routine == 1)

single_event_mismatches <- wq_single %>%
  inner_join(routine_single, by = c('Year', 'Month', 'Station')) %>%
  filter(date_wq != date_routine) %>%
  arrange(Year, Month, Station)

single_event_mismatches
```

```{r}
df_phyto <- df_phyto %>%
  mutate(Year = year(Date), Month = month(Date)) %>%
  left_join(
    single_event_mismatches %>%
      transmute(Year, Month, Station, new_date = as.Date(date_wq)),
    by = c('Year', 'Month', 'Station')
  ) %>%
  mutate(Date = if_else(!is.na(new_date), new_date, Date)) %>%
  select(-Year, -Month, -new_date)
```

## Check Times

```{r}
wq_times <- df_wq %>%
  distinct(Station, Date, Time) %>%
  mutate(Time = sub(':00$', '', Time)) %>%
  rename(Time_wq = Time)

routine_times <- df_phyto %>%
  distinct(Station, Date, Time) %>%
  rename(Time_routine = Time)

time_mismatches <- wq_times %>%
  inner_join(routine_times, by = c('Station', 'Date')) %>%
  filter(Time_wq != Time_routine | (is.na(Time_wq) != is.na(Time_routine))) %>%
  arrange(Station, Date)

time_mismatches
```

Check which Dates have > Time
```{r}
routine_times %>% count(Station, Date) %>% filter(n > 1)
```

Update Times
```{r}
df_phyto <- df_phyto %>%
  left_join(
    time_mismatches %>%
      select(Station, Date, Time_routine, Time_wq),
    by = c('Station', 'Date', 'Time' = 'Time_routine')
  ) %>%
  mutate(Time = if_else(!is.na(Time_wq), Time_wq, Time)) %>%
  select(-Time_wq)
```

## Check missing data

```{r}
# build the full Month x Year grid, then flag presence/absence
routine_coverage <- df_phyto %>%
  mutate(Year = year(Date), Month = month(Date)) %>%
  distinct(Year, Month) %>%
  mutate(has_data = TRUE) %>%
  right_join(
    expand_grid(Year = 2013:2021, Month = 1:12),
    by = c('Year', 'Month')
  ) %>%
  mutate(has_data = if_else(is.na(has_data), FALSE, has_data))

ggplot(routine_coverage, aes(x = factor(Month), y = factor(Year), fill = has_data)) +
  geom_tile(color = 'white', linewidth = 0.2) +
  scale_fill_manual(values = c('TRUE' = 'steelblue', 'FALSE' = 'grey90'),
                     labels = c('TRUE' = 'Data', 'FALSE' = 'No data'),
                     name = NULL) +
  scale_x_discrete(labels = month.abb) +
  scale_y_discrete(limits = rev(as.character(2013:2021))) +
  labs(x = 'Month', y = 'Year', title = 'data coverage, 2013–2021') +
  theme_minimal() +
  theme(panel.grid = element_blank())
```

# Format Data
TODO: add lab column, format for DB (here, Ecoanalysts)

# Save data

Write data

```{r}
save(df_phyto, file = here::here('03_Phyto/aquarius/historical-imports/data/2013-2021.rda'))
```



