Title: Data Exploration
Purpose: determine issues with current EDI dataset and how it relates to past data
Author: Sarah Perry
Date: 07/2022

# Setup
- cur data is current EDI (2021)
- old data is old EDI (2020)
- lab data is from old website
```{r}
# import packages
library(tidyverse)
library(lubridate)
'%!in%' <- function(x,y)!('%in%'(x,y))

# read in data
df_old <- read_csv('data/data_old.csv') # 2020 EDI data (v4)
df_final <- read_csv('data/edi_2010-2021.csv') # 2021 EDI data (with errors, v6)
df_lab <- read_csv('data/data_lab.csv') # lab data
df_edi <- read_csv('data/SACSJ_delta_water_quality_1975-2021_updated.csv') # 2021 data ready for EDI without historic data added in (from (add_historical_data.Rmd))

# fix dates
df_lab <- df_lab %>% rename(Date = SampleDate, Station = StationCode)
df_edi$Date <- as.Date(df_edi$Date, format = '%m/%d/%Y')

df_old$Date <- mdy(df_old$Date)
df_final$Date <- ymd(df_final$Date)
df_lab$Date <- mdy(df_lab$Date)

# fix up lab data
keep_cols <- c('Organic Nitrogen (Total)','TSS','VSS','Organic Nitrogen (Dissolved)','TDS','Kjeldahl Nitrogen (Total)','Pheophytin a','Ortho-phosphate (Dissolved)','Chloride (Total)','Silica (SiO2) (Dissolved)','Chlorophyll a','Ammonia (Total)','Nitrite + Nitrate (Dissolved)','Ammonia (Dissolved)','Chloride (Dissolved)','Organic Carbon (Dissolved)')

df_lab <- df_lab %>% filter(ConstituentName %in% keep_cols)
df_lab$ResultPrefix[is.na(df_lab$ResultPrefix)] <- '='


## remove all depths != 3
df_lab$Depth <- as.numeric(df_lab$Depth)
df_lab <- df_lab %>% filter(Depth == 3)
```

# Initial Issues
- old data missing TSS/VSS/TDS data
- data starts 12/1997
```{r}
select(df_edi, c(Date, TSS)) %>% filter(is.na(TSS)) %>% arrange(Date)
```
- check lab data
-- does start earlier
```{r}
select(df_lab, c(Date, ConstituentName)) %>% filter(ConstituentName == 'TSS' & !is.na(ConstituentName)) %>% arrange(Date)
```
- check lab doesn't have any unique stations
```{r}
unique(df_lab$Station)[unique(df_lab$Station) %!in% unique(df_old$Station)]
```
- check for dupes in lab data
```{r}
df_dupes <- df_lab[duplicated(df_lab[1:6]),]
df_dupes <- df_dupes %>% rename(Result_D = Result, ResultPrefix_D = ResultPrefix)
col_check <- colnames(df_lab)[!colnames(df_lab) %in% c('ResultPrefix','Result')]
df_check <- left_join(df_dupes, df_lab, by = col_check)
df_check <- df_check %>% select(-ResultPrefix_D, -Result_D) %>% filter(Date <= '1998-01-01')

col_check
```


```{r}
df_check
```
Need to manually check 1978-02-15, otherwise good (removed from raw file input so not shown here)

Final clean (TDS):
```{r}
df_lab <- df_lab %>% pivot_wider(id_cols = c(Station, Date),
                                 names_from = ConstituentName, values_from = c(ResultPrefix, Result, ReportingLimit, UnitName))

df_add <- df_lab %>% filter(Date < '1997-12-04') %>% select(c(Date, Station, ResultPrefix_TDS, Result_TDS, ReportingLimit_TDS, UnitName_TDS))
df_add <- df_add %>% 
  rename(TDS = Result_TDS) %>%
  distinct(Date,Station,TDS, .keep_all = TRUE) %>%
  filter(!is.na(TDS))

df_final <- left_join(df_edi, df_add[c('Date','Station','TDS')], by = c('Date', 'Station'))
df_final <- df_final %>% rename(TDS = TDS.x)

df_final <- df_final %>%
  mutate(
    TDS = case_when(
      is.na(TDS) ~ TDS.y,
      !is.na(TDS) ~ TDS)
    ) %>%
  select(-TDS.y)
  

# check
df_check <- select(df_final, c(Date, Station, TDS)) %>% filter(is.na(TDS))
```

Final clean (TSS):
```{r}
df_add <- df_lab %>% filter(Date < '1997-12-04') %>% select(c(Date, Station, ResultPrefix_TSS, Result_TSS, ReportingLimit_TSS, UnitName_TSS))
df_add <- df_add %>% 
  rename(TSS = Result_TSS) %>%
  distinct(Date,Station,TSS, .keep_all = TRUE) %>%
  filter(!is.na(TSS))

# check for unique values; all good, so can quick add
print(unique(df_add$ResultPrefix_TSS))
print(unique(df_add$UnitName_TSS))
print(unique(df_add$ReportingLimit_TSS))

df_final <- left_join(df_final, df_add[c('Date','Station','TSS')], by = c('Date', 'Station'))
df_final <- df_final %>% rename(TSS = TSS.x)
df_final <- as.numeric(df_final$TSS)

df_final <- df_final %>%
  mutate(
    TSS.z = case_when(!is.na(TSS) ~ TSS,
                      TRUE ~ TSS.y)
    )
  

# check
df_check <- select(df_final, c(Date, Station, TSS)) %>% filter(is.na(TSS))

unique(df_final$TSS)
test <- df_final %>% filter(Date, Station, TSS, TSS.y, TSS_Sign)
```

