# Load Packages/Read in Data

```{r}
library(tidyverse)
source(here::here('00_GlobalFunctions/functions.R'))
```

Read in data from SharePoint
```{r}
# Count Data
df_count <- abs_path_emp('Benthic/AQUARIUS Samples Database/Database Migration/Raw Benthic Data/Count Data/COMPILED Benthic data 1975-2025 for EDI input.csv')

df_count <- read_csv(df_count, col_types = cols(.default = col_character()))

# Taxonomist Sample Data
df_samp <- abs_path_emp('Benthic/AQUARIUS Samples Database/Database Migration/Raw Benthic Data/Sample Data/COMPILED Benthic sample data 1975-2025.csv')

df_samp <- read_csv(df_samp, col_types = cols(.default = col_character()))

# Taxa Data
fp_codes <- abs_path_emp('Benthic/AQUARIUS Samples Database/Database Migration/Raw Benthic Data/Benthic organism keylist 2026.csv')

df_codes <- read_csv(fp_codes, col_types = cols(.default = col_character()))
```

# Format

Format dates
```{r}
df_samp <- df_samp %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y'))

df_count <- df_count %>%
  mutate(Date = as.Date(Date))
```

Add Main Station
```{r}
df_samp <- df_samp %>%
  mutate(MainStation = sub('-.*', '', Station))

df_count <- df_count %>%
  mutate(MainStation = sub('-.*', '', Station))
```


# Remove bad samples
```{r}
df_samp %>%
  filter((Date == '2023-10-10' & Station == 'D41A-C') |
           (Date == '2022-12-21' & Station == 'P8-R'))
```

```{r}
df_samp <- df_samp %>%
  filter(!((Date == '2023-10-10' & Station == 'D41A-C') |
           (Date == '2022-12-21' & Station == 'P8-R')))
```

# Check Codes

All codes are in keylist
```{r}
setdiff(df_count$OrganismCode, df_codes$OrganismCode)
```

Only one code per lab ID
```{r}
df_count %>% group_by(LabSampleNumber, OrganismCode) %>% filter(n() > 1)
```

Only one code per Date/Station/Grab
```{r}
df_count %>% group_by(Date, Station, Grab) %>% filter(duplicated(OrganismCode))
```

# Check multiple dates per station

```{r}
df_samp %>%
  distinct(Station, Year, Month, Date) %>%
  group_by(Station, Year, Month) %>%
  filter(n() > 1) %>%
  summarize(dates = paste(sort(Date), collapse = ', '), n = n(), .groups = 'drop')

df_count %>%
  distinct(Station, Year, Month, Date) %>%
  group_by(Station, Year, Month) %>%
  filter(n() > 1) %>%
  summarize(dates = paste(sort(Date), collapse = ', '), n = n(), .groups = 'drop')
```

```{r}
df_samp %>%
  filter(Date == '1988-04-01') %>%
  select(Station, Month) %>%
  distinct()
```

```{r}
# fix samp issues
df_samp <- df_samp %>%
  # Remove specified rows
  filter(!(Date == as.Date('2006-06-09') & Station == 'C9-L')) %>%
  filter(!(Date == as.Date('2006-08-08') & Station == 'C9-L')) %>%
  # Change Month from "April" to "March" for 1988-04-01 rows at specific stations
  mutate(Month = ifelse(Date == as.Date('1988-04-01') &
                          Station %in% c('D28A-L', 'D28A-R', 'D7-C', 'D4-R', 'D4-C','D4-L'),
                        'March', Month)) %>%
  # Change Date for specific (Date, Station, Grab) rows
  mutate(Date = case_when(
    Date == as.Date('2020-02-10') & Station == 'D41A-C' ~ as.Date('2020-02-11'),
    Date == as.Date('2024-11-14') & Station == 'D7-C' ~ as.Date('2024-11-13'),
    Date == as.Date('2025-02-11') & Station == 'D6-R' & Grab == 4 ~ as.Date('2025-02-12'),
    Date == as.Date('2025-11-19') & Station == 'D7-C' & Grab == 4 ~ as.Date('2025-11-18'),
    TRUE ~ Date
  ))
```

```{r}
# fix count dates
df_count <- df_count %>%
   mutate(Date = case_when(
     Date == as.Date('2020-02-10') & Station == 'D41A-C' ~ as.Date('2020-02-11'),
    Date == as.Date('2024-11-14') & Station == 'D7-C' ~ as.Date('2024-11-13'),
    Date == as.Date('2025-02-11') & Station == 'D6-R' & Grab == 4 ~ as.Date('2025-02-12'),
    Date == as.Date('2025-11-19') & Station == 'D7-C' & Grab == 4 ~ as.Date('2025-11-18'),
    TRUE ~ Date
  ))
```

check again
```{r}
df_samp %>%
  distinct(Station, Year, Month, Date) %>%
  group_by(Station, Year, Month) %>%
  filter(n() > 1) %>%
  summarize(dates = paste(sort(Date), collapse = ', '), n = n(), .groups = 'drop')

df_count %>%
  distinct(Station, Year, Month, Date) %>%
  group_by(Station, Year, Month) %>%
  filter(n() > 1) %>%
  summarize(dates = paste(sort(Date), collapse = ', '), n = n(), .groups = 'drop')
```

# check multiple stations per main station

Keep for now
```{r}
df_samp %>%
  distinct(MainStation, Station, Year, Month) %>%
  group_by(MainStation, Year, Month) %>%
  filter(n_distinct(Station) > 1) %>%
  summarise(stations = paste(sort(unique(Station)), collapse = ', '),
            n = n_distinct(Station),
            .groups = 'drop') %>%
  arrange(Year, Month, MainStation)
```

```{r}
df_count %>%
  distinct(MainStation, Station, Year, Month) %>%
  group_by(MainStation, Year, Month) %>%
  filter(n_distinct(Station) > 1) %>%
  summarise(stations = paste(sort(unique(Station)), collapse = ', '),
            n = n_distinct(Station),
            .groups = 'drop') %>%
  arrange(Year, Month, MainStation)
```

# Check Count/Samp

## check mismatched lab samples

```{r}
bind_rows(
  df_samp  %>% distinct(LabSampleNumber, Date, Station) %>% mutate(src = 'df_samp'),
  df_count %>% distinct(LabSampleNumber, Date, Station) %>% mutate(src = 'df_count')
) %>%
  group_by(LabSampleNumber, Station) %>%
  filter(n_distinct(Date) > 1) %>%
  ungroup() %>%
  pivot_wider(names_from = src, values_from = Date,
              values_fn = ~ paste(sort(unique(.x)), collapse = ', '))
```

Fix D6 count date

```{r}
df_count <- df_count %>%
  mutate(Date = if_else(Station == 'D6-R' & Date == as.Date('2023-01-20'),
                        as.Date('2023-01-19'), Date))
```


## Check if there's missing events

```{r}
# in count but not samp
df_miss_samp <- anti_join(
  df_count %>% distinct(Date, Station, Grab),
  df_samp  %>% distinct(Date, Station, Grab),
  by = c('Date', 'Station', 'Grab')
)

# in samp but not count
df_miss_count <- anti_join(
  df_samp  %>% distinct(Date, Station, Grab),
  df_count %>% distinct(Date, Station, Grab),
  by = c('Date', 'Station', 'Grab')
)

df_miss_samp
df_miss_count
```

```{r}
df_count %>%
  distinct(Date, Station, Grab, LabSampleNumber) %>%
  full_join(
    df_samp %>% distinct(Date, Station, Grab, LabSampleNumber),
    by = c('Date', 'Station', 'Grab'),
    suffix = c('_count', '_samp')
  ) %>%
  filter(LabSampleNumber_count != LabSampleNumber_samp |
         is.na(LabSampleNumber_count) | is.na(LabSampleNumber_samp))
```


Might be real?


# Format for Export

```{r}
target_cols <- c(
  'Observation ID', 'Location ID', 'Observed Property ID', 'Observed DateTime',
  'Analyzed DateTime', 'Depth', 'Depth Unit', 'Data Classification', 'Result Value',
  'Result Unit', 'Result Status', 'Result Grade', 'Medium', 'Activity Name',
  'Activity ID', 'Collection Method', 'Field: Device ID', 'Field: Device Type',
  'Field: Comment', 'Lab: Specimen Name', 'Lab: Analysis Method',
  'Lab: Detection Condition', 'Lab: Limit Type', 'Lab: MDL', 'Lab: MRL',
  'Lab: Quality Flag', 'Lab: Received DateTime', 'Lab: Prepared DateTime',
  'Lab: Sample Fraction', 'Lab: From Laboratory', 'Lab: Sample ID',
  'Lab: Dilution Factor', 'Lab: Comment', 'QC: Type', 'QC: Source Sample ID',
  'EA_Sample Event ID', 'EA_Month', 'EA_Year', 'EA_Taxa Sample ID',
  'EA_Grab Number', 'EA_Size Bin', 'EA_Main Station', 'EA_Station Position'
)

add_common <- function(df) {
  df %>% mutate(
    `Location ID` = Station,
    `Observed DateTime` = format(as.Date(Date), '%Y-%m-%d 12:00 [PST]'),
    `Analyzed DateTime` = '1950-01-01 12:00 [PST]',
    Depth = 3,
    `Depth Unit` = 'ft',
    `Data Classification` = 'FIELD_SURVEY',
    `Result Status` = 'Reviewed',
    `Result Grade` = 'Good',
    Medium = 'Water',
    `EA_Sample Event ID` = paste0(
      format(as.Date(Date), '%m%y'), '_',
      gsub('-', '', Station), '_01'
    ),
    EA_Month = Month,
    EA_Year = Year,
    `EA_Grab Number` = Grab,
    `EA_Main Station` = sub('-.*', '', `Location ID`),
    `EA_Station Position` = ifelse(grepl('-', `Location ID`),
                                   sub('.*-', '', `Location ID`),
                                   NA_character_)
  )
}

# format df_count
# Organism/Count rows — one per organism per grab
df_count_organism <- df_count %>%
  add_common() %>%
  mutate(
    `Activity Name` = paste('All: Taxa Grab', Grab),
    `Lab: Specimen Name` = OrganismCode
  ) %>%
  pivot_longer(
    cols = c(OrganismCode, Count),
    names_to = 'Observed Property ID',
    values_to = 'Result Value'
  ) %>%
  mutate(
    `Observed Property ID` = recode(`Observed Property ID`,
      'OrganismCode' = 'Organism Number',
      'Count'        = 'Count (Taxa All)'
    ),
    `Result Unit` = ifelse(`Observed Property ID` == 'Count (Taxa All)', 'individuals', NA),
    `Result Value` = as.character(`Result Value`)
  ) %>%
  filter(!is.na(`Result Value`))

# Taxa Sample ID rows — one per grab
df_count_sampleid <- df_count %>%
  distinct(Date, Station, Month, Year, Grab, LabSampleNumber) %>%
  add_common() %>%
  mutate(
    `Observed Property ID` = 'Taxa Sample ID',
    `Result Value` = as.character(LabSampleNumber),
    `Result Unit` = NA_character_,
    `Activity Name` = 'Metadata',
    `Lab: Specimen Name` = paste('Taxa Grab', Grab)
  ) %>%
  filter(!is.na(`Result Value`))

df_count_format <- bind_rows(df_count_organism, df_count_sampleid)

# format df_samp (sediment + comment + sample volume + subsample size + Taxa Sample ID rows)
sediment_map <- c(
  VegDetritus_Lab   = 'Vegetative Detritus (Taxa)',
  ShellFragment_Lab = 'Shell Fragment (Taxa)',
  Sand_Lab          = 'Sand (Taxa)',
  Peat_Lab          = 'Peat (Taxa)',
  Mica_Lab          = 'Mica (Taxa)',
  Gravel_Lab        = 'Gravel (Taxa)',
  HardClay_Lab      = 'Hard Clay (Taxa)',
  WormTubes_Lab     = 'Worm Tubes (Taxa)',
  Other_Lab         = 'Other (Taxa)'
)

df_samp_format <- df_samp %>%
  add_common() %>%
  mutate(
    Comment_Lab = ifelse(Comment_Lab == 'n/p', NA, Comment_Lab)
  ) %>%
  pivot_longer(
    cols = c(all_of(names(sediment_map)), Comment_Lab,
             BenthicSampleVol, SubSampleSize, LabSampleNumber),
    names_to = 'Observed Property ID',
    values_to = 'Result Value',
    values_transform = list(`Result Value` = as.character)
  ) %>%
  mutate(
    `Observed Property ID` = recode(`Observed Property ID`,
      !!!sediment_map,
      'Comment_Lab'      = 'Sediment Description (Taxa)',
      'BenthicSampleVol' = 'Sample Volume',
      'SubSampleSize'    = 'SubSample Size',
      'LabSampleNumber'  = 'Taxa Sample ID'
    ),
    `Activity Name` = ifelse(`Observed Property ID` %in% c('Sample Volume', 'SubSample Size', 'Taxa Sample ID'),
                             'Metadata', 'Sediment'),
    `Lab: Specimen Name` = paste('Taxa Grab', Grab),
    `Result Unit` = case_when(
      `Observed Property ID` == 'Sediment Description (Taxa)' ~ NA_character_,
      `Observed Property ID` == 'SubSample Size'              ~ NA_character_,
      `Observed Property ID` == 'Taxa Sample ID'              ~ NA_character_,
      `Observed Property ID` == 'Sample Volume'               ~ 'mL',
      TRUE                                                    ~ 'percent'
    )
  ) %>%
  filter(!is.na(`Result Value`))

missing_cols_count <- setdiff(target_cols, colnames(df_count_format))
df_count_format[missing_cols_count] <- NA
df_count_format <- df_count_format[, target_cols]

missing_cols_samp <- setdiff(target_cols, colnames(df_samp_format))
df_samp_format[missing_cols_samp] <- NA
df_samp_format <- df_samp_format[, target_cols]
```

## Check

all the lab sample numbers agree
```{r}
df_count_sampleids <- df_count %>%
  distinct(Date, Station, Grab, count_id = LabSampleNumber)

df_samp_sampleids <- df_samp %>%
  distinct(Date, Station, Grab, samp_id = LabSampleNumber)

comparison <- full_join(df_count_sampleids, df_samp_sampleids,
                        by = c('Date', 'Station', 'Grab')) %>%
  mutate(status = case_when(
    is.na(count_id) & !is.na(samp_id) ~ 'only in samp',
    !is.na(count_id) & is.na(samp_id) ~ 'only in count',
    count_id == samp_id                ~ 'agree',
    TRUE                               ~ 'disagree'
  ))

comparison %>% filter(status != 'agree')
```

all agree. can get rid of one in df_count

```{r}
df_count_format <- df_count_format %>%
  filter(`Observed Property ID` != 'Taxa Sample ID')
```


```{r}
df_count <- df_count_format
df_samp <- df_samp_format

save(df_count, file = here::here('02_Benthic/Aquarius/historical-imports/data/count_data.rda'))
save(df_samp, file = here::here('02_Benthic/Aquarius/historical-imports/data/samp_data.rda'))
```
