# Load Packages/Read in Data

```{r}
library(tidyverse)
source(here::here('00_GlobalFunctions/functions.R'))
```

Read in data from Sharepoint
```{r}
fp_raw <- abs_path_emp('Benthic/AQUARIUS Samples Database/Database Migration/Raw Benthic Data/Field Data/COMPILED Benthic field data 1975-2024.csv')

df_raw <- read_csv(fp_raw, col_types = cols(.default = col_character()))
```

# Initial Clean

Remove unneeded columns
```{r}
# saved in location data
df_field <- df_raw %>%
  select(-c(Survey, StationName, StationPosition))

# no data
df_field <- df_field %>%
  select(-starts_with('...'))
```

Check/Fix Stations
```{r}
unique(df_field$Station)

df_field %>%
  filter(Station == 'D41-A')


# fix D41-A to D41A-C
df_field <- df_field %>%
  mutate(Station = ifelse(Station == 'D41-A', 'D41A-C', Station))
```

Convert Date
```{r}
df_field$Date <- as.Date(df_field$Date, format = '%m/%d/%Y')

# unique(df_field$Date)
```

Fix bad date
```{r}
df_field <- df_field %>%
  mutate(Date = if_else(Month == 'September' & Year == 2024 & Station == 'D7-C',
                        as.Date('2024-09-12'), Date))
```



Remove row with NA date (not sampled)
```{r}
df_field <- df_field %>%
  filter(!is.na(Date))
```


# Ask Betsy

## Time_PST

Should it be rounded to nearest 5 minutes like WQ?

Fix '12:10 PST' to '12:10'. ALso '.' to ':'

```{r}
df_field <- df_field %>%
  mutate(
    Time_PST = case_when(
      Time_PST == '12:10 PST' ~ '12:10',
      TRUE ~ Time_PST
    )
  )

df_field <- df_field %>%
  mutate(Time_PST = gsub('\\.', ':', Time_PST))
```

Fix non-military

```{r}
df_field <- df_field %>%
  mutate(Time_PST = ifelse(
    !is.na(Time_PST) & as.numeric(sub(':.*', '', Time_PST)) < 3,
    sprintf('%d:%s',
            as.numeric(sub(':.*', '', Time_PST)) + 12,
            sub('^[^:]*:', '', Time_PST)),
    Time_PST
  ))
```


```{r}
unique(df_field$Time_PST)
```

```{r}
round_time <- function(x) {
  out <- rep(NA_character_, length(x))
  ok <- !is.na(x)
  out[ok] <- format(round_date(as_datetime(paste0('2000-01-01 ', x[ok], ':00')),
                               '5 minutes'), '%H:%M')
  out
}

df_field <- df_field %>% mutate(Time_PST = round_time(Time_PST))

unique(df_field$Time_PST)
```

## Tide

Should tide be saved? Would you default to 'high' after a certain point? Ignore for current data? 96% NA.

```{r}
df_field <- df_field %>%
  mutate(
    Tide = tolower(Tide),
    Tide = case_when(
      Tide == 'flooding' ~ 'flood',
      TRUE ~ Tide
    )
  )

df_field %>% count(Tide)
```

## Sampling Staff

Keep? Extended attribute? Part of sampling event? 72% NA.

```{r}
# remove \n's
df_field$SamplingStaff <- gsub('\\s*\n\\s*', ' ', df_field$SamplingStaff)
df_field$SamplingStaff <- gsub(' +', ' ', df_field$SamplingStaff)

# unique(df_field$SamplingStaff)
```


## WQ Data taken on benthic

Keep? Maybe at sampling event level? Or field metadata observation? 0% NA

```{r}
unique(df_field$WQ_data_taken_on_benthic)
```

## WQ benthic combined run 

Keep? Maybe sampling event level? Or field metadata observation? < 1 % NA (only one NA).

Set to "TRUE", "FALSE", and NA

```{r}
unique(df_field$WQ_benthic_combined_run)

df_field <- df_field %>%
  mutate(WQ_benthic_combined_run = case_when(
    WQ_benthic_combined_run == 'yes' ~ TRUE,
    WQ_benthic_combined_run == 'no' ~ FALSE,
    WQ_benthic_combined_run == 'unknown' ~ NA,
    TRUE ~ NA 
  ))
```

## DO percent

Lots of missing DO percent data, even when other data was taken:

```{r}
missing_doperc <- df_field %>% filter(is.na(DO_perc) & !is.na(DO_ugL))
```

## Missing Metadata

Missing things like preservative and ponar size when I'm fairly sure it should be known.

## Comments

Field Comments commonly seem to be Sediment Comments + Biota combined.

eg.:

```{r}
df_field %>%
  filter(Comments_Field == 'Sediment was observed to be primarily silt with some fine clay on bottom.  Samples mostly contained shell fragments and worms. One corbula and one isopod were also observed in the samples.')
```

```{r}
df_field %>%
  filter(Comments_Field == 'Sand and mica.Worms, Corbicula')
```

Also sometimes exactly the same as Sediment Description, eg.:

```{r}
df_field %>%
  filter(Comments_Field == 'Sediment was silt with sand and organics. Sediment was uniform throughout grab')
```

```{r}
df_field %>%
  filter(Comments_Field == 'Gray sand and tough peat, lots of amphipods, Corbicula (some small ~3 mm, a few v. large)')
```

Should this be addressed?

Either way, remove the newlines:

```{r}
df_field$Comments_Field <- gsub('\\s*\n\\s*', ' ', df_field$Comments_Field)
df_field$Comments_Field <- gsub(' +', ' ', df_field$Comments_Field)
df_field$SedDesc_Field <- gsub('\\s*\n\\s*', ' ', df_field$SedDesc_Field)
df_field$SedDesc_Field <- gsub(' +', ' ', df_field$SedDesc_Field)
```


# Metadata Check

## Collection Method

One odd value.

```{r}
unique(df_field$CollectionMethod)
```

Change F to PO and PO to 'Ponar Grab'

```{r}
df_field <- df_field %>%
  mutate(
    CollectionMethod = case_when(
      CollectionMethod == 'PO' | CollectionMethod == 'F' ~ 'Ponar Grab',
      TRUE ~ CollectionMethod
    )
  )
```

## Station, Ponar Size, Preservative, Year, Month

All look good; change 'FO' to formalin.

```{r}
unique(df_field$Station)
```

```{r}
unique(df_field$PonarSize)
```
fill in missing ponar size/preservative
```{r}
df_field <- df_field %>%
  mutate(
    PonarSize = ifelse(is.na(PonarSize), '0.052', PonarSize),
    Preservative = ifelse(is.na(Preservative), 'FO', Preservative)
  )
```


```{r}
unique(df_field$Preservative)

df_field <- df_field %>%
  mutate(
    Preservative = case_when(
      Preservative == 'FO' ~ 'formalin',
      TRUE ~ Preservative
    )
  )

unique(df_field$Preservative)
```



```{r}
unique(df_field$Year)
unique(df_field$Month)
```

# Sediment Check

Most seem plausible. However, a few don't add to 100 (or are NA).

```{r}
# no error; all good
df_field <- df_field %>%
  mutate(across(c(OrgMatter_Field, FineSand_Field, CoarseSand_Field,
                  Mica_Field, Gravel_Field, Clay_Field, Silt_Field), 
                as.numeric))

unique(df_field$OrgMatter_Field)
unique(df_field$FineSand_Field)
unique(df_field$CoarseSand_Field)
unique(df_field$Mica_Field)
unique(df_field$Gravel_Field)
unique(df_field$Clay_Field)
unique(df_field$Silt_Field)
```

Odd rows:

```{r}
df_field %>%
  mutate(row_sum = rowSums(pick(OrgMatter_Field, FineSand_Field, CoarseSand_Field,
                                Mica_Field, Gravel_Field, Clay_Field, Silt_Field), 
                           na.rm = TRUE)) %>%
  count(row_sum)
```


```{r}
odd_sed_counts <- df_field %>%
  mutate(row_sum = rowSums(pick(OrgMatter_Field, FineSand_Field, CoarseSand_Field,
                                Mica_Field, Gravel_Field, Clay_Field, Silt_Field), 
                           na.rm = TRUE)) %>%
  filter(!row_sum %in% c(0, 100))
```

Biota loook good.

```{r}
# unique(df_field$Biota)

# remove newline characters
df_field$Biota <- gsub('\\s*\n\\s*', ' ', df_field$Biota)
df_field$Biota <- gsub(' +', ' ', df_field$Biota)
```

# Analyte Check

Ones that are numeric are assumed okay (spot checked). Looking more into logical ones

```{r}
str(df_field)
```

## Temperature

Convert to numeric.

```{r}
df_field <- df_field %>%
  mutate(across(c(Temp_C), 
                as.numeric))

unique(df_field$Temp_C)
```

## pH

Correct two values (7073 and 707).

```{r}
df_field <- df_field %>%
  mutate(across(c(pH), as.numeric)) %>%
  mutate(pH = round(pH, 2)) %>%
  mutate(pH = case_when(
    pH == 7073 ~ 7.73,
    pH == 707 ~ 7.07,
    TRUE ~ pH
  ))

unique(df_field$pH)
```

## Turbidity

Fix a double decimal, round to 2 like DWQ

```{r}
df_field$Turb_FNU <- gsub('4\\.\\.3', '4.3', df_field$Turb_FNU)

df_field <- df_field %>%
  mutate(across(c(Turb_FNU), as.numeric)) %>%
  mutate(Turb_FNU = round(Turb_FNU, 2))

unique(df_field$Turb_FNU)
```
## Chla

Round to 2 like DWQ

```{r}
df_field <- df_field %>%
  mutate(across(c(Chla_mgL), as.numeric)) %>%
  mutate(Chla_mgL = round(Chla_mgL, 2))
```

## DO mg/L

Round to 2 like DWQ

```{r}
df_field <- df_field %>%
  mutate(across(c(DO_ugL), as.numeric)) %>%
  mutate(DO_ugL = round(DO_ugL,2))

unique(df_field$DO_ugL)
```

## DO %

Round to 1 like DWQ

```{r}
df_field <- df_field %>%
  mutate(across(c(DO_perc), as.numeric)) %>%
  mutate(DO_perc = round(DO_perc, 1))

unique(df_field$DO_perc)
```

## Water Depth

Round 1 like DWQ

```{r}
df_field <- df_field %>%
  mutate(across(c(WaterDepth_ft), as.numeric)) %>%
  mutate(WaterDepth_ft = round(WaterDepth_ft, 1))

unique(df_field$WaterDepth_ft)
```

## Air Temp

Round 1 like DWQ

```{r}
df_field <- df_field %>%
  mutate(across(c(AirTemp_F), as.numeric)) %>%
  mutate(AirTemp_F = round(AirTemp_F, 1))

unique(df_field$AirTemp_F)
```

## Wind

Check 'PC' value. Sky is 1, so a typo. Wind should be 15.5 here.

Also note that Rain (1 -> N), WaveScale (N -> 1), Water Depth (50.9 -> 7.5),
Air Temp (15.5 -> 50.9), DO percent (NA -> 89.8), and MC Score (NA -> 1) are wrong, so fixing those here as well.

```{r}
df_field %>%
  filter(Wind_mph == 'PC')
```

```{r}
df_field <- df_field %>%
  mutate(
    Wind_mph = case_when(
      Station == 'D7-C' & Date == '2022-01-05' ~ '15.5',
      TRUE ~ Wind_mph
    ),
    Sky = case_when(
      Station == 'D7-C' & Date == '2022-01-05' ~ 'PC',
      TRUE ~ Sky
    ),
    WaveScale = case_when(
      Station == 'D7-C' & Date == '2022-01-05' ~ '1',
      TRUE ~ WaveScale
    ),
    Rain = case_when(
      Station == 'D7-C' & Date == '2022-01-05' ~ 'N',
      TRUE ~ Rain
    ),
    MicrocystisScore = case_when(
      Station == 'D7-C' & Date == '2022-01-05' ~ '1',
      TRUE ~ MicrocystisScore
    ),
    AirTemp_F = case_when(
      Station == 'D7-C' & Date == '2022-01-05' ~ 50.9,
      TRUE ~ AirTemp_F
    ),
    WaterDepth_ft = case_when(
      Station == 'D7-C' & Date == '2022-01-05' ~ 7.5,
      TRUE ~ WaterDepth_ft
    ),
    DO_perc = case_when(
      Station == 'D7-C' & Date == '2022-01-05' ~ 89.8,
      TRUE ~ DO_perc
    )
  )
```

```{r}
df_field %>%
  filter(Station == 'D7-C' & Date == '2022-01-05')
```

Convert to numeric, round to 1 like DWQ

```{r}
df_field <- df_field %>%
  mutate(across(c(Wind_mph), as.numeric)) %>%
  mutate(Wind_mph = round(Wind_mph, 1))

unique(df_field$Wind_mph)
```

## SpC

Has commas, so fix that first. Round to nearest whole number as with DWQ.

```{r}
df_field$SC_uScm <- gsub(',', '', df_field$SC_uScm)
df_field$SC_uScm <- as.integer(df_field$SC_uScm)

unique(df_field$SC_uScm)
```

# Environment Variables

## Sky

D24 on 8/20/2024, change:
- Sky (2 -> PC)
- WaveScale (N -> 2)
- Rain (1 -> N)
- MC Score (NA -> 1)

```{r}
unique(df_field$Sky)

df_field %>%
  filter(Sky == '2')
```

```{r}
df_field <- df_field %>%
  mutate(
    Sky = case_when(
      Station == 'D24-L' & Date == '2024-08-20' ~ 'PC',
      TRUE ~ Sky
    ),
    WaveScale = case_when(
      Station == 'D24-L' & Date == '2024-08-20' ~ '2',
      TRUE ~ WaveScale
    ),
    Rain = case_when(
      Station == 'D24-L' & Date == '2024-08-20' ~ 'N',
      TRUE ~ Rain
    ),
    MicrocystisScore = case_when(
      Station == 'D24-L' & Date == '2024-08-20' ~ '1',
      TRUE ~ MicrocystisScore
    )
  )
```

```{r}
df_field %>%
  filter(Station == 'D24-L' & Date == '2024-08-20')
```

## WaveScale

Looks good.

```{r}
unique(df_field$WaveScale)
```

## Rain

Looks good.

```{r}
unique(df_field$Rain)
```
## Microcystis Score

Looks good.

```{r}
unique(df_field$MicrocystisScore)
```
# Plot numeric variables

```{r}
vars <- c("Temp_C", "SC_uScm", "pH", "Turb_FNU", "Chla_mgL", "DO_ugL", "DO_perc", "WaterDepth_ft", "AirTemp_F", "Wind_mph")

for (var in vars) {
  p <- df_field %>%
    ggplot(aes(x = "", y = .data[[var]])) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(alpha = 0.7, width = 0.2, size = 1, color = "steelblue") +
    theme_bw() +
    labs(x = NULL, y = var, title = var)
  
  print(p)
}
```

```{r}
df_field <- df_field %>%
  mutate(AirTemp_F = case_when(
    AirTemp_F == 798 ~ 79.8,
    TRUE ~ AirTemp_F
  ))

df_field <- df_field %>%
  mutate(Temp_C = case_when(
    Temp_C == 927 ~ NA,
    Temp_C == 3.34 ~ 9.34,
    TRUE ~ Temp_C
  ))

df_field <- df_field %>%
  mutate(SC_uScm = case_when(
    Station == 'D16-L' & Date == '2021-06-21' ~ 927,
    TRUE ~ SC_uScm
  ))

df_field <- df_field %>%
  mutate(SC_uScm = case_when(
    SC_uScm < 5 ~ NA,
    TRUE ~ SC_uScm
  ))
```

```{r}
vars <- c("Temp_C", "SC_uScm", "pH", "Turb_FNU", "Chla_mgL", "DO_ugL", "DO_perc", "WaterDepth_ft", "AirTemp_F", "Wind_mph")

for (var in vars) {
  p <- df_field %>%
    ggplot(aes(x = "", y = .data[[var]])) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(alpha = 0.7, width = 0.2, size = 1, color = "steelblue") +
    theme_bw() +
    labs(x = NULL, y = var, title = var)
  
  print(p)
}
```

```{r}
# zoomed SC_uScm
df_field %>%
  ggplot(aes(x = "", y = SC_uScm)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(alpha = 0.7, width = 0.2, size = 2, color = "steelblue") +
  coord_cartesian(ylim = c(0, 500)) +
  theme_bw()
```


```{r}
for (var in vars) {
  p <- df_field %>%
    mutate(Month = month(Date, label = TRUE),
           Year = factor(year(Date))) %>%
    filter(year(Date) >= 2005) %>%
    group_by(Year, Month) %>%
    summarise(has_data = any(!is.na(.data[[var]])), .groups = "drop") %>%
    ggplot(aes(x = Month, y = fct_rev(Year), fill = has_data)) +
    geom_tile(color = "white") +
    scale_fill_manual(values = c("TRUE" = "steelblue", "FALSE" = "grey90")) +
    theme_bw() +
    labs(title = var, fill = "Has Data", y = "Year")
  
  print(p)
}
```




# Rename columns

```{r}
names(df_field)
```



```{r}
df_field <- df_field %>%
  rename(
    'Air Temperature' = AirTemp_F,
    'Chlorophyll a' = Chla_mgL,
    'Sediment Description (Field)' = SedDesc_Field,
    'Water Temperature' = Temp_C,
    'Water Depth' = WaterDepth_ft,
    'Specific Conductance' = SC_uScm,
    'Turbidity' = Turb_FNU,
    'Dissolved Oxygen' = DO_ugL,
    'Dissolved Oxygen (percent)' = DO_perc,
    'Sky Conditions' = Sky,
    'MVI' = MicrocystisScore,
    'Wind Speed' = Wind_mph,
    'Wave Scale' = WaveScale,
    'Fine Sand (Field)' = FineSand_Field,
    'Coarse Sand (Field)' = CoarseSand_Field,
    'Gravel (Field)' = Gravel_Field,
    'Mica (Field)' = Mica_Field,
    'Clay (Field)' = Clay_Field,
    'Silt (Field)' = Silt_Field,
    'Organic Matter (Field)' = OrgMatter_Field,
    'Comments (Field)' = Comments_Field,
    'Ponar Size' = PonarSize
  )

# remove
df_field <- df_field %>% select(-c(WQ_data_taken_on_benthic))
```

```{r}
names(df_field)
```


```{r}
target_cols <- c(
  'Observation ID', 'Location ID', 'Observed Property ID', 'Observed DateTime',
  'Analyzed DateTime', 'Depth', 'Depth Unit', 'Data Classification', 'Result Value',
  'Result Unit', 'Result Status', 'Result Grade', 'Medium', 'Activity Name',
  'Activity ID', 'Collection Method', 'Field: Device ID', 'Field: Device Type',
  'Field: Comment', 'Lab: Specimen Name', 'Lab: Analysis Method',
  'Lab: Detection Condition', 'Lab: Limit Type', 'Lab: MDL', 'Lab: MRL',
  'Lab: Quality Flag', 'Lab: Received DateTime', 'Lab: Prepared DateTime',
  'Lab: Sample Fraction', 'Lab: From Laboratory', 'Lab: Sample ID',
  'Lab: Dilution Factor', 'Lab: Comment', 'QC: Type', 'QC: Source Sample ID',
  'EA_Sample Event ID', 'EA_Month', 'EA_Year', 'EA_Taxa Sample ID',
  'EA_Grab Number', 'EA_Size Bin', 'EA_Main Station', 'EA_Station Position'
)

# Property groups
metadata_field_props <- c('Ponar Size', 'Preservative', 'Water Depth',
                          'Sampling Staff', 'Combined Run', 'Time', 'Comments (Field)')
sediment_props <- c('Organic Matter (Field)', 'Fine Sand (Field)', 'Coarse Sand (Field)',
                    'Mica (Field)', 'Gravel (Field)', 'Clay (Field)', 'Silt (Field)',
                    'Sediment Description (Field)', 'Biota')
wq_props <- c('Water Temperature', 'Specific Conductance', 'pH', 'Turbidity',
              'Chlorophyll a', 'Dissolved Oxygen', 'MVI')
weather_props <- c('Air Temperature', 'Wind Speed', 'Sky Conditions', 'Wave Scale', 'Rain')

# Property -> Result Unit
unit_map <- c(
  'Ponar Size'                   = 'mm',
  'Preservative'                 = NA,
  'Water Depth'                  = 'ft',
  'Sampling Staff'               = NA,
  'Combined Run'                 = NA,
  'Time'                         = NA,
  'Organic Matter (Field)'       = 'percent',
  'Fine Sand (Field)'            = 'percent',
  'Coarse Sand (Field)'          = 'percent',
  'Mica (Field)'                 = 'percent',
  'Gravel (Field)'               = 'percent',
  'Clay (Field)'                 = 'percent',
  'Silt (Field)'                 = 'percent',
  'Sediment Description (Field)' = NA,
  'Comments (Field)'             = NA,
  'Biota'                        = NA,
  'Water Temperature'            = 'degC',
  'Specific Conductance'         = 'uS/cm',
  'pH'                           = 'unitless',
  'Turbidity'                    = 'NTU',
  'Chlorophyll a'                = 'mg/L',
  'Dissolved Oxygen'             = 'ug/L',
  'Air Temperature'              = 'degF',
  'Wind Speed'                   = 'mph',
  'Sky Conditions'               = NA,
  'Wave Scale'                   = NA,
  'Rain'                         = NA,
  'MVI'                          = NA
)

df_format <- df_field %>%
  mutate(
    `Location ID` = Station,
    `Observed DateTime` = paste0(format(as.Date(Date), '%Y-%m-%d'), ' 12:00 [PST]'),
    `Analyzed DateTime` = '1950-01-01 12:00 [PST]',
    Depth = 3,
    `Depth Unit` = 'ft',
    `Data Classification` = 'FIELD_SURVEY',
    `Result Status` = 'Reviewed',
    `Result Grade` = 'Good',
    Medium = 'Water',
    `EA_Sample Event ID` = paste0(
      format(as.Date(Date), '%m%y'), '_',
      gsub('-', '', Station), '_01'
    ),
    EA_Month = Month,
    EA_Year = Year,
    `EA_Main Station` = sub('-.*', '', `Location ID`),
    `EA_Station Position` = ifelse(grepl('-', `Location ID`),
                                   sub('.*-', '', `Location ID`),
                                   NA_character_)
  ) %>%
  rename(
    `Sampling Staff` = SamplingStaff,
    `Combined Run`   = WQ_benthic_combined_run,
    `Time`           = Time_PST
  ) %>%
  pivot_longer(
    cols = c(`Ponar Size`, `Preservative`, `Water Depth`,
             `Sampling Staff`, `Combined Run`, `Time`,
             `Organic Matter (Field)`, `Fine Sand (Field)`, `Coarse Sand (Field)`,
             `Mica (Field)`, `Gravel (Field)`, `Clay (Field)`, `Silt (Field)`,
             `Comments (Field)`, `Sediment Description (Field)`,
             `Biota`, `Water Temperature`, `Specific Conductance`, `pH`,
             `Turbidity`, `Chlorophyll a`, `Dissolved Oxygen`, `Air Temperature`,
             `Wind Speed`, `Sky Conditions`, `Wave Scale`, `Rain`, `MVI`),
    names_to = 'Observed Property ID',
    values_to = 'Result Value',
    values_transform = list(`Result Value` = as.character)
  ) %>%
  mutate(
    `Activity Name` = case_when(
      `Observed Property ID` %in% sediment_props        ~ 'Sediment',
      `Observed Property ID` %in% metadata_field_props  ~ 'Metadata',
      TRUE                                              ~ 'Field Data'
    ),
    `Lab: Specimen Name` = case_when(
      `Activity Name` == 'Sediment'                    ~ 'Field',
      `Observed Property ID` %in% metadata_field_props ~ 'Field',
      `Observed Property ID` %in% wq_props             ~ 'Water Quality',
      `Observed Property ID` %in% weather_props        ~ 'Weather',
      TRUE                                             ~ NA_character_
    ),
    `Collection Method` = case_when(
      `Observed Property ID` == 'Comments (Field)'      ~ NA_character_,
      `Observed Property ID` %in% weather_props         ~ NA_character_,
      `Observed Property ID` %in% metadata_field_props  ~ NA_character_,
      `Observed Property ID` == 'MVI'                   ~ NA_character_,
      `Lab: Specimen Name` == 'Water Quality'           ~ 'YSI Sonde',
      TRUE                                              ~ 'Ponar Grab'
    ),
    `Result Unit` = unname(unit_map[`Observed Property ID`])
  ) %>%
  filter(!is.na(`Result Value`))

# Add any missing cols as NA
missing_cols <- setdiff(target_cols, colnames(df_format))
df_format[missing_cols] <- NA
df_format <- df_format[, target_cols]
```

```{r}
df_field <- df_format

save(df_field, file = here::here('02_Benthic/Aquarius/historical-imports/data/field_data.rda'))
```

