# Load Packages/Read in Data

```{r}
library(tidyverse)
source(here::here('00_GlobalFunctions/functions.R'))
options(scipen = 999)
```

Read in data from SharePoint
```{r}
# Sediment Data
fp_raw <- abs_path_emp('Benthic/AQUARIUS Samples Database/Database Migration/Raw Benthic Data/Bin Weight Data/COMPILEDlive_sort_biomass_for_database.csv')

df_raw <- read_csv(fp_raw, col_types = cols(.default = col_character()))
```

# Format

```{r}
df_ls <- df_raw %>%
  mutate(across(c(Dry_pan_mass, Pan_mass, Count, Ash_pan_mass), as.numeric))
```


```{r}
df_ls <- df_ls  %>%
  mutate(Date = as.Date(Date, format = '%Y-%m-%dT%H:%M:%SZ'))
```

fix date
```{r}
df_ls <- df_ls %>%
  mutate(Date = if_else(Year == 2024 & Month == 'May' & Station == 'D7' & Date == as.Date('2024-05-23'),
                        as.Date('2024-05-20'), Date))
```


## issues

multiple events per Date/Station/Species/Size bin; flag
```{r}
issues <- df_ls %>%
  group_by(Date, Station, Species_name, Size_class) %>%
  filter(n() > 1) %>%
  ungroup() %>%
  arrange(Date, Station, Species_name, Size_class)
```

also multiple no clams (some have comments)
```{r}
df_ls %>%
  filter(Species_name == 'No clams') %>%
  group_by(Date, Station) %>%
  filter(n() > 1) %>%
  ungroup() %>%
  arrange(Date, Station)
```

remove the rows without comments (or just the duplciate)
```{r}
no_clams_deduped <- df_ls %>%
  filter(Species_name == 'No clams') %>%
  arrange(desc(!is.na(Comments) & Comments != '')) %>%  # comment-having rows first
  distinct(Date, Station, .keep_all = TRUE)

df_ls <- df_ls %>%
  filter(Species_name != 'No clams') %>%
  bind_rows(no_clams_deduped)
```

and for now, just add a letter after for the others
```{r}
df_ls <- df_ls %>%
  group_by(Date, Station, Species_name, Size_class) %>%
  mutate(
    .n = n(),
    .idx = row_number(),
    Size_class = ifelse(.n > 1,
                        paste0(Size_class, letters[.idx]),
                        Size_class)
  ) %>%
  ungroup() %>%
  select(-.n, -.idx)
```

## format (again)

```{r}
df_ls <- df_ls %>%
  rename(
    'Count (Live Clams)' = Count,
    'Pan Mass' = Pan_mass,
    'Dried Pan Mass' = Dry_pan_mass,
    'Ashed Pan Mass' = Ash_pan_mass,
    'Size Bin' = Size_class
  ) %>%
  mutate(
    `Total Clam Mass` = round(`Dried Pan Mass` - `Ashed Pan Mass`, 4),
    `Average Clam Mass` = round(`Total Clam Mass` / `Count (Live Clams)`, 4),
    `Pan Mass` = round(`Pan Mass`, 3),
    `Dried Pan Mass` = round(`Dried Pan Mass`, 3),
    `Ashed Pan Mass` = round(`Ashed Pan Mass`, 3),
  )

df_ls <- df_ls %>%
  mutate(
    Species_name = case_when(
      Species_name == 'Corbicula fluminea' ~ '6730',
      Species_name == 'Potamocorbula amurensis' ~ '6890',
      TRUE ~ Species_name
    )
  )
```

TODO: Are no record and no clams the same??

```{r}
df_ls %>%
  filter(Species_name == 'No record')
```

```{r}
df_ls <- df_ls %>%
  filter(!(Species_name == 'No record'))
```

Mass should be NA if no clams

```{r}
df_ls <- df_ls %>%
  mutate(
    across(
      c(`Pan Mass`, `Dried Pan Mass`, `Ashed Pan Mass`, `Total Clam Mass`, `Average Clam Mass`),
      ~ ifelse(Species_name == 'No clams', NA, .)
    ),
    `Count (Live Clams)` = ifelse(Species_name == 'No clams', 0, `Count (Live Clams)`)
  )
```

# Fix Station
```{r}
df_ls <- df_ls %>%
  mutate(Station = case_when(
    Date == as.Date('2011-12-07') & Station == 'D28' ~ 'D28A',
    Month == 'November' & Year == '2025' & Station == 'P9' ~ 'P8',
    Date == as.Date('2016-03-10') & Station == 'D14' ~ 'D4',
    TRUE ~ Station
  ))
```


# Fix Dates

```{r}
df_ls <- df_ls %>%
  mutate(Date = case_when(
    Date == as.Date('2019-08-14') & Station == 'D24'  ~ as.Date('2019-08-15'),
    Date == as.Date('2026-01-20') & Station == 'D24'  ~ as.Date('2026-01-21'),
    Date == as.Date('2016-05-18') & Station == 'D28A' ~ as.Date('2016-05-19'),
    Date == as.Date('2020-02-11') & Station == 'D4'   ~ as.Date('2020-02-10'),
    Date == as.Date('2017-10-05') & Station == 'D41'  ~ as.Date('2017-10-04'),
    Date == as.Date('2019-11-13') & Station == 'D6'   ~ as.Date('2019-11-12'),
    Date == as.Date('2019-10-16') & Station == 'D6'   ~ as.Date('2019-10-15'),
    Date == as.Date('2023-08-16') & Station == 'D6'   ~ as.Date('2023-08-15'),
    Date == as.Date('2025-10-21') & Station == 'D6'   ~ as.Date('2025-10-20'),
    Date == as.Date('2008-11-19') & Station == 'D7'   ~ as.Date('2008-11-18'),
    Date == as.Date('2024-05-23') & Station == 'D7'   ~ as.Date('2024-05-20'),
    TRUE ~ Date
  ))
```

## remove for now

```{r}
df_ls <- df_ls %>%
  filter(!(Date == as.Date('2008-10-01') & Station %in% c('D16', 'D41', 'D41A')))
```

```{r}
df_ls %>%
  filter(Month == 'November', Year == '2025', Station %in% c('P8','C9','P9'))
```


# Format for Export

```{r}
target_cols <- c(
  'Observation ID', 'Location ID', 'Observed Property ID', 'Observed DateTime',
  'Analyzed DateTime', 'Depth', 'Depth Unit', 'Data Classification', 'Result Value',
  'Result Unit', 'Result Status', 'Result Grade', 'Medium', 'Activity Name',
  'Activity ID', 'Collection Method', 'Field: Device ID', 'Field: Device Type',
  'Field: Comment', 'Lab: Specimen Name', 'Lab: Analysis Method',
  'Lab: Detection Condition', 'Lab: Limit Type', 'Lab: MDL', 'Lab: MRL',
  'Lab: Quality Flag', 'Lab: Received DateTime', 'Lab: Prepared DateTime',
  'Lab: Sample Fraction', 'Lab: From Laboratory', 'Lab: Sample ID',
  'Lab: Dilution Factor', 'Lab: Comment', 'QC: Type', 'QC: Source Sample ID',
  'EA_Sample Event ID', 'EA_Month', 'EA_Year', 'EA_Taxa Sample ID',
  'EA_Grab Number', 'EA_Size Bin', 'EA_Main Station', 'EA_Station Position'
)

# Column -> Observed Property ID + Result Unit
prop_units <- list(
  Species_name         = list(prop = 'Organism Number',     unit = NA),
  `Count (Live Clams)` = list(prop = 'Count (Live Clams)',  unit = 'individuals'),
  `Pan Mass`           = list(prop = 'Pan Mass',            unit = 'g'),
  `Dried Pan Mass`     = list(prop = 'Dried Pan Mass',      unit = 'g'),
  `Ashed Pan Mass`     = list(prop = 'Ashed Pan Mass',      unit = 'g'),
  `Total Clam Mass`    = list(prop = 'Total Clam Mass',     unit = 'g'),
  `Average Clam Mass`  = list(prop = 'Average Clam Mass',   unit = 'g')
)
prop_map <- sapply(prop_units, function(x) x$prop)
unit_map <- sapply(prop_units, function(x) x$unit)

# Expand 'No_clams' into two rows (6730 + 6890), Count=0, Size Bin='None'
no_clams <- df_ls %>%
  filter(Species_name == 'No clams')

if (nrow(no_clams) > 0) {
  no_clams_expanded <- no_clams %>%
    slice(rep(1:n(), each = 2)) %>%
    mutate(
      Species_name = rep(c('6730', '6890'), times = nrow(no_clams)),
      `Count (Live Clams)` = 0,
      `Size Bin` = 'None'
    )
  df_ls <- df_ls %>%
    filter(Species_name != 'No clams') %>%
    bind_rows(no_clams_expanded)
}

# format
df_format <- df_ls %>%
  mutate(
    `Location ID` = Station,
    `Observed DateTime` = paste0(format(as.Date(Date), '%Y-%m-%d'), ' 12:00 [PST]'),
    `Analyzed DateTime` = '1950-01-01 12:00 [PST]',
    Depth = 3,
    `Depth Unit` = 'ft',
    `Data Classification` = 'FIELD_SURVEY',
    `Result Status` = 'Reviewed',
    `EA_Grab Number` = 'Live Sort',
    Medium = 'Water',
    `Activity Name` = paste0(Species_name, ': Live Sort'),
    `Collection Method` = 'Ponar Grab',
    `Lab: Specimen Name` = paste('Size Bin', `Size Bin`),
    `Field: Comment` = Comments,
    `EA_Sample Event ID` = paste0(
      format(as.Date(Date), '%m%y'), '_',
      gsub('-', '', Station), '_01'
    ),
    EA_Month = Month,
    EA_Year = Year,
    `EA_Size Bin` = `Size Bin`,
    `Result Grade` = case_when(
      QC_flag == 'y' & QC_exclude == 'y' ~ 'Bad',
      QC_flag == 'y' & QC_exclude == 'n' ~ 'Suspect',
      QC_flag == 'n' & QC_exclude == 'y' ~ 'Bad',
      QC_flag == 'n' & QC_exclude == 'n' ~ 'Good'
    ),
    `EA_Main Station` = sub('-.*', '', `Location ID`),
    `EA_Station Position` = ifelse(grepl('-', `Location ID`),
                                   sub('.*-', '', `Location ID`),
                                   NA_character_)
  ) %>%
  pivot_longer(
    cols = c(Species_name, `Count (Live Clams)`, `Pan Mass`, `Dried Pan Mass`,
             `Ashed Pan Mass`, `Total Clam Mass`, `Average Clam Mass`),
    names_to = '_src_col',
    values_to = 'Result Value',
    values_transform = list(`Result Value` = as.character)
  ) %>%
  mutate(
    `Observed Property ID` = unname(prop_map[`_src_col`]),
    `Result Unit` = unname(unit_map[`_src_col`])
  ) %>%
  select(-`_src_col`) %>%
  filter(!is.na(`Result Value`))

# Add any missing cols as NA
missing_cols <- setdiff(target_cols, colnames(df_format))
df_format[missing_cols] <- NA
df_format <- df_format[, target_cols]
```

```{r}
issues <- df_format %>%
  group_by(`EA_Sample Event ID`, `EA_Grab Number`, `EA_Size Bin`,
           `Location ID`, `Observed DateTime`, `Activity Name`,
           `Observed Property ID`) %>%
  filter(n() > 1) %>%
  ungroup() %>%
  arrange(`EA_Sample Event ID`, `Observed DateTime`, `Location ID`,
          `EA_Grab Number`, `EA_Size Bin`, `Activity Name`,
          `Observed Property ID`)

issues
```

```{r}
df_ls %>%
  group_by(Date, Station, Species_name, `Size Bin`) %>%
  filter(n() > 1) %>%
  ungroup() %>%
  arrange(Date, Station, Species_name, `Size Bin`)
```


```{r}
df_ls <- df_format

save(df_ls, file = here::here('02_Benthic/Aquarius/historical-imports/data/ls_data.rda'))
```
