# Load Packages/Read in Data

```{r}
library(tidyverse)
source(here::here('00_GlobalFunctions/functions.R'))
```

Read in data from SharePoint
```{r}
# Sediment Data
df_raw <- abs_path_emp('Benthic/AQUARIUS Samples Database/Database Migration/Raw Benthic Data/Sediment Data/COMPILED DWR EMP Sediment Data 1975-2022.csv')

df_raw <- read_csv(df_raw, col_types = cols(.default = col_character()))

# Missing Data
df_miss <- abs_path_emp('Benthic/AQUARIUS Samples Database/Database Migration/Raw Benthic Data/Metadata/benthic_missing_data.csv')

df_miss <- read_csv(df_miss, col_types = cols(.default = col_character()))
```

# Format

```{r}
load(here::here('02_Benthic/Aquarius/historical-imports/data/field_data.Rda'))
```

```{r}
df_field %>%
  filter(!is.na(`Observed DateTime`)) %>%
  distinct(EA_Month, EA_Year, `Location ID`, `Observed DateTime`) %>%
  count(EA_Month, EA_Year, `Location ID`) %>%
  filter(n > 1)
```

Check stations

```{r}
sort(unique(df_raw$Station))
```

# Dates

## Field/Sed Dates

```{r}
field_dates <- df_field %>%
  mutate(Field_Date = as.Date(`Observed DateTime`)) %>%
  group_by(`Location ID`, EA_Month, EA_Year) %>%
  filter(n_distinct(Field_Date) == 1) %>%
  distinct(
    Station = `Location ID`,
    Month = month(Field_Date, label = TRUE, abbr = FALSE),
    Year = as.character(year(Field_Date)),
    Field_Date
  ) %>%
  ungroup()

raw_dates <- df_raw %>%
  mutate(
    Raw_Date = as.Date(Date),
    Month = month(Raw_Date, label = TRUE, abbr = FALSE) |> as.character(),
    Year = as.character(year(Raw_Date))
  ) %>%
  group_by(Station, Month, Year) %>%
  filter(n_distinct(Raw_Date) == 1) %>%
  distinct(Station, Month, Year, Raw_Date) %>%
  ungroup()

field_dates %>%
  inner_join(raw_dates, by = c('Station', 'Month', 'Year')) %>%
  filter(Field_Date != Raw_Date) %>%
  arrange(Year, Month, Station)

date_mismatches <- field_dates %>%
  inner_join(
    raw_dates,
    by = c('Station', 'Month', 'Year')
  ) %>%
  filter(Field_Date != Raw_Date) %>%
  arrange(Station) %>%
  select(-c(Month, Year, `Location ID`))

date_mismatches
```

```{r}
date_updates <- field_dates %>%
  mutate(
    Month_Join = month(Field_Date, label = TRUE, abbr = FALSE) %>% 
      as.character(),
    Year_Join = year(Field_Date)
  ) %>%
  inner_join(
    df_raw %>%
      mutate(
        Raw_Date = as.Date(Date),
        Month_Join = month(Raw_Date, label = TRUE, abbr = FALSE) %>%
          as.character(),
        Year_Join = year(Raw_Date)
      ) %>%
      group_by(Station, Month_Join, Year_Join) %>%
      filter(n_distinct(Raw_Date) == 1) %>%
      distinct(Station, Month_Join, Year_Join, Raw_Date) %>%
      ungroup(),
    by = c('Station', 'Month_Join', 'Year_Join')
  ) %>%
  select(Station, Month_Join, Year_Join, Field_Date)

df_raw <- df_raw %>%
  mutate(
    Raw_Date = as.Date(Date),
    Month_Join = month(Raw_Date, label = TRUE, abbr = FALSE) %>%
      as.character(),
    Year_Join = year(Raw_Date)
  ) %>%
  left_join(
    date_updates,
    by = c('Station', 'Month_Join', 'Year_Join')
  ) %>%
  mutate(
    Date = coalesce(as.character(Field_Date), Date)
  ) %>%
  select(-Raw_Date, -Month_Join, -Year_Join, -Field_Date)
```

## Missing Dates

check which are missing
```{r}
missing <- df_raw %>%
  filter(is.na(Date)) %>%
  count(Month, Year, Station, name = 'n_rows') %>%
  inner_join(
    df_field %>%
      filter(!is.na(`Observed DateTime`)) %>%
      mutate(Date = as.Date(`Observed DateTime`)) %>%
      distinct(EA_Month, EA_Year, `Location ID`, Date),
    by = c('Month' = 'EA_Month', 'Year' = 'EA_Year', 'Station' = 'Location ID')
  ) %>%
  rename(Field_Date = Date)

missing
```

```{r}
missing %>% filter(n_rows > 1)
```

remove one for now
```{r}
target_rows <- which(df_raw$Month == 'April' & df_raw$Year == '2022' &
                     df_raw$Station == 'D7-C' & is.na(df_raw$Date))
df_raw <- df_raw[-target_rows[2], ]
```


```{r}
date_lookup <- df_field %>%
  filter(!is.na(`Observed DateTime`)) %>%
  mutate(Date = as.character(as.Date(`Observed DateTime`))) %>%
  distinct(EA_Month, EA_Year, `Location ID`, Date)

date_lookup %>% count(EA_Month, EA_Year, `Location ID`) %>% filter(n > 1)

df_raw <- df_raw %>%
  left_join(date_lookup,
            by = c('Month' = 'EA_Month', 'Year' = 'EA_Year', 'Station' = 'Location ID'),
            suffix = c('', '_fill')) %>%
  mutate(Date = coalesce(Date, Date_fill)) %>%
  select(-Date_fill)

df_raw %>% filter(is.na(Date))
```


```{r}
df_raw %>% filter(
  Year == 2022, Month == 'February', is.na(Station)
)
```

## Fix Manual Dates

```{r}
df_raw <- df_raw %>%
  mutate(
    Date = case_when(
      Year == 2022 & Month == 'February' & is.na(Station) ~ '2022-02-09',
      TRUE ~ Date
    ),
    Station = case_when(
      Date == '2022-02-09' & is.na(Station) ~ 'P8-R',
      TRUE ~ Station
    )
  )
```

## All 100 Sieves
Usually missing data

```{r}
df_hund <- df_raw %>%
  filter(
    if_all(starts_with('Sieve'), ~ .x == 100),
    is.na(Percent_Organic)
  ) %>%
  arrange(Date)
```

## Duplicates (no 100s)

```{r}
df_dups <- df_raw %>%
  mutate(
    Month_Join = month(Date),
    Year_Join = year(Date)
  ) %>%
  group_by(Station, Month, Month_Join, Year_Join) %>%
  filter(n() > 1) %>%
  ungroup()

df_dups_100 <- df_dups %>%
  group_by(Station, Month, Month_Join, Year_Join) %>%
  filter(
    any(
      if_all(starts_with('Sieve'), ~ .x == 100) &
        is.na(Percent_Organic)
    )
  ) %>%
  ungroup() %>%
  arrange(Year_Join, Month_Join, Station)

df_dups_not_100 <- df_dups %>%
  anti_join(
    df_dups_100 %>%
      distinct(Station, Month, Month_Join, Year_Join),
    by = c('Station', 'Month', 'Month_Join', 'Year_Join')
  ) %>%
  arrange(Year_Join, Month_Join, Station)

field_lookup <- df_field %>%
  transmute(
    Station = `Location ID`,
    Date_Join = as.Date(`Observed DateTime`)
  ) %>%
  distinct()

df_dups_not_100 <- df_dups_not_100 %>%
  mutate(Date_Join = as.Date(Date)) %>%
  left_join(
    field_lookup %>%
      mutate(has_field_data = TRUE),
    by = c('Station', 'Date_Join')
  ) %>%
  mutate(
    has_field_data = coalesce(has_field_data, FALSE)
  ) %>%
  select(-Date_Join) %>%
  relocate(has_field_data, .before = 1)

df_dups_not_100
```


## Fix Months

```{r}
df_raw <- df_raw %>%
  mutate(Month = ifelse(Date == as.Date('1988-04-01') &
                          Station %in% c('D11-C', 'D19-C', 'D28A-L', 'D28A-R',
                                               'D4-C', 'D4-L', 'D4-R', 'D7-C'),
                        'March', Month))
```

# Check Missing Month-Year
Further check if they correspond to field data

```{r}
df_missing_month_year <- df_raw %>%
  filter(is.na(Month) | is.na(Year)) %>%
  mutate(Date_Join = as.Date(Date)) %>%
  left_join(
    df_field %>%
      transmute(
        Station = `Location ID`,
        Date_Join = as.Date(`Observed DateTime`),
        has_field_data = TRUE
      ) %>%
      distinct(),
    by = c('Station', 'Date_Join')
  ) %>%
  mutate(
    has_field_data = coalesce(has_field_data, FALSE)
  ) %>%
  select(-Date_Join) %>%
  relocate(has_field_data, .before = 1) %>%
  arrange(Date, Station)

df_missing_month_year
```

Update Month-Year for ones with Field Data
```{r}
field_lookup <- df_field %>%
  transmute(
    Station = `Location ID`,
    Date_Join = as.Date(`Observed DateTime`),
    EA_Month,
    EA_Year
  ) %>%
  distinct()

df_raw <- df_raw %>%
  mutate(
    Date_Join = as.Date(Date),
    missing_month_year = is.na(Month) | is.na(Year)
  ) %>%
  left_join(
    field_lookup,
    by = c('Station', 'Date_Join')
  ) %>%
  mutate(
    Month = if_else(
      missing_month_year & !is.na(EA_Month),
      EA_Month,
      Month
    ),
    Year = if_else(
      missing_month_year & !is.na(EA_Year),
      EA_Year,
      Year
    )
  ) %>%
  select(-Date_Join, -missing_month_year, -EA_Month, -EA_Year)
```

Check sieve logic
```{r}
sieve_cols <- c(
  'Sieve76200', 'Sieve38100', 'Sieve19050', 'Sieve09525',
  'Sieve04750', 'Sieve02360', 'Sieve01180', 'Sieve00600',
  'Sieve00300', 'Sieve00150', 'Sieve00075'
)

df_sieve_flag <- df_raw %>%
  rowwise() %>%
  mutate(
    sieve_order_flag = {
      x <- suppressWarnings(
        as.numeric(c_across(all_of(sieve_cols)))
      )
      
      any(x[-1] > x[-length(x)], na.rm = TRUE)
    }
  ) %>%
  ungroup() %>%
  filter(sieve_order_flag) %>%
  select(
    Date,
    Station,
    all_of(sieve_cols)
  )

df_sieve_flag
```

```{r}
unique(df_raw$Sieve76200)
```


# Format for Export
```{r}
target_cols <- c(
  'Observation ID', 'Location ID', 'Observed Property ID', 'Observed DateTime',
  'Analyzed DateTime', 'Depth', 'Depth Unit', 'Data Classification', 'Result Value',
  'Result Unit', 'Result Status', 'Result Grade', 'Medium', 'Activity Name',
  'Activity ID', 'Collection Method', 'Field: Device ID', 'Field: Device Type',
  'Field: Comment', 'Lab: Specimen Name', 'Lab: Analysis Method',
  'Lab: Detection Condition', 'Lab: Limit Type', 'Lab: MDL', 'Lab: MRL',
  'Lab: Quality Flag', 'Lab: Received DateTime', 'Lab: Prepared DateTime',
  'Lab: Sample Fraction', 'Lab: From Laboratory', 'Lab: Sample ID',
  'Lab: Dilution Factor', 'Lab: Comment', 'QC: Type', 'QC: Source Sample ID',
  'EA_Sample Event ID', 'EA_Month', 'EA_Year', 'EA_Taxa Sample ID',
  'EA_Grab Number', 'EA_Size Bin', 'EA_Main Station', 'EA_Station Position'
)

# col -> obs property map
prop_map <- c(
  'Percent_Organic'      = 'Organic Matter (Lab)',
  'Sediment_Description' = 'Sediment Description (Lab)'
)

df_format <- df_raw %>%
  mutate(
    `Location ID` = Station,
    `Observed DateTime` = paste0(format(as.Date(Date), '%Y-%m-%d'), ' 12:00 [PST]'),
    `Analyzed DateTime` = '1950-01-01 12:00 [PST]',
    Depth = 3,
    `Depth Unit` = 'ft',
    `Data Classification` = 'FIELD_SURVEY',
    `Result Status` = 'Reviewed',
    `Result Grade` = 'Good',
    Medium = 'Water',
    `Activity Name` = 'Sediment',
    `Collection Method` = 'Ponar Grab',
    `Lab: Specimen Name` = 'Lab',
    `EA_Sample Event ID` = paste0(
      format(as.Date(Date), '%m%y'), '_',
      gsub('-', '', Station), '_01'
    ),
    EA_Month = Month,
    EA_Year = Year,
    `EA_Grab Number` = 'Sediment',
    `EA_Main Station` = sub('-.*', '', `Location ID`),
    `EA_Station Position` = ifelse(grepl('-', `Location ID`),
                                   sub('.*-', '', `Location ID`),
                                   NA_character_)
  ) %>%
  pivot_longer(
    cols = c(Sieve76200, Sieve38100, Sieve19050, Sieve09525, Sieve04750,
             Sieve02360, Sieve01180, Sieve00600, Sieve00300, Sieve00150,
             Sieve00075, Percent_Organic, Sediment_Description),
    names_to = 'Observed Property ID',
    values_to = 'Result Value',
    values_transform = list(`Result Value` = as.character)
  ) %>%
  mutate(
    `Observed Property ID` = ifelse(`Observed Property ID` %in% names(prop_map),
                                    unname(prop_map[`Observed Property ID`]),
                                    `Observed Property ID`),
    `Result Unit` = ifelse(`Observed Property ID` == 'Sediment Description (Lab)', NA, 'percent')
  ) %>%
  filter(!is.na(`Result Value`))

# add cols
missing_cols <- setdiff(target_cols, colnames(df_format))
df_format[missing_cols] <- NA
df_format <- df_format[, target_cols]
```


```{r}
df_sed <- df_format

save(df_sed, file = here::here('02_Benthic/Aquarius/historical-imports/data/sed_data.rda'))
```
