
# Load packages/read in data

```{r}
library(tidyverse)
library(lubridate)
source(here::here('00_GlobalFunctions/functions.R'))

rda_files <- list.files('data', pattern = '\\.rda$', full.names = TRUE, ignore.case = TRUE)

for (f in rda_files) load(f, envir = .GlobalEnv)
```

```{r}
# load missing sampling events
fp_dates <- abs_path_emp('Benthic/AQUARIUS Samples Database/Database Migration/Raw Benthic Data/Metadata/benthic_not_sampled.csv')

df_dates <- read_csv(fp_dates)
```

```{r}
dfs <- list(df_clamlen, df_count, df_samp, df_field, df_ls, df_sed)

dfs <- lapply(dfs, function(df) {
  df %>% mutate(
    Date = as.Date(`Observed DateTime`),
    MainStation = sub('-.*', '', `Location ID`)
  )
})

list2env(setNames(dfs, c('df_clamlen', 'df_count', 'df_samp', 'df_field', 'df_ls', 'df_sed')),
         envir = .GlobalEnv)

dfs <- list(df_clamlen = df_clamlen, df_count = df_count, df_samp = df_samp,
            df_field = df_field, df_ls = df_ls, df_sed = df_sed)
```

## Split live sort data

main data vs grts (number stations vs not)

```{r}
df_grts <- df_ls %>% filter(grepl('^[0-9]+$', `Location ID`))
df_mls  <- df_ls %>% filter(!grepl('^[0-9]+$', `Location ID`))
```

```{r}
df_sed %>%
  filter(is.na(EA_Year) | is.na(EA_Month)) %>%
  select(`Location ID`, Date, EA_Year, EA_Month) %>%
  distinct()
```


remove these for now, easier to compare
```{r}
df_fullsed <- df_sed %>%
  filter(!(is.na(EA_Year) | is.na(EA_Month)))
```

```{r}
dfs <- list(df_clamlen = df_clamlen, df_count = df_count, df_samp = df_samp,
            df_field = df_field, df_ls = df_mls, df_sed = df_sed)
```

# LS stations

base on field data
```{r}
# Build field lookup: unique Location ID per (EA_Month, EA_Year, Date, MainStation)
field_lookup <- df_field %>%
  distinct(EA_Month, EA_Year, Date, MainStation, `Location ID`)

# any (EA_Month, EA_Year, Date, MainStation) with >1 Location ID in df_field?
ambiguous <- field_lookup %>%
  group_by(EA_Month, EA_Year, Date, MainStation) %>%
  filter(n() > 1) %>%
  arrange(EA_Year, EA_Month, MainStation, Date)

# Rows in df_mls with no matching event in df_field
no_field_match <- df_mls %>%
  distinct(EA_Month, EA_Year, Date, MainStation, `Location ID`) %>%
  anti_join(df_field %>% distinct(EA_Month, EA_Year, Date, MainStation),
            by = c('EA_Month', 'EA_Year', 'Date', 'MainStation'))

date_disagree <- df_mls %>%
  distinct(EA_Month, EA_Year, MainStation, mls_Date = Date) %>%
  inner_join(
    df_field %>% distinct(EA_Month, EA_Year, MainStation, field_Date = Date),
    by = c('EA_Month', 'EA_Year', 'MainStation')
  ) %>%
  filter(mls_Date != field_Date) %>%
  arrange(EA_Year, EA_Month, MainStation)

ambiguous
no_field_match
date_disagree
```

check dates per main station that disagree with each other/df_ls

```{r}
events <- imap_dfr(dfs, ~ .x %>%
                     distinct(EA_Month, EA_Year, MainStation, Date) %>%
                     mutate(df = .y)) %>%
  mutate(df = factor(df, levels = names(dfs)))

date_disagree <- events %>%
  group_by(EA_Month, EA_Year, MainStation) %>%
  filter(n_distinct(Date) > 1) %>%
  ungroup() %>%
  pivot_wider(names_from = df, values_from = Date,
              values_fn = ~ paste(sort(unique(.x)), collapse = ', ')) %>%
  arrange(EA_Year, EA_Month, MainStation)

date_disagree
```

```{r}

```



```{r}
df_samp %>%
  filter(Date == '1988-04-01') %>%
  distinct(MainStation)

df_samp %>%
  filter(Date == '1988-04-28') %>%
  distinct(MainStation)
```

```{r}
df_sed %>%
  filter(Date == '1988-04-28') %>%
  distinct(MainStation)
```




for non-ls dfs, check for any discrepancies that would affect ls dates (none after corrections)

```{r}
# Compute agreed date per row from non-ls cols
analyzed <- date_disagree %>%
  rowwise() %>%
  mutate(
    other_dates = list(unique(unlist(
      str_split(c(df_clamlen, df_count, df_samp, df_field, df_sed), ',\\s*')
    ))),
    other_dates = list(other_dates[!is.na(other_dates) & other_dates != '']),
    agreed_date = ifelse(length(other_dates) == 1, other_dates[[1]], NA_character_)
  ) %>%
  ungroup() %>%
  select(-other_dates) %>%
  mutate(disposition = case_when(
    is.na(df_ls)         ~ 'NA ls',
    is.na(agreed_date)   ~ 'dfs disagree',
    df_ls == agreed_date ~ 'all good',
    TRUE                 ~ 'will be fixed'
  ))

fixes <- analyzed %>%
  filter(disposition == 'will be fixed') %>%
  select(EA_Month, EA_Year, MainStation, from_date = df_ls, to_date = agreed_date)

unfixed <- analyzed %>%
  filter(disposition != 'will be fixed') %>%
  select(EA_Month, EA_Year, MainStation, df_clamlen, df_count, df_samp,
         df_field, df_ls, df_sed, disposition)

df_mls <- df_mls %>%
  mutate(EA_Year_chr = as.character(EA_Year)) %>%
  left_join(
    fixes %>% mutate(EA_Year = as.character(EA_Year),
                     from_date = as.Date(from_date),
                     to_date   = as.Date(to_date)),
    by = c('EA_Month', 'EA_Year_chr' = 'EA_Year',
           'MainStation', 'Date' = 'from_date')
  ) %>%
  mutate(Date = coalesce(to_date, Date)) %>%
  select(-EA_Year_chr, -to_date)

dfs <- list(df_clamlen = df_clamlen, df_count = df_count, df_samp = df_samp,
            df_field = df_field, df_ls = df_mls, df_sed = df_fullsed)

fixes
unfixed
```

# Update LS stations

no disagreement in stations (with L/R/C) since live sort started, so can set to others

```{r}
events_loc <- imap_dfr(dfs[setdiff(names(dfs), 'df_ls')],
                       ~ .x %>%
                         distinct(EA_Month, EA_Year, MainStation, `Location ID`) %>%
                         mutate(df = .y)) %>%
  mutate(df = factor(df, levels = setdiff(names(dfs), 'df_ls')))

loc_disagree <- events_loc %>%
  group_by(EA_Month, EA_Year, MainStation) %>%
  filter(n_distinct(`Location ID`) > 1) %>%
  ungroup() %>%
  pivot_wider(names_from = df, values_from = `Location ID`,
              values_fn = ~ paste(sort(unique(.x)), collapse = ', ')) %>%
  arrange(EA_Year, EA_Month, MainStation)

loc_disagree
```

```{r}
# Pool Location IDs from all dfs except df_ls
other_locs <- imap_dfr(dfs[setdiff(names(dfs), 'df_ls')],
                       ~ .x %>%
                         distinct(Date, MainStation, `Location ID`) %>%
                         mutate(df = .y))

# Per (Date, MainStation), unique Location IDs across the other dfs
loc_lookup <- other_locs %>%
  group_by(Date, MainStation) %>%
  summarise(
    locs = list(sort(unique(`Location ID`))),
    n_locs = n_distinct(`Location ID`),
    .groups = 'drop'
  )

conflicts <- loc_lookup %>%
  filter(n_locs > 1) %>%
  semi_join(df_mls %>% distinct(Date, MainStation), by = c('Date', 'MainStation')) %>%
  mutate(locations = map_chr(locs, ~ paste(.x, collapse = ', '))) %>%
  select(Date, MainStation, locations)

clean_lookup <- loc_lookup %>%
  filter(n_locs == 1) %>%
  mutate(new_LocationID = map_chr(locs, 1)) %>%
  select(Date, MainStation, new_LocationID)

# Apply to df_mls
df_mls <- df_mls %>%
  left_join(clean_lookup, by = c('Date', 'MainStation')) %>%
  mutate(`Location ID` = coalesce(new_LocationID, `Location ID`)) %>%
  select(-new_LocationID)

conflicts

unmatched <- df_mls %>%
  distinct(Date, MainStation, `Location ID`) %>%
  anti_join(loc_lookup, by = c('Date', 'MainStation'))

unmatched

clean_lookup_mls <- loc_lookup %>%
  filter(n_locs == 1) %>%
  semi_join(df_mls %>% distinct(Date, MainStation), by = c('Date', 'MainStation')) %>%
  mutate(new_LocationID = map_chr(locs, 1)) %>%
  select(Date, MainStation, new_LocationID)

clean_lookup_mls
```

also fix 2026, checking there's only one option per main station

```{r}
location_map <- map_dfr(dfs[setdiff(names(dfs), 'df_ls')],
        ~ .x %>%
          filter(EA_Year %in% c(2023,2024,2025)) %>%
          distinct(MainStation, `Location ID`)) %>%
  distinct() %>%
  arrange(MainStation, `Location ID`) %>%
  rename(new_LocationID = `Location ID`)

location_map
```

```{r}
unchanged <- df_mls %>%
  filter(EA_Year == 2026) %>%
  distinct(MainStation, `Location ID`) %>%
  anti_join(location_map, by = 'MainStation')

# Apply the update
df_mls <- df_mls %>%
  left_join(location_map, by = 'MainStation') %>%
  mutate(`Location ID` = if_else(EA_Year == 2026 & !is.na(new_LocationID),
                                 new_LocationID, `Location ID`)) %>%
  select(-new_LocationID)
```

re-update station position
```{r}
df_mls <- df_mls %>%
  mutate(
    `EA_Main Station` = sub('-.*', '', `Location ID`),
    `EA_Station Position` = ifelse(grepl('-', `Location ID`),
                                   sub('.*-', '', `Location ID`),
                                   NA_character_)
  )
```


update main dfs
```{r}
dfs <- list(df_clamlen = df_clamlen, df_count = df_count, df_samp = df_samp,
            df_field = df_field, df_ls = df_mls, df_sed = df_fullsed)
```

```{r}
map_dfr(dfs, ~ .x %>% select(`Location ID`)) %>%
  count(`Location ID`, name = 'n', sort = TRUE)
```

check all stations have coordinates

```{r}
fp_stations <- abs_path_emp('Benthic/AQUARIUS Samples Database/Metadata/Station_Coords_Benthic.csv')

df_stations <- read_csv(fp_stations)
```

```{r}
all_locs <- map_dfr(dfs, ~ .x %>% select(`Location ID`))

in_dfs_not_stations <- all_locs %>%
  anti_join(df_stations, by = 'Location ID')

in_stations_not_dfs <- df_stations %>%
  anti_join(all_locs, by = 'Location ID')

in_dfs_not_stations
in_stations_not_dfs
```


# Dates


## check no sampling

None should have data where sampling didn't occur

```{r}
flagged <- lapply(dfs, function(df) {
  df %>% filter(Date %in% df_dates$Date)
})

sapply(flagged, nrow)
```

# check multiple stations per "main station"

```{r}
multi_station <- imap_dfr(dfs, ~ .x %>%
  distinct(Date, MainStation, `Location ID`) %>%
  group_by(Date, MainStation) %>%
  filter(n() > 1) %>%
  ungroup() %>%
  mutate(df = .y))

multi_station
```


```{r}
cross_df_mismatch <- imap_dfr(dfs, ~ .x %>%
    distinct(Date, MainStation, `Location ID`) %>%
    mutate(df = .y)) %>%
  group_by(Date, MainStation) %>%
  filter(n_distinct(`Location ID`) > 1) %>%
  ungroup() %>%
  arrange(Date, MainStation, df)

cross_df_mismatch
```


# check mismatched Dates

```{r}
dfs <- lapply(dfs, function(df) {
  df %>% mutate(
    Date = as.Date(`Observed DateTime`),
    MainStation = sub('-.*', '', `Location ID`)
  )
})

station_combos <- imap_dfr(dfs, ~ .x %>%
                             distinct(Date, MainStation, `Location ID`) %>%
                             mutate(df = .y)) %>%
  mutate(df = factor(df, levels = names(dfs))) %>%
  pivot_wider(names_from = df, values_from = `Location ID`,
              values_fn = ~ paste(sort(unique(.x)), collapse = ', ')) %>%
  relocate(Date, MainStation)

disagreements <- station_combos %>%
  rowwise() %>%
  filter({
    vals <- c_across(c(df_clamlen, df_count, df_samp, df_field, df_ls, df_sed))
    n_distinct(vals[!is.na(vals)]) > 1
  }) %>%
  ungroup()

disagreements
```

# check missing seds
```{r}
combos <- imap_dfr(dfs, ~ .x %>%
                     distinct(Date, `Location ID`, EA_Month, EA_Year) %>%
                     mutate(df = .y)) %>%
  mutate(present = TRUE,
         df = factor(df, levels = names(dfs))) %>%
  pivot_wider(names_from = df, values_from = present, values_fill = FALSE) %>%
  relocate(EA_Month, EA_Year, Date, `Location ID`)

mismatches <- combos %>%
  filter(!if_all(all_of(names(dfs)), ~ replace_na(.x, FALSE)))

mismatches
```

```{r}
sed_missing <- mismatches %>%
  filter(!df_sed) %>%
  select(Date, EA_Month, EA_Year, `Location ID`)

df_field %>%
  filter(`Observed Property ID` == 'Comments (Field)') %>%
  inner_join(sed_missing, by = c('Date', 'EA_Month', 'EA_Year', 'Location ID')) %>%
  select(Date, EA_Month, EA_Year, `Location ID`, `Result Value`) %>%
  mutate(has_no_sed_note = grepl('no sediment', `Result Value`, ignore.case = TRUE))
```

## check IDs

```{r}
issues <- df_mls %>%
  group_by(`EA_Sample Event ID`, `EA_Grab Number`, `EA_Size Bin`,
           `Location ID`, `Observed DateTime`, `Activity Name`,
           `Observed Property ID`) %>%
  filter(n() > 1) %>%
  ungroup() %>%
  arrange(`EA_Sample Event ID`, `Observed DateTime`, `Location ID`,
          `EA_Grab Number`, `EA_Size Bin`, `Activity Name`,
          `Observed Property ID`)

issues
```

```{r}
unique(issues$`Observed Property ID`)
```

# heat maps


```{r}
heat_data <- imap_dfr(dfs, ~ .x %>%
                       distinct(`Location ID`, Date) %>%
                       mutate(Year = year(Date), Month = month(Date), df = .y)) %>%
  distinct(df, Year, Month, `Location ID`) %>%
  mutate(present = TRUE) %>%
  pivot_wider(names_from = df, values_from = present, values_fill = FALSE) %>%
  pivot_longer(all_of(names(dfs)), names_to = 'df', values_to = 'present') %>%
  group_by(df, Year, Month) %>%
  summarise(pct_present = mean(present) * 100, .groups = 'drop')

for (d in unique(heat_data$df)) {
  p <- ggplot(heat_data %>% filter(df == d),
              aes(Month, Year, fill = pct_present)) +
    geom_tile(color = 'black', linewidth = 0.5) +
    scale_fill_viridis_c(limits = c(0, 100), name = '% present') +
    scale_x_continuous(breaks = 1:12) +
    ggtitle(d) +
    theme_minimal()
  print(p)
  
  ggsave(
    filename = paste0('C:/Users/sperry/Desktop/', d, '_coverage.jpg'),
    plot = p,
    width = 8, height = 6, dpi = 300
  )
}
```

```{r}
combos %>%
  filter(year(Date) %in% c(2007), month(Date) %in% c(7)) %>%
  rowwise() %>%
  mutate(any_present = any(c_across(all_of(names(dfs))), na.rm = TRUE)) %>%
  ungroup() #%>%
  #filter(!any_present)
```


# Export

```{r}
df_final <- bind_rows(dfs) %>%
  select(-MainStation, Date)
```

fix activity names sigh
```{r}
df_final <- df_final %>%
  mutate(`Activity Name` = paste0(
    `Activity Name`, ' (',
    gsub('-', '', `Location ID`),
    sprintf('%02d', match(EA_Month, month.name)),
    substr(EA_Year, 3, 4),
    ')'
  ))
```

and woops fix Data Classification

```{r}
df_final <- df_final %>%
  mutate(`Data Classification` = ifelse(
    grepl('Metadata|Field Data|Sediment', `Activity Name`),
    'LAB',
    `Data Classification`
  ))
```



```{r}
cur_export <- df_final %>%
  filter(EA_Year == 2022, EA_Month %in% c('March'))#,'February','March'))#,'April','May','June'))
```

```{r}
write_csv(cur_export, 'C:/Users/sperry/Desktop/cur_ben.csv', na = '')
```

